### 2017-01-18 writing/outline.Rmd
added initial version of outline and manuscript

---
title: "Your time series is (probably) too short"
author: "Easton R. White"
date: "January 18, 2017"
output: pdf_document
---

```{r,echo=F,include=F,eval=F}
# General manuscript notes here
```

# Outline

## Titles: "Your time series is (probably) too short" "Minimum time series calculations for population inferences"
## Length: about 150 words for the abstract, 2500 words, not including a short abstract, not more than 20 references, and a total of 3 figures, tables, and/or panels

## Outline

## Abstract (150 words)

## Introduction to problem (600 words)

- ~~Have longer data sets now, Short funding cycles~~
- Little work has empirically investigated how long time series need to be
- Power series
- Empirically answer this question
- What determines minimum time series
    - Deterministic skeleton (dynamics), stochasticity, question being asked, biology of organism, management and conservation practices, power and alpha
- Past work that has looked at this question
- Population dynamics database
- What we plan to do

## Methods (750 words)
- How to subsample
- Theoretical results (maybe in the appendix mostly)
- Min time series calculation for each question
- Database I will be using

## Results (500 words)
- Simulated data with cycles and other types of attractors where we know what time series length you need to sample
- 1 good case study example (maybe this is in methods?) 
   - Isle Royale
- Results from X number of species

## Discussion and generality (800 words)

## Conclusions

## Figures/tables

http://www.frontiersinecology.org/fron/instructions-to-authors/research-communications/