### 2017-07-07 writing/supp_mat_draft.Rmd
cleaned up supp mat

---
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: true
fontsize: 12pt
geometry: margin=1in
bibliography: White_bib.bib
header-includes: 
    \usepackage{float}
    \renewcommand{\thepage}{S\arabic{page}} 
    \renewcommand{\thesection}{S\arabic{section}}  
    \renewcommand{\thetable}{S\arabic{table}}  
    \renewcommand{\thefigure}{S\arabic{figure}}
    \usepackage[round]{natbib} \bibpunct[; ]{(}{)}{,}{a}{}{;}
---

\vspace{2cm}

\begin{center}
 \textbf{Supplementary Material: Article title here}

Easton R. White$^{*}$
\vspace{3 mm}

Address: \\ \emph{Center for Population Biology, University of California-Davis, Davis, California 95616 USA}

*Corresponding author: eawhite@ucdavis.edu

June 22, 2017
 \end{center}

\vspace{2cm}

\tableofcontents

\vspace{1cm}

Code for all the figures and tables can be found at (https://github.com/erwhite1).

\vspace{2cm}

\clearpage

# Detailed example of subsampling and power calculations

```{r,echo=F}
 setwd("~/Desktop/Research/time-series-project")
 load("cleaned-data/cleaned_timeseries_database_with_min_time_linear_regression.Rdata")
    pop = subset(long_dat,long_dat$ID=='7558')
    model_example = lm(pop$popvalue~c(1:nrow(pop)))
```

Here, we illustrate how we performed the subsampling and power calculations for a specific population. We provide more detail here than in the main manuscript. As an example, we examine a 35-year time series of Bigeye tuna (\emph{Thunnus obesus}), one species in the Global Population Dynamics Database [@GPDD2010]. <!--Simple linear regression indicates a significant increase for this population with an estimated intercept of `r as.character(round(as.numeric(coef(model_example)[1])))` and a slope coefficient of `r as.numeric(coef(model_example)[2])`.--> We assume that this significant increase over 35 years is in fact the ``true trend". In statistical jargon, the 35-year trend is an effect that is actually present, we can reject the null hypothesis of no trend. We can then use this as a benchmark to see if subsamples of the time series also indicate a significant increase.

We first extract all contiguous subsamples of the time series. This leads to 34 two-year subsamples, 33 two-year subsamples, and so forth until a single 35-year subsample. We can call each set of subsamples, of the same length, a set. For each subsample, we conduct linear regression and extract model coefficients and p-values. Then, the fraction of subsamples within a set that show significant trends (significant slope coefficient) is the statistical power. It is important to note that we only consider subsamples to be significant if they are significant in the same direction as the complete 35-year time series. In other words, we are conducting a one-tailed test.

We can then plot statistical power as a function of time series length (Fig. \ref{fig:empirical_approach_example}b). As expected, we can see that power increases with the more years that are sampled. Then, we determine an appropriate level of statistical power that we find acceptable. Traditionally, this has been at 0.8, however, this is purely historical. Statistical power of 0.8 implies if a true trend is present, or a real change in abundance, then we will detect this trend 0.8 fraction of the time.

<!-- example of how subsampling and empirical approach works -->
```{r, echo=F,message=F,warning=F,fig.cap='(a) Population size of Bigeye tuna (\\emph{Thunnus obesus}) over time. The line is the best fit line from linear regresssion. (b) Statistical power for different subsets of the time series in panel a.\\label{fig:empirical_approach_example}'}
# Example of subsampling empirical data
 setwd("~/Desktop/Research/time-series-project")
 load("cleaned-data/cleaned_timeseries_database_with_min_time_linear_regression.Rdata")
  par(mfrow=c(1,2),oma=c(4,4,0.5,0.5))
    pop = subset(long_dat,long_dat$ID=='7558') #13192 for Harp seals, 7558 for Thunnus obesus
    plot(pop$year,pop$popvalue,cex.axis=1.2,cex.lab=1.2,ylab='',xlab='time (years)',pch=16,las=1)
    abline(lm(pop$popvalue~pop$year),lwd=2,col='red')
    mtext(text = '(a)',side = 3,line = -1.2,adj=0.95,font = 2,cex=1.2)
    mtext(text='population size', side=2,line=4,cex=1.2)
# could insert all possible regressions of length n
source('scripts/calculate_power_metric.R')
    power_example=calculate_power_for_subsamples(pop$popvalue,0.05)
    plot(power_example,cex.axis=1.2,cex.lab=1.2,ylab='statistical power',xlab='years sampled',las=1,pch=16)
    mtext(text = '(b)',side = 3,line = -1.2,adj=0.05,font = 2,cex=1.2)
# could also insert standard error bars
    abline(h=0.8,col='red',lwd=2,lty=2)
    abline(v=head(which(power_example>0.8),1),col='red',lwd=2,lty=2)

```

With statistical power of 0.8, we then determine the minimum time series length ($T_{min}$) required to achieve that level of statistical power. Here, $T_{min}$ is the first point in Fig. \ref{fig:empirical_approach_example}b where following points are also above 0.8. In this example, $T_{min}$ is `r head(which(power_example>0.8),1)`. Therefore, a minimum of 22 years of continuous monitoring are required to have 0.8 statistical power in determining long-term changes in abundance.

# Additional results from the main manuscript

## Statistical model of minimum time required and time series characteristics

In the main text, we explained how the minimum time required strongly correlated with the trend strength, temporal autocorrelation, and variance in population size. Here, we use a generalized linear model framework with a Poisson error structure to determine drivers of the minimum time required. In figure, \ref{fig:poisson_model} we show a set of residual plots for the regression. We then show the coefficient estimates and levels of significance in table \ref{table:model_output}.

```{r poisson_model, echo=F,include=T,fig.cap='Output of generalized linear model with a Poisson error structure for predicting the minimum time required with explanatory variables of the absolute value of the slope coefficient (or trend strength), temporal autocorrelation, and variability in population size.\\label{fig:poisson_model}'}
setwd("~/Desktop/Research/time-series-project")
load("cleaned-data/cleaned_timeseries_database_with_min_time_linear_regression.Rdata")

# GLM with Poisson error structure (include all terms - it is best not to include abs_overall_trend with variance as they are strongly corrrelated to one another)
pop_info$abs_overall_trend = abs(pop_info$overall_trend)
poisson_model <-glm(min_time_for_power~ abs_overall_trend + autocorrelation + variance,family = 'poisson',data=pop_info)

  par(mfrow=c(2,2),mar=c(5,5,1,2),oma=c(0,0,0.5,0))
    plot(poisson_model,cex.lab=1.2,las=1,cex.axis=1.2,main=' ',pch=16, col=rgb(0.5,0.5,0.5,0.4))
#R^2 calculation (pg. 218 Zuur et al. 2009 mixed effects book)
R2_fullmodel <- 100*(poisson_model$null.deviance - poisson_model$deviance)/(poisson_model$null.deviance)

poisson_model_no_variance <-glm(min_time_for_power~ abs_overall_trend + autocorrelation,family = 'poisson',data=pop_info)
R2_no_variance <- 100*(poisson_model_no_variance$null.deviance - poisson_model_no_variance$deviance)/(poisson_model_no_variance$null.deviance)

poisson_model_no_trend <-glm(min_time_for_power~ autocorrelation + variance,family = 'poisson',data=pop_info)
R2_no_trend <- 100*(poisson_model_no_trend$null.deviance - poisson_model_no_trend$deviance)/(poisson_model_no_trend$null.deviance)

```

```{r,echo=F,message=F,warning=F}

require(knitr)
knitr::kable(summary(poisson_model)$coef,caption = 'Output of generalized linear model to examine predictors of the minimum time required for determing long-term population trends.\\label{table:model_output}')

```

Our model with trend strength, autocorrelation, variance, and generation length accounted for `r round(R2_fullmodel,2)`% of the variation in the minimum time required (Table \ref{table:model_output}). However, we also found trend strength and variance to be strongly correlated with one another. Therefore, we ran two additional models with either trend strength or variance, but not both together. This resulted in lower explained deviance (analogue to $R^2$) of `r round(R2_no_variance,2)`% and `r round(R2_no_trend,2)`%, respectively.

\pagebreak

## Minimum time required and biological correlates
```{r, echo=F}
setwd("~/Desktop/Research/time-series-project")
load('cleaned-data/combined_databases_with_min_time_linear_regression.Rdata')
```

In the main manuscript, we examined the minimum time required to detect a significant trend in abundance over time using linear regression. As detailed in the main manuscript the minimum time required was around 15, but there was a wide distribution. Therefore, we were interested in potential explanatory variables of the minimum time required. In the main manuscript, we examined characteristics of the time series itself, like variability, autocorrelation, and the trend in abundance over time. Here, we combined our time series database with a database on life history characteristics of amniotes from @Myhrvold2015. There was life history information available for `r nrow(LPI_pop_info)` populations representing `r length(table(LPI_pop_info$Binomial))` different species, all of which were in the Aves class.

We then correlated minimum time required for each population with its corresponding life history characteristics. In figure \ref{fig:biological_correlates} we examined minimum time required versus generation length (years), litter size (n), adult body mass (grams), maximum longevity (years), egg mass (grams), and incubation (days). None of these variables had much explanatory power in accounting for the variance in the minimum time required. see table \ref{table:model_output}}

```{r, echo=F, fig.cap='Mimimum time required versus (a) generation length (years), (b) litter size (n), (c) adult body mass (grams), (d) maximum longevity (years), (e) egg mass (grams), and (f) incubation (days). The lines in each plot represent the best fit line from linear regression.\\label{fig:biological_correlates}',fig.height=6}
setwd("~/Desktop/Research/time-series-project")
load('cleaned-data/combined_databases_with_min_time_linear_regression.Rdata')
par(mfrow=c(3,2),oma=c(5,4,1,1),mar=c(5,3,0,0))

n=1
x_axis_name_list = c('generation length (years)','litter size (n)','adult body mass (grams)','maximum longevity (years)','egg mass (grams)','incubation (days)')
for (x_axis_name in names(LPI_pop_info)[21:26]){
  plot(LPI_pop_info[[x_axis_name]],LPI_pop_info$min_time_for_power,pch=16,col=rgb(0.5,0.5,0.5,0.5),las=1,ylab='',xlab='',cex.axis=1.2)
  abline(lm(LPI_pop_info$min_time_for_power~LPI_pop_info[[x_axis_name]]),lwd=3)

  mtext(text = x_axis_name_list[n],side = 1,line = 3,cex = 1.2)
  mtext(text = paste('(',letters[n],')',sep=''),side = 3,line=-1.5,adj = 0.98,cex=1.2)
  n=n+1
}

mtext(text = 'minimum time required',side = 2,line = 2,cex = 1.4,outer = TRUE)

```

\pagebreak

## Minimum time series and biological class

When we examined differences in the minimum time required there were similar patterns with biological classes. For instance, more time is required for species within the Actinopterygii class compared to other species (Fig. \ref{fig:class}a). These differences between biological classes can be explained by differences in population variability, with species in Actinopterygii have larger variability in population size from year-to-year.

```{r,echo=F,fig.cap='(a) Minimum time required to estimate change in abundance for species class, (b) long-term trend (estimated slope coefficient) by species class, (c) interannual variability in population size by species class, and (d) temporal autocorrelation by species class.\\label{fig:class}',fig.height=6}

# Pull in data
setwd("~/Desktop/Research/time-series-project")
load("~/Desktop/Research/time-series-project/cleaned-data/cleaned_timeseries_database_with_min_time_linear_regression.Rdata")
# Biological correlates plots
par(mfrow=c(2,2),oma=c(8,1,1.5,0.5),mar=c(0,4,0,1))

boxplot(pop_info$min_time_for_power~pop_info$class, lwd = 1, ylab = 'minimum time required',col='white',cex.lab=1.4,cex.axis=1.2,las=2,xaxt="n")
mtext(text = '(a)',side = 3,line = -1.2,adj=0.95,font = 2,cex=1.2)
stripchart(pop_info$min_time_for_power~pop_info$class, vertical = TRUE,method = "jitter", add = TRUE, pch = 20, col = rgb(0.5,0.5,0.5,0.5))

boxplot(abs(pop_info$overall_trend)~pop_info$class, lwd = 1, ylab = ' ',col='white',cex.lab=1.4,cex.axis=1.2,las=2,xaxt="n")
stripchart(abs(pop_info$overall_trend)~pop_info$class, vertical = TRUE,method = "jitter", add = TRUE, pch = 20, col = rgb(0.5,0.5,0.5,0.5))
  mtext('| slope coefficient |',2,line=3.5,cex=1.2)
  mtext(text = '(b)',side = 3,line = -1.2,adj=0.95,font = 2,cex=1.2)

boxplot(pop_info$variance~pop_info$class, lwd = 1, ylab = 'population variance',col='white',cex.lab=1.4,cex.axis=1.2,las=2)
stripchart(pop_info$variance~pop_info$class, vertical = TRUE,method = "jitter", add = TRUE, pch = 20, col = rgb(0.5,0.5,0.5,0.5))
mtext(text = '(c)',side = 3,line = -1.2,adj=0.95,font = 2,cex=1.2)

boxplot(pop_info$autocorrelation~pop_info$class, lwd = 1, ylab = 'autocorrelation',col='white',cex.lab=1.4,cex.axis=1.2,las=2,ylim=c(-0.2,1.2))
stripchart(pop_info$autocorrelation~pop_info$class, vertical = TRUE,method = "jitter", add = TRUE, pch = 20, col = rgb(0.5,0.5,0.5,0.5))
mtext(text = '(d)',side = 3,line = -1.2,adj=0.95,font = 2,cex=1.2)
```

\pagebreak

## Sensitivity analysis of significance level and power

Estimates of $T_{min}$ depend strongly on the values used for the significance level ($\alpha$) and the probability of type II error ($\beta$), both of which are set by the practitioner. Here we explore how estimates of $T_{min}$ are affected by changes in each of these parameters. We see that the minimum time required increases with increases in statistical power or decreases with increases in the threshold for statistical significance (Fig \ref{fig:min_time_vs_alpha_beta}).

```{r,echo=F,message=F,error=F,fig.cap='Minimum time required to assess long-term trends in abundance for values of statistical significance ($\\alpha$) and power ($1-\\beta$).\\label{fig:min_time_vs_alpha_beta}'}
# Bring in theoretical code
# Build plot of $T_min$ vs alpha and another versus beta
# Explain results after building figures
setwd("~/Desktop/Research/time-series-project")
source('scripts/calculate_slope.R') 
source('scripts/theoretical-models/simulate_pop.R')
# calculate_power = function(r,phi,sigma,trials,max.time,alpha){
#   trials   = trials
#   max.time = max.time
#   r        = r
#   phi      = phi
#   sigma    = sigma
#   
#   y=replicate(n=trials,simulate_pop(r,phi,sigma,max.time)) 
#   
#   power=sum((apply(y,2,calculate_p_value)<alpha) & sign(apply(y,2,calculate_slope))==sign(r))/trials 
#   
#   return(power)
# }
# 
# 
# min_time_needed= function(alpha,power){
#   years_to_sample=seq(2,30,by=1)
#   y=lapply(X=years_to_sample,FUN = calculate_power,phi =0.5,r=0.5,sigma = 1,trials=100,alpha=alpha)
#   min_time_needed = years_to_sample[tail(which(y<power),1)] + 1
#   return(min_time_needed)
# }
# 
# alpha_values=seq(0.001,0.2,by=0.01)
# power_values=seq(0.65,0.95,by=0.01)
# parameter_combinations = expand.grid(alpha_values,power_values)
# names(parameter_combinations) = c('alpha','power')
# parameter_combinations$min_time_required = NA
# 
# for (i in 1:nrow(parameter_combinations)){
# 
#   parameter_combinations$min_time_required[i] =  min_time_needed(parameter_combinations[i,1],parameter_combinations[i,2])
#  
#   #print(i)
# }

#write.csv(parameter_combinations,file = 'analysis-outputs/min_time_vs_alpha_and_power.csv',quote = F,row.names = F)

# Pull in data to prevent R from running for a long time to create data from scratch
parameter_combinations = read.csv('analysis-outputs/min_time_vs_alpha_and_power.csv',header=T)

require(ggplot2)
ggplot(parameter_combinations, aes(x = alpha, y = power, z = min_time_required, fill=min_time_required)) +
  geom_tile() + 
  labs(title = " ", x = expression(alpha), y = expression(paste('power ',(1-beta))), fill = expression(T[min])) +
  scale_fill_distiller(palette="Spectral", na.value="white") +
  theme_bw()
```

# Minimum time calculations testing exponential growth

In the main text we evaluated the minimum time required to determine long-term trends in abundance via linear regression. This process examined linear trends in abundance over time. Here we examine the minimum time required to estimate long-term trends that are either exponential growth or decay. We use the same methods as described in the previous sections, but we take the $log$ of population density, or abundance.

```{r,echo=F,message=F,error=F,fig.cap='Distribution of the minimum time required in order to detect a significant trend (at the 0.05 level) in log(abundance) given power of 0.8.\\label{fig:min_time_dist_log_pop}'}
setwd("~/Desktop/Research/time-series-project")
# load('cleaned-data/cleaned_timeseries_database.Rdata')
# 
# source('scripts/calculate_slope.R')
# source('scripts/calculate_power_metric.R')
# 
# for (j in 1:nrow(pop_info)){
#     pop= subset(long_dat,long_dat$ID==pop_info$ID[j])
#     pop = pop[(nrow(pop)-34):nrow(pop),]
#     pop$popvalue=as.numeric(pop$popvalue)
#     pop$popvalue=pop$popvalue/max(pop$popvalue) #should I standaridize this in a better way
#     ## Take the log of population size and add a small number
#     pop$popvalue = log(pop$popvalue + 0.001)
#     
#     pop_info$year[j] = pop$year[1]
#     pop_info$variance[j]= var(pop$popvalue)
#     pop_info$overall_trend[j] = calculate_slope(pop$popvalue)
#     pop_info$autocorrelation[j] = acf(pop$popvalue,plot = FALSE)$acf[2]
#     pop_info$trend_p_value[j]=calculate_p_value(pop$popvalue)
#     pop_info$popvalue = pop$popvalue[1] #initial abundance
#     pop_info$series_length = nrow(pop)
#     
# 
#      pop_info$min_time_for_power[j] = min_time_needed(pop$popvalue,0.05,0.8)
#     print(paste(j,':',pop_info$min_time_for_power[j],sep=' '))
#     
# }
# 
# pop_info=pop_info[pop_info$trend_p_value<0.05,]
# long_dat=subset(long_dat,long_dat$ID %in% pop_info$ID)
# long_time_names=subset(long_time_names,long_time_names %in% pop_info$ID)
# 
# save(pop_info,long_dat,long_time_names,file = 'cleaned-data/cleaned_timeseries_database_with_min_time_linear_regression_for_log_pop.Rdata')

# LOAD in previously run script from this code chuck and make histogram plot
load('cleaned-data/cleaned_timeseries_database_with_min_time_linear_regression_for_log_pop.Rdata')

plot(density(pop_info$min_time_for_power),las=1,ylab='',xlab='',cex.axis=1.2,main='')
mtext(text = 'Frequency',2,line=3,cex=1.2)
mtext(text = 'Minimum time required',1,line=3,cex=1.2)

```

We see that the distribution of $T_{min}$ is almost identical to that in the main manuscript. This is perhaps not surprising as most time series that would significantly increase or decrease linearly would probably also significantly increase or decrease at an exponential rate. Further, the calculations here and in the main manuscript both use linear regression, and both estimate the same number of parameters.

\pagebreak

# Simulations with more complicated population model

In the main text, we showed how a simple population model could be simulated repeatedly to estimate the power obtained with time series of increasing length. The model in the main text simulated linear population growth with only a slope coefficient, y-intercept, and noise parameter required. This model is purely phenomenological and does not include any species life history. Here, we use the same routine as the main text, but simulate from a more biologically-realistic population model. We simulate the model described in @White2014. The model is a stochastic, age-structured population model that includes density-dependence for lemon sharks in Bimini, Bahamas.

```{r, echo=F,fig.cap='Statistical power for different length of time series simulations for a lemon shark population in Bimini, Bahamas.\\label{fig:shark_example}'}

setwd("~/Desktop/Research/time-series-project")
#source('scripts/LemonShark_PopulationModel.R')
years_to_sample=seq(5,50,by=1)
#shark_power=sapply(years_to_sample,power)

load('analysis-outputs/example_power_for_shark_model.Rdata')
par(mfrow=c(1,1),oma=c(3,3,0.5,0.5),mar=c(2,2,0,0))
   plot(years_to_sample,shark_power,type='l',lwd=2,las=1,ylab='',xlab='',cex.axis=1.2)
      abline(v=years_to_sample[tail(which(shark_power<0.8),1)+1],lty=2,lwd=2,col='red')
      abline(h=0.8,lty=2,lwd=2,col='red')
      mtext(text = 'time sampled' ,side = 1,cex = 1.2,line=3)
      mtext(text = 'statistical power' ,side = 2,cex = 1.2,line=3.5)
      #mtext(text = '(b)',side = 3,line = -1.2,adj=0.05,font = 2,cex=1.2)

#save(shark_power,file='analysis-outputs/example_power_for_shark_model.Rdata')

```

Here, we parameterize the model for a situation where adult mortality rate is high enough to cause a population decline. In the same way we simulated models in the main text, we simulate this more biologically-realistic model for different lengths of time. For each length of time, we calculate the statistical power. Similar to results with the simpler model, statistical power generally increases with longer sampling time (Fig. \ref{fig:shark_example}). In this example, the minimum time required ($T_{min}$) to obtain at least 0.8 statistical power, given a significance level of 0.05, is `r years_to_sample[tail(which(shark_power<0.8),1)+1]` years.

<!--
## Optional section looking at Gompertz model
In the main text, we showed how a simple population model could be simulated repeatedly to estimate the power obtained with time series of increasing length. The model in the main text simulated linear population growth with only a slope coefficient, y-intercept, and noise parameter required. This model is purely phenomenological and does not include any species life history. Here, we use the same routine as the main text, but simulate from a more biologically-realistic population model, the stochastic Gompertz population model:

  \begin{equation}
    N_{t+1} = N_t e^{a - b \mbox{ log} N_t + \epsilon_t}
  \end{equation}

Here, $N_t$ is the population density in year $t$, $a$ is an intercept term, and $b$ is a parameter measuring the strength of density-dependence. Lastly, $\epsilon_t$ is process error that is assumed to be normally distributed. @Knape2012 used this model to investigate the frequency of density-dependence in time series of populations.

We simulate this model repetitively for different lengths of time series. Statistical power is the fraction of time series of a set length that showed significant trends in abundance, given a true trend in abundance. *How do I determine significance here? Through linear regression again, or do I look at non-zero estimation of the parameter values?*-->

# Minimum time required to estimate geometric growth rate

Instead of detecting a trend over time with linear regression, we could also calculate the geometric growth rate of the population. In figure \ref{fig:growth_rate}, we show how to calculate growth rates for subsamples of a time series. First, we created subsamples of each possible length from the full 35 year time series, as we did in the main next. Next, we calculated the mean and standard deviation of growth rates for each possible time series length (Fig. \ref{fig:growth_rate}b). Lastly, we calculated the percent error $\mbox{percent error} = 100 \times \left| \frac{\mbox{observed} - \mbox{theoretical}}{\mbox{theoretical}} \right|$ between the mean of each time series length (observed) and the overall population growth rate (theoretical). In Fig. \ref{fig:growth_rate}c), we show the percent error as a function of time series length. Here we define the minimum time required as the minimum number of years to achieve less than 20% error.

```{r, echo=F, fig.cap='Example of calculating minimum time required for growth rate estimation. (a) European herring gull (\\emph{Larus argentatus}) scaled density over time, (b) mean and standard deviation of growth rate for subsamples of entire time series, and (c) the percent error between mean estimated growth rate and the true long-term growth rate. The vertical bar denotes the minimum time required to estimate growth rate within 20% error.\\label{fig:growth_rate}'}
#Example of using population growth rate instead of power
setwd("/Users/eastonwhite/Desktop/Research/time-series-project")
source("scripts/calculate_geometric_growth.R")
load('cleaned-data/cleaned_timeseries_database.Rdata')

par(mfrow=c(1,3),oma=c(0,1,0,0))

  pop = subset(long_dat,long_dat$ID=='10008') #13192 for Harp seals, 10007 is California Gull, 10008 for European herring gull (Larus argentatus)
  source('scripts/create_subsamples.R')
  source('scripts/percent_error.R')
  aaa=calculate_average_geometric_growth(pop$popvalue)
  plot(pop$year,pop$popvalue,cex.axis=1.2,cex.lab=1.2,ylab='',xlab='',pch=16,las=1)
    #abline(lm(pop$popvalue~pop$year),lwd=2,col='red')
    mtext(text = 'year',1,line=3,cex=1.4)
    mtext(text = 'population size',2,line=3.5,cex=1.4)
    mtext(text = '(a)',side = 3,line = -1.2,adj=0.95,font = 2,cex=1.2)

  plot(rowMeans(aaa,na.rm=T),cex.axis=1.2,pch=16,ylab='',xlab='',las=1)
    mtext(text = 'years sampled',1,line=3,cex=1.4)
    mtext(text = 'population growth rate',2,line=3.7,cex=1.4)

    abline(h= -0.01*rowMeans(aaa,na.rm = T)[34] +rowMeans(aaa,na.rm = T)[34] )
    abline(h= 0.01*rowMeans(aaa,na.rm = T)[34] +rowMeans(aaa,na.rm = T)[34] )

    arrows(1:34,rowMeans(aaa,na.rm=T),1:34,rowMeans(aaa,na.rm=T)+sd(apply(aaa,MARGIN = 1,FUN=var,na.rm=T),na.rm=T),angle = 90,length = 0.1)
    arrows(1:34,rowMeans(aaa,na.rm=T),1:34,rowMeans(aaa,na.rm=T)-sd(apply(aaa,MARGIN = 1,FUN=var,na.rm=T),na.rm=T),angle = 90,length = 0.1)
    mtext(text = '(b)',side = 3,line = -1.2,adj=0.95,font = 2,cex=1.2)

    plot(percent_error(rowMeans(aaa,na.rm=T),rowMeans(aaa,na.rm=T)[34]),cex.axis=1.2,pch=16,ylab='',xlab='',las=1)
    mtext(text = 'years sampled',1,line=3,cex=1.4)
    mtext(text = 'percent error in growth rate',2,line=3.7,cex=1.4)
    abline(h=0.2,lty=2)
    abline(v=tail(which(percent_error(rowMeans(aaa,na.rm=T),rowMeans(aaa,na.rm=T)[34])>0.1),1),lty=2)
    mtext(text = '(c)',side = 3,line = -1.2,adj=0.95,font = 2,cex=1.2)
```

We applied the same calculations to `r nrow(pop_info)` populations, as we did in the main text. We then obtain a distribution of the minimum time required to measure the ``true" long-term growth rate (Fig. \ref{fig:min_time_growth_dist}). We see a bimodal distribution with many populations required 30+ years to estimate the long-term growth rate. The large number of short years required is due to cases where the entire time series is consistently increasing or decreasing at the same rate.

```{r, echo=F,eval=T,fig.cap='Histogram of the minimum time required in order to estimate the long-term growth rate within 20% error.\\label{fig:min_time_growth_dist}'}
# Growth rates calculations for each species
setwd("/Users/eastonwhite/Desktop/Research/time-series-project")
# load('cleaned-data/cleaned_timeseries_database.Rdata')
# 
# source('scripts/create_subsamples.R')
# source('scripts/calculate_geometric_growth.R')
# source('scripts/percent_error.R')
#  for (j in 1:nrow(pop_info)){
#     pop= subset(long_dat,long_dat$ID==pop_info$ID[j])
#     pop$popvalue=as.numeric(pop$popvalue)
#     pop$popvalue=pop$popvalue/max(pop$popvalue) #should I standaridize this in a better way
#     
#    # source('scripts/calculate_power_metric.R')
#     aaa=calculate_average_geometric_growth(pop$popvalue)
#     if (length(which(percent_error(rowMeans(aaa,na.rm=T),rowMeans(aaa,na.rm=T)[34])>0.2))==0){
#       pop_info$min_time_growth_rate[j] = 2
#     }else{
#       pop_info$min_time_growth_rate[j] = tail(which(percent_error(rowMeans(aaa,na.rm=T),rowMeans(aaa,na.rm=T)[34])>0.2),1)+1}
#     #print(paste(j,':',pop_info$min_time_for_power[j],sep=' '))
#  }
# save(pop_info,file = 'analysis-outputs/min_time_for_growth_rate0_2.Rdata')

load('analysis-outputs/min_time_for_growth_rate0_2.Rdata')

# Create plot
par(mfrow=c(1,1))
hist(pop_info$min_time_growth_rate,las=1,ylab='',xlab='',cex.axis=1.2,main='',breaks=20,xlim=c(0,40),ylim=c(0,500))
mtext(text = 'frequency',2,line=3,cex=1.2)
mtext(text = 'minimum time required',1,line=3,cex=1.2)

```

\clearpage

# Using Generalized additive model to identify significant trends

In the main text, we examined the minimum time required to identify a trend in abundance via linear regression. This approach allowed us to identify increases or decreases, but a linear model may not always be a good fit. Generalized additive models (GAMs) are more general than general linear models and allow more flexible. GAMs are models where a response variable depends on unknown smooth functions of explanatory variables. GAMs, therefore, can identify relationships between response variables and explanatory variables that are non-linear and perhaps more complicated. The downside of GAMs is they typically require more data and are also prone to overfitting (citations).

Here, we conduct the same analyses in the main text, but instead calculate the minimum time series required to detect trends over time according to a GAM model. We hypothesize that GAMs should require less time to detect a trend as they are more flexible than linear regression. We provide an example that shows statistical power increases with more time sampled (Fig. \ref{fig:gam_example}).

```{r,echo=F,message=F,warning=F,fig.cap='(a) Time series for Bigeye tuna (\\emph{Thunnus obesus}) with corresponding fitted GAM model in red and (b) statistical power as a function of the number of years sampled. The horizontal ine at 0.8 indicates the minimum threshold for statistical power and the vertical line denotes the minimum time required to achieve 0.8 statistical power.\\label{fig:gam_example}'}
# show gam example side-by-side plot
 setwd("~/Desktop/Research/time-series-project")
 load("cleaned-data/cleaned_timeseries_database_with_min_time_linear_regression.Rdata")
 source('scripts/calculate_power_metric_gam.R')
  par(mfrow=c(1,2),oma=c(4,4,0.5,0.5))
    pop = subset(long_dat,long_dat$ID=='7558') #13192 for Harp seals, 7558 for Thunnus obesus
    plot(pop$year,pop$popvalue,cex.axis=1.2,cex.lab=1.2,ylab='',xlab='time (years)',pch=16,las=1)
    example_gam_model=gam(pop$popvalue ~ s(pop$year,k=3),family=gaussian)
    points(pop$year,example_gam_model$fitted.values,col='red',type='l',lwd=2)
    mtext(text = '(a)',side = 3,line = -1.2,adj=0.95,font = 2,cex=1.2)
    mtext(text='population size', side=2,line=4,cex=1.2)
# could insert all possible regressions of length n

    power_example=calculate_power_for_subsamples(pop$popvalue,0.05)
    plot(power_example,cex.axis=1.2,cex.lab=1.2,ylab='statistical power',xlab='years sampled',las=1,pch=16)
    mtext(text = '(b)',side = 3,line = -1.2,adj=0.05,font = 2,cex=1.2)
# could also insert standard error bars
    abline(h=0.8,col='red',lwd=2,lty=2)
    abline(v=head(which(power_example>0.8),1),col='red',lwd=2,lty=2)
```

```{r,echo=F,fig.cap='Distribution of the minimum time required in order to detect a significant trend (at the 0.05 level) in abundance according to a GAM model given statistical power of 0.8. The smoothing parameter was set to 3 for each population.\\label{fig:min_time_dist_gam}'}
# show gam results for minimium time required
setwd("/Users/eastonwhite/Desktop/Research/time-series-project")
load("cleaned-data/cleaned_timeseries_database_with_min_time_linear_regression.Rdata")
main_text = pop_info
# load('cleaned-data/cleaned_timeseries_database.Rdata')
#  source('scripts/calculate_power_metric_gam.R')
#  for (j in 1:nrow(pop_info)){
#     pop= subset(long_dat,long_dat$ID==pop_info$ID[j])
#     pop$popvalue=as.numeric(pop$popvalue)
#     pop$popvalue=pop$popvalue/max(pop$popvalue) #should I standaridize this in a better way
#   
#     pop$trend_p_value = calculate_p_value_gam(pop$popvalue)
#  
#     pop_info$min_time_for_power_gam[j] = min_time_needed(pop$popvalue,0.05,0.8)
#     print(paste(j,':',pop_info$min_time_for_power[j],sep=' '))
#  }
# 
# #pop_info$min_time_for_power_gam[pop_info$min_time_for_power_gam==-999]=33
# 
# pop_info=pop_info[pop_info$trend_p_value<0.05,]
# long_dat=subset(long_dat,long_dat$ID %in% pop_info$ID)
# long_time_names=subset(long_time_names,long_time_names %in% pop_info$ID)
# 
# save(pop_info,long_dat,long_time_names,file='analysis-outputs/min_time_required_GAM.Rdata')

load('analysis-outputs/min_time_required_GAM.Rdata')
# Create plot
par(mfrow=c(1,1))
min_time_gam = pop_info$min_time_for_power_gam[pop_info$min_time_for_power_gam!=-999]
plot(density(min_time_gam),las=1,ylab='',xlab='',cex.axis=1.2,main='',xlim=c(0,40))
mtext(text = 'Frequency',2,line=3,cex=1.2)
mtext(text = 'Minimum time required',1,line=3,cex=1.2)
```

We then fit GAM models for `r length(pop_info$min_time_for_power_gam[pop_info$min_time_for_power_gam!=-999])` populations. We found a similar distribution of minimum time required as in the main text for linear regression (Fig. \ref{fig:min_time_dist_gam}). However, in line with our hypothesis, the GAM models did result in a lower mean minimum time required of `r round(mean(pop_info$min_time_for_power_gam[pop_info$min_time_for_power_gam!=-999]),2)` years compared to the results from the main text of `r round(mean(main_text$min_time_for_power),2)` years.

<!--
## Minimum time to detect density-dependence in time series

For decades, the importance of density-dependence in natural populations has been an important question in ecology (citations). To address this question, @Knape2012 used data from the Global Population Dynamics Database [@GPDD2010]. They fitted 627 time series with a Gompertz population model and found evidence for density-dependence in 45% of the time series. Each time series in their database contained at least 15 data points, and ranged from 15 to 157. A follow up question to their work would be, what time series length is required to confidently estimate density-dependence. 
-->

\clearpage

# References