### 2017-07-07 writing/manuscript_draft.Rmd
added citations and modified code for figure 1

---
bibliography: White_bib.bib
fontsize: 12pt
geometry: margin=1in
header-includes: \usepackage{float} \usepackage{setspace}\doublespacing \usepackage{lineno}
  \usepackage[round]{natbib} \bibpunct[; ]{(}{)}{,}{a}{}{;} \linenumbers

output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
  word_document: default
---

Running head: 
\vspace{3 mm}

Title: Your time series is (probably) too short
\vspace{7 mm}

Author(s): \textsc{Easton R. White$^{1,2}$} 
\vspace{3 mm}

Addresses:\emph{$^1$ Center for Population Biology, University of California-Davis, Davis, California 95616 USA}
\vspace{3 mm}

$^2$Corresponding author: eawhite@ucdavis.edu
\vspace{3 mm}

Number of words: X,XXX (less than 4,000 words or 20 pages)
\vspace{3 mm}

To be submitted to: \emph{Biological Conservation} (Short Communication)
\vspace{3 mm}

Available as a preprint at: www.peerj.com
\vspace{3 mm}

Keywords: ecological time series, experimental design, monitoring, power analysis, statistical power, sampling design

\vspace{3 mm}

\pagebreak

\linenumbers

# Abstract
Long-term census data is necessary to better understand ecological processes and make management decisions. However, census data is often expensive, requiring a lot of people-hours and equipment. Little work has addressed the question, when is a time series of census data long enough to address a question of interest? Here, we explore two approaches to address this question: through simulations and through an empirical approach. We specifically determine the minimum time series length required to estimate significant increases or decreases in population abundance. Importantly, We examine the ability to detect a trend with a set level of statistical power, which is often neglected in ecological time series analyses. Using a simulation approach, we demonstrate how the minimum time series length required increases with weaker trends in abundance and with higher population variability. In addition, we examine 868 populations of vertebrate species to determine the minimum time required to detect changes in their abundance. We show that 10-15 years of continuous monitoring are required in order to achieve a high level of statistical power. Similar to simulation-approaches, the minimum time required for field census data strongly depends on trend strength, population variability, and temporal autocorrelation. <!--Perhaps surprisingly, the minimum time required did not correlate well with biological explanatory variables, like body size or generation length.--> These results point to the importance of sampling populations over long periods of time. <!--Here, We stress the need for verifying a study has enough statistical power to accurately determine long-term changes in population abundance.--> Most studies, especially those less than 10-15 years, are probably underpowered and potentially misleading.

# Introduction

Observational studies and population censuses have become a cornerstone of modern ecological research and conservation biology [@Magurran2010; @Hughes2017]. Long-term data is necessary to both understand population dynamics and to assess extinction risk of species. Even though many time series may now be considered "long-term", most are still short. Time series are typically short for a variety of reasons. They are often coupled with an experiment, which may only last a couple of years. In addition, short funding cycles make it difficult to examine populations over longer periods of time [@Hughes2017]]. There are, of course, notable exceptions where long-term monitoring has occured, for example the continuous plankton recorder [@Giron-Nava2017].

Naturally, one might ask, how long of a time series is actually necessary? This question has important implications for both research and management. Scientifically, we need to know the time series length required to address a specific question. Too short of a time series may lead to wrong conclusions given large natural year-to-year variability (citations). From a management perspective, it is important to understand when a trend in abundance over time is actually meaningful or not. The IUCN Red List Categories and Criteria suggest under Criterian A2, a species qualifies as vulnerable if it has experienced a 30% decline over 10 years, or 3 generations [@IUCN2012]. For both scientific and management questions, because sampling is typically expensive, we do not want to sample for longer than is neccessary [@Gerber1999].

In determining the number of samples required for a particular experiment four quantities are intricately linked: statistical power (1-$\beta$), effect size, $\alpha$, and sample size [@Legg2006]. The exact relationship between these quantities depends on the specific statistical test. Formally, statistical power ($1-\beta$) is one minus the probability of a type II error ($\beta$), or false negative. In other words, it is a measure of how well you could detect a trend given it is actually significant. Prior to an experiment, one could set appropriate levels of power, significance threshold ($\alpha$), and the effect size to estimate the sample size required for the experiment. This approach, however, is not straight-forward for a time series, or more complicated scenarios [@Johnson2015], as data points are clearly non-independent.

For time series data, two general approaches can be used. First, as can be done with experimental data, we can relax the definition of statistical power [@Bolker2008]. Instead of using classic statistical tests (e.g. t-test), and corresponding formulas for determinig sample size, simulation approaches can be used [@Gerrodette1987, Bolker2008]. Simple models can be simulated with parameter values corresponding to a population of interest. Out of a large number of simulations, power can be calculated as the fraction of simulations that meet some criteria. The specific criteria depend on the question at hand. Then, time series simulations of different lengths are run to determine the minimum time series length required to be confident that you will be able to answer your question [@Bolker2008]. In addition to using simulated data, empirical time series can also be used. We usually do not have multiple replicates of similar populations, but we can subsample an empirical time series [@Gerber1999]. Subsamples of different lengths can then be evaluated to see which fraction of subsamples meet some criteria, again a measure of statistical power. Similar to the simulation approach, we can use these measures of power to determine the minimum time series required for a particular question of interest.

Past work has investigated questions related to the minimum time series required to estimate trends in population size over time [@Giron-Nava2017]. For example, @Rhodes2011 examined the optimal allocation of effort between spatial and temporal replicates. Using simple populations models, they found that the allocation of effort depends on environmental variation, spatial and temporal autocorrelation, and observer error. @Rueda-Cediel2015 also used a modeling approach, but parameterized amodel specifically for a threatened snail, \emph{Tasmaphena lamproides}. They found that for this short-lived organim, 15 years was adequate to assess long-term trends in abundance. However, these studies, and other past work, have typically focused either on theoretical aspects of monitoring design or focused on single species.

In this paper, we use both simulations and a database of empirical time series to determine the minimum number of years required to address various questions. More specifically, we estimate the minimum time series length required to asssess long-term changes in abundance via simple linear regression. We examine 878 population time series to estimate this minimum time series length required.
<!-- In this paper, we evaluate the minimal time series length required to answer X questions. 1) How sensitivive are IUCN asssessments to time series lengths? 2) What minimum time series length is required to measure the entire time series trend? 3) How long does a time series need to be to measure the variance in population size (rare event questions)? We demonstrate our approach with X case studies and then show results for X species from the Blah Blah Database.  -->

# Methods

## Simulation approach

One approach to determining the minimum time series length needed is through repetitive simulations of a population model [@Gerrodette1987]. This is the same approach one might use in sample size calculations for any experimental design too complicated for simple power analyses [@Johnson2015] (and cite Bolker). We only briefly discuss this approach as it has been desribed elsewhere (citations). Essentially, we take a stochastic population model and simulate it for a number of years repetitively. We can then determine the minimum time series required to achieve designed levels of significance and power for a particular question. This approach requires us to determine values for our model parameters (e.g. growth rate). As an example, we can take the following population model for population size $N$ at time $t$:
$$
N(t + 1) = N(t) + r(t) \mbox{ with } r(t) \sim N(\mu, \sigma)
$$
where $\epsilon$ is a normally-distributed random noise term with mean $\mu$ and standard deviation $\sigma$. The rate of growth $r$ is also the trend strength of increase or decrease.

By simulating this model repetitively for different lengths of time, we can determine the minimum time series required. Here statistical power is simply the fraction of simulations with significant trends in abundance. A significant trend in abundance is determined by estimating the significance of the slope parameter from simple linear regression. Then, the minimum time required ($T_{min}$) is the number of years required to obtain statistical power above some pre-determined threshold (historically at 0.8). Statistical power of 0.8 would indicate that, if there was a true trend in abundance, we would have a 0.8 probability of detecting the trend.

In Fig. \ref{fig:theoretical_approach}a, a number of simulated time series are shown for a set number of time periods ($t=40$). It is clear that statistical power increases quickly with increases in length of time sampled (Fig. \ref{fig:theoretical_approach}b). Once power is greater than 0.8 (the dotted line), that is the minimum time required ($T_{min}$).

```{r,echo=F,eval=T,fig.cap='(a) example of simulated time series for 40 time periods and (b) statistical power versus the simulated time series length. The horizontal, dashed line is our desired statistical power of 0.8. The vertical, dashed line is the minimum time required to achieve our desired statistical power. (c) Minimum time required for different values of the trend strength ($r$) and (d) minimum time required for different levels of population variability ($\\sigma$).\\label{fig:theoretical_approach}'}
source("../scripts/theoretical-models/calculate_power.R")
source("../scripts/theoretical-models/simulate_pop.R")

  # years_to_sample=seq(2,100,by=1)
  # power=lapply(X=years_to_sample,FUN = calculate_power,phi = 0.6,r=0.2,sigma = .5,trials=100)#should use 300+ trials
  # 
  # simulations = replicate(n=30,simulate_pop(phi = 0.6,r=0.2,sigma = .5,max.time=100))
  # min_time_needed = years_to_sample[tail(which(power<0.8),1)] + 1
  # 
  # save(simulations,years_to_sample,power,min_time_needed,file='../analysis-outputs/example_simulations_fig1.Rdata')
  load('../analysis-outputs/example_simulations_fig1.Rdata')

  #load command
  # Plotting commands
  par(mfrow=c(2,2),oma=c(1,2,0,0.5),mar=c(4,5,0.5,0.5))
    matplot(simulations,type='l',col=rgb(0.5,0.5,0.5,0.5),lty=1,ylab='',xlab='',las=1,cex.axis=1.2)
      mtext(text = 'time' ,side = 1,cex = 1.2,line=3)
      mtext(text = 'population size' ,side = 2,cex = 1.2,line=4)
      mtext(text = '(a)',side = 3,line = -1.2,adj=0.05,font = 2,cex=1.2)

    plot(years_to_sample,power,type='l',lwd=2,las=1,ylab='',xlab='',cex.axis=1.2)
      abline(v=min_time_needed,lty=2,lwd=2,col='red')
      abline(h=0.8,lty=2,lwd=2,col='red')
      mtext(text = 'time sampled' ,side = 1,cex = 1.2,line=3)
      mtext(text = 'statistical power' ,side = 2,cex = 1.2,line=3.5)
      mtext(text = '(b)',side = 3,line = -1.2,adj=0.05,font = 2,cex=1.2)

    #   
    # # Plot versus trend strength
    #   r_values = seq(1,3,by=0.1) #by=0.2
    #   sigma_values=seq(1,7,by=0.3) #by=0.2
    # 
    #   parameter_combinations = expand.grid(r_values,5)
    #   parameter_combinations$min_time_required = NA
    #   names(parameter_combinations)= c('r','sigma','min_time_required')
    #   
    #   for (i in 1:nrow(parameter_combinations)){
    #     years_to_sample=seq(2,50,by=1)
    #     y=lapply(X=years_to_sample,FUN = calculate_power,phi =0.5,r=parameter_combinations$r[i],sigma = parameter_combinations$sigma[i],trials=150)
    #     if (length(tail(which(y<0.8),1))>0){
    #     parameter_combinations$min_time_required[i] = years_to_sample[tail(which(y<0.8),1)] + 1
    #     }
    #     #print(i)
    #   }
    #   
    #   save(parameter_combinations,file='../analysis-outputs/min_time_vs_trend_strength.Rdata')
      load('../analysis-outputs/min_time_vs_trend_strength.Rdata')

      plot(parameter_combinations$r,parameter_combinations$min_time_required,type='l',lwd=2,las=1,ylab='',xlab='',cex.axis=1.2,ylim=c(0,50))
      mtext(text = 'trend strength (slope coefficient)' ,side = 1,cex = 1.2,line=3)
      mtext(text = 'minimum time required' ,side = 2,cex = 1.2,line=3.5)
      mtext(text = '(c)',side = 3,line = -1.2,adj=0.05,font = 2,cex=1.2)

      # 
      # 
      # 
      # # Plot versus sigma values
      # parameter_combinations = expand.grid(1.5,sigma_values)
      # parameter_combinations$min_time_required = NA
      # names(parameter_combinations)= c('r','sigma','min_time_required')
      # 
      # for (i in 1:nrow(parameter_combinations)){
      #   years_to_sample=seq(2,50,by=1)
      #   y=lapply(X=years_to_sample,FUN = calculate_power,phi =0.5,r=parameter_combinations$r[i],sigma = parameter_combinations$sigma[i],trials=150)
      #   if (length(tail(which(y<0.8),1))>0){
      #   parameter_combinations$min_time_required[i] = years_to_sample[tail(which(y<0.8),1)] + 1
      #   }
      #   #print(i)
      # }

      #save(parameter_combinations,file='../analysis-outputs/min_time_vs_pop_var.Rdata')
      load('../analysis-outputs/min_time_vs_pop_var.Rdata')
      plot(parameter_combinations$sigma,parameter_combinations$min_time_required,type='l',lwd=2,las=1,ylab='',xlab='',cex.axis=1.2,ylim=c(0,50))
      mtext(text = 'population variability' ,side = 1,cex = 1.2,line=3)
      mtext(text = 'minimum time required' ,side = 2,cex = 1.2,line=3.5)
      mtext(text = '(d)',side = 3,line = -1.2,adj=0.05,font = 2,cex=1.2)
```

<!-- code to pull in databases-->
```{r, echo=F}
# Pull in original database from Keith et al. 2015 and the cleaned version after our series of filtering steps

#Load keith et al. database (from GPDD database)
  load('../data/Keith_et_al_2015_data.Rdata')

#Load our cleaned data (see other code for cleaning steps)
  load('../cleaned-data/cleaned_timeseries_database_with_min_time_linear_regression.Rdata') #Do i need to filter out non-significant populations as well (where no trend was present)
 # pop_info=pop_info[pop_info$trend_p_value<0.05,]
 # long_dat=subset(long_dat,long_dat$ID %in% pop_info$ID)
 # long_time_names=subset(long_time_names,long_time_names %in% pop_info$ID)
```

## Data source
We use a database of `r length(table(dat$ID))` population time series compiled in [@Keith2015]. The data is originally from the Global Population Dynamics Database [@GPDD2010]. We filtered out short time series (less than `r pop_info$series_length[1]` years), and those with missing data, and were left with `r nrow(pop_info)` time series. The database includes information on `r length(table(pop_info$Binomial))` vertebrate species with a focus on mammals, birds, and fish. The database also includes information on generation length and census specifications. For each time series, we also calculated variables of interest like variance in population size, long-term trend in abundance (slope coefficient from simple linear regression), and temporal autocorrelation. Lastly, we also calculated the minimum number of years required (10 years or 3 generations) under IUCN Red List Criterian A2 to classify a species as vulnerable [@IUCN2012].

```{r,echo=F}
# Pull in results of biological characteristics comparison to min time series length required
load('../cleaned-data/combined_databases_with_min_time_linear_regression.Rdata')
```
For a subset of populations in the database ($n =$ `r nrow(LPI_pop_info)`), we were able to connect a database on biological characteristics [@Myhrvold2015] like body size and generation length. All `r nrow(LPI_pop_info)` populations are within the Aves class. In the supplementary material, we explore how the minimum time required does not strongly correlate with any biological characteristics.

## Empirical approach

<!--Our objective is to determine the minimum time series length ($T_{min}$) required to answer a specific question of interest. As we describe below, the minimum time series length calculation is particular to each question. Intuitively, we might expect a longer minimum time series length ($T_{min}$) required given more complicated questions (citations).-->

We assume that each time series in our database is long enough to include all necessary information (e.g. variability) about the population. In other words, each time series is a representative sample. We first take all possible contiguous subsamples of each time series. For example, a time series of 40 years would have 39 possible contiguous subsamples of length 2, 38 possible contiguous subsamples of length 3, and continuing until 1 possible contiguous subsample of length 40 [@Gerber1999; @Giron-Nava2017]. Then we determined the $T_{min}$ required to be confident (i.e. have high enough statistical power) in an estimate of the slope coefficient from linear regression. Therefore, we run a linear regression for each subsampled time series. Then, we determine the fraction of subsamples of a particular length that have estimated slope coefficients which are statistically different from zero given the long-term, or ``true", time series also has a significant slope. This is a measure of statistical power (often denoted as $1 - \beta$, where $\beta$ is the probability of a type II error). Lastly, we determine which subsample length is required to guarantee we have above a certain threshold of statistical power, which we will assume is 0.8 by convention (citation).

<!--
## Questions addressed in the appendix

We also examine how minimum time series calculations changed depending on the question of interest. In the supplementary material, we use similar calculations to determine the minimum time series required to estimate overall population growth rate (Fig. \ref{fig:growth_rate}). We also explore how the method can be applied to more biologically-realistic population models (Fig. \ref{fig:shark_example}. More specifically, we determine the minimum time required to estimate long-term population trends using a stochastic, age-structured model of lemon shark population dynamics in the Bahamas [@White2014]. Lastly, we determine the minimum time required to estimate long-term trends according to generalized additive models, instead of the simple linear models used here.
-->

# Results

We determined the minimum time series length required to address a particular question of interest. More specifically we asked, what is the minimum time series length required to determine, via linear regression, the long-term trend in population abundance? The minimum time series length required is that which had high enough statistical power. We investigated this question through two approaches: simulation-based and empirical analyses. We briefly desribe results from the simulation approach and then discuss our emprirical approach.

## Simulation approach

We constructed a general population model where the trend over time could be a model parameter. We chose a simple model, but any other population model could be used. The specific model choice should be tailored to the population of interest. We then ran simulated time series of different lengths. From these simulations we determined the minimum time series length required to achieve a certain level of power (which we set to 0.8). We were also able to alter the parameter values (trend strength and population variability) of the population model to see how this altered $T_{min}$.

In line with past work, we found the $T_{min}$ increases (i.e. more time is required) with decreases in trend strength and with increases in variability (citation).

## Empirical test

```{r,echo=F,fig.cap='Distribution of the minimum time required in order to detect a significant trend (at the 0.05 level) in abundance given power of 0.8.\\label{fig:min_time_dist}'}
# Pull in analysis of average min time needed for **empirical species** (save analysis as Rdata file or something)
 load("../cleaned-data/cleaned_timeseries_database_with_min_time_linear_regression.Rdata")

#only look at pops with significant long-term declines or increases
#pop_info=pop_info[pop_info$trend_p_value<0.05,]
#long_dat=subset(long_dat,long_dat$ID %in% pop_info$ID)
#long_time_names=subset(long_time_names,long_time_names %in% pop_info$ID)

plot(density(pop_info$min_time_for_power),las=1,ylab='',xlab='',cex.axis=1.2,main='')
mtext(text = 'Frequency',2,line=3,cex=1.2)
mtext(text = 'Minimum time required',1,line=3,cex=1.2)
```

```{r poisson_model, echo=F,include=F}
# I should put this in the appendix
#Poisson regression
pop_info$abs_overall_trend = abs(pop_info$overall_trend)
poisson_model <-glm(min_time_for_power~ abs_overall_trend + autocorrelation + variance + gen_len,family = 'poisson',data=pop_info)
par(mfrow=c(2,2),mar=c(5,5,1,2),oma=c(0,0,0.5,0))
#plot(poisson_model,cex.lab=1.2,las=1,cex.axis=1.2,main=' ')

#R^2 calculation (pg. 218 Zuur)
R2 <- 100*(poisson_model$null.deviance - poisson_model$deviance)/(poisson_model$null.deviance)

```

```{r,echo=F, fig.cap='Minimum time required to estimate change in abundance correlated with (a) trend strength (absolute value of slope coefficient estimated from linear regression), (b) population variance (interannual variability in population size), and (c) temporal autocorrelation.\\label{fig:correlates}'}

# Plot minimum time required against trend strength, population variance, and autocorrelation  
  par(mfrow=c(3,1),mar=c(4,1,1,1),oma=c(0,4.5,0,0.5))
  plot(abs(pop_info$overall_trend),pop_info$min_time_for_power,pch=16,col=rgb(0.5,0.5,0.5,0.5),las=1,ylab='',xlab='',cex.axis=1.2)
  mtext('| trend strength |',1,line=3,cex=1.2)
  mtext('(a)',3,line=-1.5,adj = 0.985,cex=1.2,font = 2)
    plot(pop_info$variance,pop_info$min_time_for_power,pch=16,col=rgb(0.5,0.5,0.5,0.5),las=1,ylab='',xlab='',cex.axis=1.2)
    mtext('population variance',1,line=3,cex=1.3)
    mtext('(b)',3,line=-1.5,adj = 0.985,cex=1.2,font = 2)
      plot(pop_info$autocorrelation,pop_info$min_time_for_power,pch=16,col=rgb(0.5,0.5,0.5,0.5),las=1,ylab='',xlab='',cex.axis=1.2)
      mtext('autocorrelation',1,line=3,cex=1.2)
       mtext('(c)',3,line=-1.5,adj = 0.985,cex=1.2,font = 2)
       mtext('minimum time required',2,line=2.5,cex=1.2,outer = T)

```

We examined a database of `r nrow(pop_info)` separate population time series representing `r length(table(pop_info$Binomial))` species. This database consists of species with a variety of life history characteristics. We limited our analyses to populations with at least 35 years of continuous sampling.

Across all the populations we examined, we found an average minimum time series length required ($T_{min}$) of `r round(mean(pop_info$min_time_for_power),1)` ($\sigma$=`r round((var(pop_info$min_time_for_power))^0.5,1)`), with a wide distribution (Fig. \ref{fig:min_time_dist}). In addition, as expected based on theoretical results (cite papers), the minimum time series length required is strongly correlated with trend strength, population variance, and autocorrelation (Fig \ref{fig:correlates}). Using a generalized linear model, with a Poisson error structure, all three of these explanatory variables were significant and had large effect sizes (see Fig. SX). Combined, trend strength, population variance, and autocorrelation account for about `r round(100*(poisson_model$null.deviance - poisson_model$deviance)/(poisson_model$null.deviance),1)` of the explained deviance, which is similar to $R^2$ [@Zuur2009], in minimum time series length required.

We also examined a subset of populations where more explanatory variables were known (e.g. body mass, longevity, location). None of these explanatory variables were predictive of the minimum time series length required (**supp mat**). Some differences if $T_{min}$ did occur between biological class (**supp mat**), but this can be attributed to the amount populations are known to fluctuate (i.e. the variability in population size).

```{r,echo=F,include=FALSE,eval=FALSE,fig.cap='Minimum time series required versus the generation length of each species\\label{fig:generation_length}'}
par(mfrow=c(1,1))
  plot(pop_info$gen_len,pop_info$min_time_for_power,pch=16,col=rgb(0.5,0.5,0.5,0.5),las=1,ylab='',xlab='',cex.axis=1.2)
  mtext('generation time (years)',1,line=3,cex=1.2)
  mtext('minimum time required',2,line=2.5,cex=1.2,outer = F)
```

## Questions in the supplmentary material

In the supplmentary material, we explore several additional questions related to statistical power and estimating $T_{min}$. First, we showed that other metrics besides estimating the slope coefficient from linear regression could be used. More specifically, we show how the same approach desribed here can be used to determine the minimum time required to estimate a population's geometric growth rate (Figs. \ref{fig:growth_rate},\ref{fig:min_time_growth_dist}). We also explored how the simulation approach can be applied to more biologically-realistic population models (Fig. \ref{fig:shark_example}). More specifically, we determine the minimum time required to estimate long-term population trends using a stochastic, age-structured model of lemon shark population dynamics in the Bahamas [@White2014]. We found that over 27 years of continous monitoring were needed in this particular scenario (Fig. \ref{fig:shark_example}. Lastly, we determine the minimum time required to estimate long-term trends according to generalized additive models, instead of the simple linear models used here (Fig. \ref{fig:gam_example}). We found that although we obtain a similar distribution of minimum times required for GAMs, the minimum time required for GAMs is on average shorter than for linear regression (Fig. \ref{fig:min_time_dist_gam}). We expand on these results in the supplementary material.

# Discussion

We explored two different approaches to estimate the minimum time series length required to address a particular question of interest. We specifically asked, what is the minimum time series length required to determine long-term changes in abundance using linear regression? The simulation-based approach has been suggested by others, especially in situations more complicated than that suited for classic power analysis [@Gerrodette1987;@Johnson2015;@Bolker2008]. Our simulations support past work that longer time series are needed when the trend strength (i.e. rate of increase or decrease) is weak or when population variability is high [@Gerrodette1987].

Here, we focus on an empirical approach to estimate the minimum time series length requried to assess changes in abundance over time. We examined long time series and then subsampled each to determine the minimum time required to achieve a desired level of significance and power for linear regression. Statistical power is important as it provides on information as to the neccessary samples required to determine a significant trend [@Legg2006]. We found that for `r nrow(pop_info)` populations, at least 10-15 years of continuous monitoring were neccessary (Fig. \ref{fig:min_time_dist}). This timeframe is in line with past work on a short-lived snail species [@Rueda-Cediel2015] and a long-lived whale species [@Gerber1999]. In line with theoretical predictions [@Rhodes2011], we also found $T_{min}$ was strongly correlated with the trend strength, variability in population size, and temporal autocorrelation (Fig. \ref{fig:correlates}). Contrary to our prior expectations, we also found that $T_{min}$ did not correlate with any biological variables of interest (Fig. \ref{fig:biological_correlates}). We initially hypothesized that species with longer lifespans or generation times may require a longer sampling period. Our result could have been a result of at least two effects. First, the database we used may not include a diverse enough set of species with different life history traits. Second, the question we poised, whether a population is increasing or decreasing, was specifically concerned with trends in population density over time. Therefore, life-history characteristics may be more important for other questions (e.g. extinction risk) where estimates minimum time required are needed.

There has been interest in determining the minimum time series length in the bird research community [@Hatch2003; @Bart2004; @Seavy2007]. For example, @Hatch2003 used seabird monitoring data to estimate the minimum time needed to achieve appropriate levels of power. He found that the minimum time required ranged between 11 and 69 years depending on species, trend strength, and study design. Similarly, @Bart2004 recommended a standard of 80% power to detect a 50% decline over 20 years in landbird populations. They found that only about 42% of populations in the Breeding Bird Survey met their criteria. @Seavy2007 asked whether statistical power was even a useful framework for assessing long-term population trends. They used 24 years of census data on Red-tailed Tropicbirds (\emph{Phaethon rubricauda}) in Hawaii and showed that to detect a 50% decline over 10 years almost always resulting in high statistical power (above 0.8). Therefore, they cautioned against only using power analyses to design monitoring schemes and instead argued for metrics that would increase precision. For example, @Seavy2007 suggest improving randomization, reducing bias, and increases detection probability when designing and evaluating monitoring programs.

An important related question, is the optimal allocation of sampling effort in space versus time. In a theoretical investigation of this question, @Rhodes2011 found that the optimal allocation of sampling depended strongly on temporal and spatial autocorrelation. If spatial population dynamics were highly correlated, then it was better to sample more temporally, and vice versa. Our work supports this idea as populations with strong temporal autocorrelation needed less years of sampling (Fig. \ref{fig:correlates}). @Morrison2008 also studied the optimal allocation of sampling effort in space versus time, but used emprical data from a long-term census of the collared pika (\emph{Ochotona collaris}) found in the Yukon. They estimated long-term growth rates among three subpopulations over a 10-year period. They found that censuses less than 5 years may be misleading and that extrapolating from one population to another, even when nearby geographically, may be untenable.

<!--
@Rueda-Cediel2015 found that 10-15 years was adequate to characterize population trends in a short lived animal. They used a matrix model of a particular species...

@Gerber1999 investigated the minimum time series required to estimate population growth of an endagered, but recovering, species. They used a long-term census to retroactively determine the minimum time series required to assess threat status. They found that only XX years were needed, XX years before the delisting decision was actually made. This highlights the importance of estimating the minimunm time series required as an earlier decision would have saved time and money [@Gerber1999].
-->

## Limitations

Our work has some limitations in determining the minimum time series length required. First, $T_{min}$ is particular to the specific question of interest. For example, in the supplementary material, we used generalized additive models to detect population trends and came up with a similar, but quantivaely different value for $T_{min}$ (Fig. \ref{fig:min_time_dist_gam}). An additional complication is that for our empirical approach, the subsampling of the full time series allows for estimates of power, but the individual subsamples are clearly not independent of one another. Further, estimates of $T_{min}$ depend on chosen values of $\alpha$ and $\beta$ (Fig. \ref{fig:min_time_vs_alpha_beta}). In an ideal setting, we would build a specific population model parameterized for each population of interest. Then, model simulations could be used to estimate the minimum time series required to address each specific question of interest. Clearly, this is not always practical, especially if conducting analyses for a wide array of species as we do here. In addition, our statistical models suggest that $T_{min}$ does not correlate with any biological variables of interest, at least for the question of linear regression (Fig. \ref{fig:biological_correlates}). Therefore, it is not possible to use these results to predict $T_{min}$ for another population, even if the population is of a species with a similar life-history to one in our database.

## Conclusions

We use a database of `r nrow(pop_info)` populations to determine the minimum time series length required to determine population trends. This goes beyond previous work that either focused on theoretical investigations or a limited number of species. We show that to identify long-term changes in abundance, 10-15 years of continuos monitoring are required (Fig. \ref{fig:min_time_dist}). In line with theoretical predictions [@Gerrodette1987], we also show that $T_{min}$ is strongly correlated with the long-term population trend (i.e. rate of increase), variability in population size, and the temporal autocorrelation (Fig. \ref{fig:correlates}). Our work implies that for many populations, census data less than 10-15 years are probably not reliable. This stresses the importance of long-term monitoring programs. From both a scientific and management perspective estimates of $T_{min}$ are important. If a time series is too short we lack statistical power to reliably detect long-term population trends. In addition, a time series that is too long may be a poor use of already limited funds [@Gerber1999]. Future work should examine other species, with a wider range of life history characteristics. In addition, similar approaches can be used to determine the minimum time series length required to address additional questions of interest.

<!--
## Time series length recommendations from other papers

- greater than 5 years [@Morrison2008]
- more than 10-15 years [@Rueda-Cediel2015]
- general advice (variability, correlations) [@Rhodes2011]
- more than 10 years [@Gerber1999]
- specific recommendations depend on amount of power and strength of trend (Bart 2004, Hatch 2003) [@Seavy2007]
- paper on syncrony from James
-->

# Supporting Information

In the supporting material, we provide an expanded methods sections, additional figures, minimum time calculations for determining exponetial growth, simulations with a more complicated population model, minimum time calculations for determining long-term growth rates, and the use of generalized additive models to identify population trends. All code and data can be found at www.github.com/erwhite1

# Acknowledgements

ERW was partially supported by a National Science Foundation Graduate Fellowship. We would like to thank members of the Ecological Theory group at the University of California, Davis for their insight. We would also like to thank T. Dallas and E. Malcolm-White<!--X anonymous reviewers -->for their helpful comments.

# References

<div id="refs"></div>

# Figures

stuff here \ref{fig:empirical_approach_example}

\clearpage

# Supplementary material

 \setcounter{figure}{0}
 \renewcommand{\thefigure}{S\arabic{figure}}

\begin{figure}[htbp]
\centering
\caption{(a) Population size of Bigeye tuna (\emph{Thunnus obesus}) over
time. The line is the best fit line from linear regresssion. (b)
Statistical power for different subsets of the time series in panel
a.\label{fig:empirical_approach_example}}
\end{figure}

\begin{figure}[htbp]
\centering
\caption{Output of generalized linear model with a Poisson error
structure for predicting the minimum time required with explanatory
variables of the absolute value of the slope coefficient (or trend
strength), temporal autocorrelation, and variability in population
size.\label{fig:poisson_model}}
\end{figure}

\begin{figure}[htbp]
\centering
\caption{Mimimum time required versus (a) generation length (years), (b)
litter size (n), (c) adult body mass (grams), (d) maximum longevity
(years), (e) egg mass (grams), and (f) incubation (days). The lines in
each plot represent the best fit line from linear
regression.\label{fig:biological_correlates}}
\end{figure}

\begin{figure}[htbp]
\centering
\caption{(a) Minimum time required to estimate change in abundance for
species class, (b) long-term trend (estimated slope coefficient) by
species class, (c) interannual variability in population size by species
class, and (d) temporal autocorrelation by species
class.\label{fig:class}}
\end{figure}

\begin{figure}[htbp]
\centering
\caption{Minimum time required to assess long-term trends in abundance
for values of statistical significance (\(\alpha\)) and power
(\(1-\beta\)).\label{fig:min_time_vs_alpha_beta}}
\end{figure}

\begin{figure}[htbp]
\centering
\caption{Distribution of the minimum time required in order to detect a
significant trend (at the 0.05 level) in log(abundance) given power of
0.8.\label{fig:min_time_dist_log_pop}}
\end{figure}

\begin{figure}[htbp]
\centering
\caption{Statistical power for different length of time series
simulations for a lemon shark population in Bimini,
Bahamas.\label{fig:shark_example}}
\end{figure}

\begin{figure}[htbp]
\centering
\caption{Example of calculating minimum time required for growth rate
estimation. (a) European herring gull (\emph{Larus argentatus}) scaled
density over time, (b) mean and standard deviation of growth rate for
subsamples of entire time series, and (c) the percent error between mean
estimated growth rate and the true long-term growth rate. The vertical
bar denotes the minimum time required to estimate growth rate within
20\% error.\label{fig:growth_rate}}
\end{figure}

\begin{figure}[htbp]
\centering
\caption{Histogram of the minimum time required in order to estimate the
long-term growth rate within 20\%
error.\label{fig:min_time_growth_dist}}
\end{figure}

\begin{figure}[htbp]
\centering
\caption{(a) Time series for Bigeye tuna (\emph{Thunnus obesus}) with
corresponding fitted GAM model in red and (b) statistical power as a
function of the number of years sampled. The horizontal ine at 0.8
indicates the minimum threshold for statistical power and the vertical
line denotes the minimum time required to achieve 0.8 statistical
power.\label{fig:gam_example}}
\end{figure}

\begin{figure}[htbp]
\centering
\caption{Distribution of the minimum time required in order to detect a
significant trend (at the 0.05 level) in abundance according to a GAM
model given statistical power of 0.8. The smoothing parameter was set to
3 for each population.\label{fig:min_time_dist_gam}}
\end{figure}