### 2017-07-07 writing/manuscript_draft.tex

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{}
  \pretitle{\vspace{\droptitle}}
  \posttitle{}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

\usepackage{float}
\usepackage{setspace}\doublespacing \usepackage{lineno}
\usepackage[round]{natbib} \bibpunct[; ]{(}{)}{,}{a}{}{;} \linenumbers

\begin{document}

Running head: \vspace{3 mm}

Title: Statistical power and detecting population trends: the need for
long-term monitoring programs \vspace{7 mm}

Author(s): \textsc{Easton R. White$^{1,2}$} \vspace{3 mm}

Addresses:\emph{$^1$ Center for Population Biology, University of California-Davis, Davis, California 95616 USA}
\vspace{3 mm}

\(^2\)Corresponding author:
\href{mailto:eawhite@ucdavis.edu}{\nolinkurl{eawhite@ucdavis.edu}}
\vspace{3 mm}

Number of words: X,XXX (less than 4,000 words or 20 pages) \vspace{3 mm}

To be submitted to: \vspace{3 mm}

Available as a preprint at: www.peerj.com \vspace{3 mm}

Keywords: ecological time series, experimental design, monitoring, power
analysis, statistical power, sampling design

\vspace{3 mm}

\pagebreak

\linenumbers

\section{Abstract}\label{abstract}

Long-term census data are necessary to better understand ecological
processes and make management decisions. However, census data are often
expensive, requiring a lot of people-hours and equipment. Little work
has addressed the question, when is a time series of census data long
enough to address a question of interest? Here, we explore two
approaches to address this question: simulations and an empirical
approach. We specifically determine the minimum time series length
required to estimate significant increases or decreases in population
abundance. Importantly, we examine the ability to detect a trend with a
set level of statistical power, which is often neglected in ecological
time series analyses. Using a simulation approach, we demonstrate how
the minimum time series length required increases with weaker trends in
abundance and with higher population variability. In addition, we
examine 868 populations of vertebrate species to determine the minimum
time required to detect changes in their abundance. We show that 10-15
years of continuous monitoring are required in order to achieve a high
level of statistical power. Similar to simulation-approaches, the
minimum time required for field census data strongly depends on trend
strength, population variability, and temporal autocorrelation. These
results point to the importance of sampling populations over long
periods of time. Most studies, especially those less than 10-15 years,
are probably under-powered and potentially misleading.

\section{Introduction}\label{introduction}

Observational studies and population censuses have become a cornerstone
of modern ecological research and conservation biology (Magurran et al.
2010; Hughes et al. 2017). Long-term data are necessary to both
understand population dynamics and to assess extinction risk of species.
Even though many time series may now be considered ``long-term'', most
are still short. Time series are typically short for a variety of
reasons. They are often coupled with an experiment, which may only last
a couple of years. In addition, short funding cycles make it difficult
to examine populations over longer periods of time (Hughes et al. 2017).
There are, of course, notable exceptions where long-term monitoring has
occurred, for example the continuous plankton recorder (Giron-Nava et
al. 2017).

Naturally, one might ask, how long of a time series is actually
necessary? This question has important implications for both research
and management. Scientifically, we need to know the time series length
required to address a specific question. Too short of a time series may
lead to wrong conclusions given large natural year-to-year variability
(citations). From a management perspective, it is important to
understand when a trend in abundance over time is actually meaningful or
not. The IUCN Red List Categories and Criteria suggest under Criterion
A2, a species qualifies as vulnerable if it has experienced a 30\%
decline over 10 years, or 3 generations (IUCN 2012). For both scientific
and management questions, because sampling is typically expensive, we do
not want to sample for longer than is necessary (Gerber, DeMaster, and
Kareiva 1999).

In determining the number of samples required for a particular
experiment four quantities are intricately linked: statistical power
(1-\(\beta\)), effect size, \(\alpha\), and sample size (Legg and Nagy
2006). The exact relationship between these quantities depends on the
specific statistical test. Formally, statistical power (\(1-\beta\)) is
one minus the probability of a type II error (\(\beta\)), or false
negative. In other words, it is a measure of how well you could detect a
trend given it is actually significant. Prior to an experiment, one
could set appropriate levels of power, significance threshold
(\(\alpha\)), and the effect size to estimate the sample size required
for the experiment. This approach, however, is not straight-forward for
a time series, or more complicated scenarios (P. C. Johnson et al.
2015), as data points are clearly non-independent.

For time series data, two general approaches can be used. First, as can
be done with experimental data, we can relax the definition of
statistical power (Bolker 2008). Instead of using classic statistical
tests (e.g.~t-test), and corresponding formulas for determining sample
size, simulation approaches can be used (Gerrodette 1987; Bolker 2008;
P. C. Johnson et al. 2015). Simple models can be simulated with
parameter values corresponding to a population of interest. Out of a
large number of simulations, power can be calculated as the fraction of
simulations that meet some criteria. The specific criteria depend on the
question at hand. Then, time series simulations of different lengths are
run to determine the minimum time series length required to be confident
that you will be able to answer your question (Bolker 2008). In addition
to using simulated data, empirical time series can also be used. We
usually do not have multiple replicates of similar populations, but we
can subsample an empirical time series (Gerber, DeMaster, and Kareiva
1999). Subsamples of different lengths can then be evaluated to see
which fraction of subsamples meet some criteria, again a measure of
statistical power. Similar to the simulation approach, we can use these
measures of power to determine the minimum time series required for a
particular question of interest.

Past work has investigated questions related to the minimum time series
required to estimate trends in population size over time (Giron-Nava et
al. 2017). For example, Rhodes and Jonzen (2011) examined the optimal
allocation of effort between spatial and temporal replicates. Using
simple populations models, they found that the allocation of effort
depends on environmental variation, spatial and temporal
autocorrelation, and observer error. Rueda-Cediel et al. (2015) also
used a modeling approach, but parameterized a model specifically for a
threatened snail, \emph{Tasmaphena lamproides}. They found that for this
short-lived organism, 15 years was adequate to assess long-term trends
in abundance. However, these studies, and other past work, have
typically focused either on theoretical aspects of monitoring design or
focused on single species.

In this paper, we use both simulations and a database of empirical time
series to determine the minimum number of years required to address
various questions. More specifically, we estimate the minimum time
series length required to assess long-term changes in abundance via
simple linear regression. We examine 878 population time series to
estimate this minimum time series length required for detecting
population trends.

\section{Methods}\label{methods}

\subsection{Simulation approach}\label{simulation-approach}

One approach to determining the minimum time series length needed is
through repetitive simulations of a population model (Gerrodette 1987).
This is the same approach one might use in sample size calculations for
any experimental design too complicated for simple power analyses (P. C.
Johnson et al. 2015) (and cite Bolker). We only briefly discuss this
approach as it has been described elsewhere (citations). Essentially, we
take a stochastic population model and simulate it for a number of years
repetitively. We can then determine the minimum time series required to
achieve designed levels of significance and power for a particular
question. This approach requires us to determine values for our model
parameters (e.g.~growth rate). As an example, we can take the following
population model for population size \(N\) at time \(t\): \[
N(t + 1) = N(t) + r(t) \mbox{ with } r(t) \sim N(\mu, \sigma)
\] where \(\epsilon\) is a normally-distributed random noise term with
mean \(\mu\) and standard deviation \(\sigma\). The rate of growth \(r\)
is also the trend strength of increase or decrease.

By simulating this model repetitively for different lengths of time, we
can determine the minimum time series required. Here statistical power
is simply the fraction of simulations with significant trends in
abundance. A significant trend in abundance is determined by estimating
the significance of the slope parameter from simple linear regression.
Then, the minimum time required (\(T_{min}\)) is the number of years
required to obtain statistical power above some pre-determined threshold
(historically at 0.8, Cohen (1992)). Statistical power of 0.8 would
indicate that, if there was a true trend in abundance, we would have a
0.8 probability of detecting the trend.

In Fig. \ref{fig:theoretical_approach}a, a number of simulated time
series are shown for a set number of time periods (\(t=40\)). It is
clear that statistical power increases quickly with increases in length
of time sampled (Fig. \ref{fig:theoretical_approach}b). Once power is
greater than 0.8 (the dotted line), that is the minimum time required
(\(T_{min}\)).

\begin{figure}[htbp]
\centering
\includegraphics{manuscript_draft_files/figure-latex/unnamed-chunk-1-1.pdf}
\caption{(a) example of simulated time series for 40 time periods and
(b) statistical power versus the simulated time series length. The
horizontal, dashed line is our desired statistical power of 0.8. The
vertical, dashed line is the minimum time required to achieve our
desired statistical power. (c) Minimum time required (\(T_{min}\)) for
simulations with different values of the trend strength (\(r\)) and (d)
different levels of population variability (\(\sigma\)). In each case,
the minimum time required is the minimum number of years to achieve 0.8
statistical power given a significance level of
0.05.\label{fig:theoretical_approach}}
\end{figure}

\subsection{Data source}\label{data-source}

We use a database of 2444 population time series compiled in (Keith et
al. 2015). The data are originally from the Global Population Dynamics
Database (NERC Centre for Population Biology 2010). We filtered out
short time series (less than 35 years), and those with missing data, and
were left with 878 time series. The database includes information on 478
vertebrate species with a focus on mammals, birds, and fish. The
database also includes information on generation length and census
specifications. For each time series, we also calculated variables of
interest like variance in population size, long-term trend in abundance
(slope coefficient from simple linear regression), and temporal
autocorrelation. Lastly, we also calculated the minimum number of years
required (10 years or 3 generations) under IUCN Red List Criterion A2 to
classify a species as vulnerable (IUCN 2012).

For a subset of populations in the database (\(n =\) 547), we were able
to connect a database on biological characteristics (Myhrvold et al.
2015) like body size and generation length. All 547 populations are
within the Aves class. In the supplementary material, we explore how the
minimum time required does not strongly correlate with any biological
characteristics (Fig. \ref{fig:biological_correlates}).

\subsection{Empirical approach}\label{empirical-approach}

We assume that each time series in our database is long enough to
include all necessary information (e.g.~variability) about the
population. In other words, each time series is a representative sample.
We first take all possible contiguous subsamples of each time series.
For example, a time series of 40 years would have 39 possible contiguous
subsamples of length 2, 38 possible contiguous subsamples of length 3,
and continuing until 1 possible contiguous subsample of length 40
(Gerber, DeMaster, and Kareiva 1999; Giron-Nava et al. 2017). Then we
determined the \(T_{min}\) required to be confident (i.e.~have high
enough statistical power) in an estimate of the slope coefficient from
linear regression. Therefore, we run a linear regression for each
subsampled time series. Then, we determine the fraction of subsamples of
a particular length that have estimated slope coefficients which are
statistically different from zero given the long-term, or ``true'', time
series also has a significant slope. This is a measure of statistical
power (often denoted as \(1 - \beta\), where \(\beta\) is the
probability of a type II error). Lastly, we determine which subsample
length is required to guarantee we have above a certain threshold of
statistical power, which we will assume is 0.8 by convention (Cohen
1992).

\section{Results}\label{results}

We determined the minimum time series length required to address a
particular question of interest. More specifically we asked, what is the
minimum time series length required to determine, via linear regression,
the long-term trend in population abundance? The minimum time series
length required is that which had high enough statistical power. We
investigated this question through two approaches: simulation-based and
empirical analyses. We briefly describe results from the simulation
approach and then discuss our empirical approach.

\subsection{Simulation approach}\label{simulation-approach-1}

We constructed a general population model where the trend over time
could be a model parameter. We chose a simple model, but any other
population model could be used. The specific model choice should be
tailored to the population of interest. We then ran simulated time
series of different lengths. From these simulations we determined the
minimum time series length required to achieve a certain level of power
(which we set to 0.8). We were also able to alter the parameter values
(trend strength and population variability) of the population model to
see how this altered \(T_{min}\).

In line with past work, we found the \(T_{min}\) increases (i.e.~more
time is required) with decreases in trend strength and with increases in
variability (citation).

\subsection{Empirical test}\label{empirical-test}

\begin{figure}[htbp]
\centering
\includegraphics{manuscript_draft_files/figure-latex/unnamed-chunk-4-1.pdf}
\caption{Distribution of the minimum time required in order to detect a
significant trend (at the 0.05 level) in abundance given power of
0.8.\label{fig:min_time_dist}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{manuscript_draft_files/figure-latex/unnamed-chunk-5-1.pdf}
\caption{Minimum time required to estimate change in abundance
correlated with (a) trend strength (absolute value of slope coefficient
estimated from linear regression), (b) population variance (interannual
variability in population size), and (c) temporal
autocorrelation.\label{fig:correlates}}
\end{figure}

We examined a database of 878 separate population time series
representing 478 species. This database consists of species with a
variety of life history characteristics. We limited our analyses to
populations with at least 35 years of continuous sampling.

Across all the populations we examined, we found an average minimum time
series length required (\(T_{min}\)) of 16.5 (\(\sigma\)=8.4), with a
wide distribution (Fig. \ref{fig:min_time_dist}). In addition, as
expected based on theoretical results (cite papers), the minimum time
series length required is strongly correlated with trend strength,
population variance, and autocorrelation (Fig \ref{fig:correlates}).
Using a generalized linear model, with a Poisson error structure, all
three of these explanatory variables were significant and had large
effect sizes (see Fig. SX). Combined, trend strength, population
variance, and autocorrelation account for about 72.6 of the explained
deviance, which is similar to \(R^2\) (Zuur et al. 2009), in minimum
time series length required.

We also examined a subset of populations where more explanatory
variables were known (e.g.~body mass, longevity, location). None of
these explanatory variables were predictive of the minimum time series
length required (Fig. \ref{fig:biological_correlates}). Some differences
if \(T_{min}\) did occur between biological class (Fig.
\ref{fig:class}), but this can be attributed to the amount populations
are known to fluctuate (i.e.~the variability in population size).

\subsection{Questions in the supplmentary
material}\label{questions-in-the-supplmentary-material}

In the supplementary material, we explore several additional questions
related to statistical power and estimating \(T_{min}\). First, we
showed that other metrics besides estimating the slope coefficient from
linear regression could be used. More specifically, we show how the same
approach described here can be used to determine the minimum time
required to estimate a population's geometric growth rate (Figs.
\ref{fig:growth_rate},\ref{fig:min_time_growth_dist}). We also explored
how the simulation approach can be applied to more
biologically-realistic population models (Fig. \ref{fig:shark_example}).
More specifically, we determine the minimum time required to estimate
long-term population trends using a stochastic, age-structured model of
lemon shark population dynamics in the Bahamas (White, Nagy, and Gruber
2014). We found that over 27 years of continuous monitoring were needed
in this particular scenario (Fig. \ref{fig:shark_example}. Lastly, we
determine the minimum time required to estimate long-term trends
according to generalized additive models, instead of the simple linear
models used here (Fig. \ref{fig:gam_example}). We found that although we
obtain a similar distribution of minimum times required for GAMs, the
minimum time required for GAMs is on average shorter than for linear
regression (Fig. \ref{fig:min_time_dist_gam}). We expand on these
results in the supplementary material.

\section{Discussion}\label{discussion}

We explored two different approaches to estimate the minimum time series
length required to address a particular question of interest. We
specifically asked, what is the minimum time series length required to
determine long-term changes in abundance using linear regression? The
simulation-based approach has been suggested by others, especially in
situations more complicated than that suited for classic power analysis
(Gerrodette 1987; P. C. Johnson et al. 2015; Bolker 2008). Our
simulations support past work that longer time series are needed when
the trend strength (i.e.~rate of increase or decrease) is weak or when
population variability is high (Gerrodette 1987).

Here, we focus on an empirical approach to estimate the minimum time
series length required to assess changes in abundance over time. We
examined long time series and then subsampled each to determine the
minimum time required to achieve a desired level of significance and
power for linear regression. Statistical power is important as it
provides on information as to the necessary samples required to
determine a significant trend (Legg and Nagy 2006). We found that for
878 populations, at least 10-15 years of continuous monitoring were
necessary (Fig. \ref{fig:min_time_dist}). This time frame is in line
with past work on a short-lived snail species (Rueda-Cediel et al. 2015)
and a long-lived whale species (Gerber, DeMaster, and Kareiva 1999). In
line with theoretical predictions (Rhodes and Jonzen 2011), we also
found \(T_{min}\) was strongly correlated with the trend strength,
variability in population size, and temporal autocorrelation (Fig.
\ref{fig:correlates}). Contrary to our prior expectations, we also found
that \(T_{min}\) did not correlate with any biological variables of
interest (Fig. \ref{fig:biological_correlates}). We initially
hypothesized that species with longer lifespans or generation times may
require a longer sampling period. Our result could have been a result of
at least two effects. First, the database we used may not include a
diverse enough set of species with different life history traits.
Second, the question we poised, whether a population is increasing or
decreasing, was specifically concerned with trends in population density
over time. Therefore, life-history characteristics may be more important
for other questions (e.g.~extinction risk) where estimates minimum time
required are needed.

There has been interest in determining the minimum time series length in
the bird research community (Hatch 2003; Bart et al. 2004; Seavy and
Reynolds 2007). For example, Hatch (2003) used seabird monitoring data
to estimate the minimum time needed to achieve appropriate levels of
power. He found that the minimum time required ranged between 11 and 69
years depending on species, trend strength, and study design. Similarly,
Bart et al. (2004) recommended a standard of 80\% power to detect a 50\%
decline over 20 years in landbird populations. They found that only
about 42\% of populations in the Breeding Bird Survey met their
criteria. Seavy and Reynolds (2007) asked whether statistical power was
even a useful framework for assessing long-term population trends. They
used 24 years of census data on Red-tailed Tropicbirds
(\emph{Phaethon rubricauda}) in Hawaii and showed that to detect a 50\%
decline over 10 years almost always resulting in high statistical power
(above 0.8). Therefore, they cautioned against only using power analyses
to design monitoring schemes and instead argued for metrics that would
increase precision. For example, Seavy and Reynolds (2007) suggest
improving randomization, reducing bias, and increases detection
probability when designing and evaluating monitoring programs.

An important related question, is the optimal allocation of sampling
effort in space versus time. In a theoretical investigation of this
question, Rhodes and Jonzen (2011) found that the optimal allocation of
sampling depended strongly on temporal and spatial autocorrelation. If
spatial population dynamics were highly correlated, then it was better
to sample more temporally, and vice versa. Our work supports this idea
as populations with strong temporal autocorrelation needed less years of
sampling (Fig. \ref{fig:correlates}). Morrison and Hik (2008) also
studied the optimal allocation of sampling effort in space versus time,
but used emprical data from a long-term census of the collared pika
(\emph{Ochotona collaris}) found in the Yukon. They estimated long-term
growth rates among three subpopulations over a 10-year period. They
found that censuses less than 5 years may be misleading and that
extrapolating from one population to another, even when nearby
geographically, may be untenable.

\subsection{Limitations}\label{limitations}

Our work has some limitations in determining the minimum time series
length required. First, \(T_{min}\) is particular to the specific
question of interest. For example, in the supplementary material, we
used generalized additive models to detect population trends and came up
with a similar, but quantitatively different value for \(T_{min}\) (Fig.
\ref{fig:min_time_dist_gam}). An additional complication is that for our
empirical approach, the subsampling of the full time series allows for
estimates of power, but the individual subsamples are clearly not
independent of one another. Further, estimates of \(T_{min}\) depend on
chosen values of \(\alpha\) and \(\beta\) (Fig.
\ref{fig:min_time_vs_alpha_beta}). In an ideal setting, we would build a
specific population model parameterized for each population of interest.
Then, model simulations could be used to estimate the minimum time
series required to address each specific question of interest. Clearly,
this is not always practical, especially if conducting analyses for a
wide array of species as we do here. In addition, our statistical models
suggest that \(T_{min}\) does not correlate with any biological
variables of interest, at least for the question of linear regression
(Fig. \ref{fig:biological_correlates}). Therefore, it is not possible to
use these results to predict \(T_{min}\) for another population, even if
the population is of a species with a similar life-history to one in our
database.

\subsection{Conclusions}\label{conclusions}

We use a database of 878 populations to determine the minimum time
series length required to determine population trends. This goes beyond
previous work that either focused on theoretical investigations or a
limited number of species. We show that to identify long-term changes in
abundance, 10-15 years of continuous monitoring are required (Fig.
\ref{fig:min_time_dist}). In line with theoretical predictions
(Gerrodette 1987), we also show that \(T_{min}\) is strongly correlated
with the long-term population trend (i.e.~rate of increase), variability
in population size, and the temporal autocorrelation (Fig.
\ref{fig:correlates}). Our work implies that for many populations,
census data less than 10-15 years are probably not reliable. This
stresses the importance of long-term monitoring programs. From both a
scientific and management perspective estimates of \(T_{min}\) are
important. If a time series is too short we lack statistical power to
reliably detect long-term population trends. In addition, a time series
that is too long may be a poor use of already limited funds (Gerber,
DeMaster, and Kareiva 1999). Future work should examine other species,
with a wider range of life history characteristics. In addition, similar
approaches can be used to determine the minimum time series length
required to address additional questions of interest.

\section{Supporting Information}\label{supporting-information}

In the supporting material, we provide an expanded methods sections,
additional figures, minimum time calculations for determining
exponential growth, simulations with a more complicated population
model, minimum time calculations for determining long-term growth rates,
and the use of generalized additive models to identify population
trends. All code and data can be found at www.github.com/erwhite1

\section{Acknowledgements}\label{acknowledgements}

ERW was partially supported by a National Science Foundation Graduate
Fellowship. We would like to thank members of the Ecological Theory
group at the University of California, Davis for their insight. We would
also like to thank T. Dallas and E. Malcolm-Whitefor their helpful
comments.

\section{References}\label{references}

\hypertarget{refs}{}
\hypertarget{ref-Bart2004}{}
Bart, Jonathon, Kenneth P. Burnham, Erica H. Dunn, Charles M. Francis,
and C. John Ralph. 2004. ``Goals and strategies for estimating trends in
landbird abundance.'' \emph{Journal of Wildlife Management} 68 (3):
611--26.

\hypertarget{ref-Bolker2008}{}
Bolker, Benjamin M. 2008. \emph{Ecological Models and Data in R}. 1st
ed. Princeton, New Jersey: Princeton University Press.

\hypertarget{ref-Cohen1992}{}
Cohen, Jacob. 1992. ``A power primer.'' \emph{Psychological Bulletin}
112 (1): 155--59.
doi:\href{https://doi.org/10.1037/0033-2909.112.1.155}{10.1037/0033-2909.112.1.155}.

\hypertarget{ref-Gerber1999}{}
Gerber, L R, D P DeMaster, and P M Kareiva. 1999. ``Gray whales and the
value of monitering data in implementing the U.S. endangered species
act.'' \emph{Conservation Biology} 13 (5): 1215--9.

\hypertarget{ref-Gerrodette1987}{}
Gerrodette, Tim. 1987. ``A power analysis for detecting trends.''
doi:\href{https://doi.org/10.2307/1939220}{10.2307/1939220}.

\hypertarget{ref-Giron-Nava2017}{}
Giron-Nava, Alfredo, Chase C James, Andrew F Johnson, David Dannecker,
Bethany Kolody, Adrienne Lee, Maitreyi Nagarkar, et al. 2017.
``Quantitative argument for long-term ecological monitoring.''
\emph{Marine Ecology Progress Series} 572: 269--74.

\hypertarget{ref-Hatch2003}{}
Hatch, S A. 2003. ``Statistical power for detecting trends with
applications to seabirds monitoring.'' \emph{Biological Conservation}
111: 317--29.

\hypertarget{ref-Hughes2017}{}
Hughes, Brent B, Rodrigo Beas-luna, Allison K Barner, Kimberly Brewitt,
Daniel R Brumbaugh, Elizabeth B. Cerny-Chipman, Sarah L. Close, et al.
2017. ``Long-term studies contribute disproportionately to ecology and
policy.'' \emph{BioScience} 67 (3): 271--81.
doi:\href{https://doi.org/10.1093/biosci/biw185}{10.1093/biosci/biw185}.

\hypertarget{ref-IUCN2012}{}
IUCN. 2012. ``IUCN Red List Categories and Criteria: Version 3.1.''
doi:\href{https://doi.org/10.9782-8317-0633-5}{10.9782-8317-0633-5}.

\hypertarget{ref-Johnson2015}{}
Johnson, Paul C, Sarah J Barry, Heather M Ferguson, and Pie Müller.
2015. ``Power analysis for generalized linear mixed models in ecology
and evolution.'' \emph{Methods in Ecology and Evolution} 6 (2): 133--42.
doi:\href{https://doi.org/10.1111/2041-210X.12306}{10.1111/2041-210X.12306}.

\hypertarget{ref-Keith2015}{}
Keith, David, H. Resit Akçakaya, Stuart H.M. Butchart, Ben Collen,
Nicholas K. Dulvy, Elizabeth E. Holmes, Jeffrey A. Hutchings, et al.
2015. ``Temporal correlations in population trends: Conservation
implications from time-series analysis of diverse animal taxa.''
\emph{Biological Conservation} 192. Elsevier B.V.: 247--57.
doi:\href{https://doi.org/10.1016/j.biocon.2015.09.021}{10.1016/j.biocon.2015.09.021}.

\hypertarget{ref-Legg2006}{}
Legg, Colin J, and Laszlo Nagy. 2006. ``Why most conservation monitoring
is, but need not be, a waste of time.'' \emph{Journal of Environmental
Management} 78: 194--99.
doi:\href{https://doi.org/10.1016/j.jenvman.2005.04.016}{10.1016/j.jenvman.2005.04.016}.

\hypertarget{ref-Magurran2010}{}
Magurran, Anne E, Stephen R Baillie, Stephen T Buckland, Jan Mcp Dick,
David A Elston, E Marian Scott, Rognvald I Smith, Paul J Somerfield, and
Allan D Watt. 2010. ``Long-term datasets in biodiversity research and
monitoring : assessing change in ecological communities through time.''
\emph{Trends in Ecology and Evolution} 25: 574--82.
doi:\href{https://doi.org/10.1016/j.tree.2010.06.016}{10.1016/j.tree.2010.06.016}.

\hypertarget{ref-Morrison2008}{}
Morrison, Shawm, and David S. Hik. 2008. ``When? Where? And for how
long? Census design considerations for an Alpine Lagomorph, the Collared
pika.'' In \emph{Lagomorph Biology}, 103--13. Springer Berlin
Heidelberg.
doi:\href{https://doi.org/10.1007/978-3-540-72446-9}{10.1007/978-3-540-72446-9}.

\hypertarget{ref-Myhrvold2015}{}
Myhrvold, Nathan P., Elita Baldridge, Benjamin Chan, Dhileep Sivam,
Daniel L. Freeman, and S.K. Morgan Ernest. 2015. ``An amniote
life-history database to perform comparative analyses with birds,
mammals, and reptiles.'' \emph{Ecology} 96 (11): 3109.

\hypertarget{ref-GPDD2010}{}
NERC Centre for Population Biology, Imperial College. 2010. ``The Global
Population Dynamics Database Version 2.''

\hypertarget{ref-Rhodes2011}{}
Rhodes, Jonathan R., and Niclas Jonzen. 2011. ``Monitoring temporal
trends in spatially structured populations: how should sampling effort
be allocated between space and time?'' \emph{Ecography} 34 (6): 1040--8.
doi:\href{https://doi.org/10.1111/j.1600-0587.2011.06370.x}{10.1111/j.1600-0587.2011.06370.x}.

\hypertarget{ref-Rueda-Cediel2015}{}
Rueda-Cediel, Pamela, Kurt E Anderson, Tracey J Regan, Janet Franklin,
and M Regan. 2015. ``Combined influences of model choice, data quality,
and data quantity when estimating population trends.'' \emph{PLoSONE} 10
(7): e0132255.
doi:\href{https://doi.org/10.1371/journal.pone.0132255}{10.1371/journal.pone.0132255}.

\hypertarget{ref-Seavy2007}{}
Seavy, Nathaniel E., and Michelle H. Reynolds. 2007. ``Is statistical
power to detect trends a good assessment of population monitoring?''
\emph{Biological Conservation} 140 (1-2): 187--91.
doi:\href{https://doi.org/10.1016/j.biocon.2007.08.007}{10.1016/j.biocon.2007.08.007}.

\hypertarget{ref-White2014}{}
White, Easton R, John D Nagy, and Samuel H Gruber. 2014. ``Modeling the
population dynamics of lemon sharks.'' \emph{Biology Direct} 9 (23):
1--18.

\hypertarget{ref-Zuur2009}{}
Zuur, Alain F., Elena N. Ieno, Neil J. Walker, Anatoly A. Saveliev, and
Graham M. Smith. 2009. \emph{Mixed Effects Models and Extensions in
Ecology with R}. New York: Springer.

\section{Figures}\label{figures}

stuff here \ref{fig:empirical_approach_example}

\clearpage

\section{Supplementary material}\label{supplementary-material}

\setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}

\begin{figure}[htbp]
\centering
\caption{(a) Population size of Bigeye tuna (\emph{Thunnus obesus}) over
time. The line is the best fit line from linear regression. (b)
Statistical power for different subsets of the time series in panel
a.\label{fig:empirical_approach_example}}
\end{figure}

\begin{figure}[htbp]
\centering
\caption{Output of generalized linear model with a Poisson error
structure for predicting the minimum time required with explanatory
variables of the absolute value of the slope coefficient (or trend
strength), temporal autocorrelation, and variability in population
size.\label{fig:poisson_model}}
\end{figure}

\begin{figure}[htbp]
\centering
\caption{Minimum time required versus (a) generation length (years), (b)
litter size (n), (c) adult body mass (grams), (d) maximum longevity
(years), (e) egg mass (grams), and (f) incubation (days). The lines in
each plot represent the best fit line from linear
regression.\label{fig:biological_correlates}}
\end{figure}

\begin{figure}[htbp]
\centering
\caption{(a) Minimum time required to estimate change in abundance for
species class, (b) long-term trend (estimated slope coefficient) by
species class, (c) inter-annual variability in population size by species
class, and (d) temporal autocorrelation by species
class.\label{fig:class}}
\end{figure}

\begin{figure}[htbp]
\centering
\caption{Minimum time required to assess long-term trends in abundance
for values of statistical significance (\(\alpha\)) and power
(\(1-\beta\)).\label{fig:min_time_vs_alpha_beta}}
\end{figure}

\begin{figure}[htbp]
\centering
\caption{Distribution of the minimum time required in order to detect a
significant trend (at the 0.05 level) in log(abundance) given power of
0.8.\label{fig:min_time_dist_log_pop}}
\end{figure}

\begin{figure}[htbp]
\centering
\caption{Statistical power for different length of time series
simulations for a lemon shark population in Bimini,
Bahamas.\label{fig:shark_example}}
\end{figure}

\begin{figure}[htbp]
\centering
\caption{Example of calculating minimum time required for growth rate
estimation. (a) European herring gull (\emph{Larus argentatus}) scaled
density over time, (b) mean and standard deviation of growth rate for
subsamples of entire time series, and (c) the percent error between mean
estimated growth rate and the true long-term growth rate. The vertical
bar denotes the minimum time required to estimate growth rate within
20\% error.\label{fig:growth_rate}}
\end{figure}

\begin{figure}[htbp]
\centering
\caption{Histogram of the minimum time required in order to estimate the
long-term growth rate within 20\%
error.\label{fig:min_time_growth_dist}}
\end{figure}

\begin{figure}[htbp]
\centering
\caption{(a) Time series for Bigeye tuna (\emph{Thunnus obesus}) with
corresponding fitted GAM model in red and (b) statistical power as a
function of the number of years sampled. The horizontal line at 0.8
indicates the minimum threshold for statistical power and the vertical
line denotes the minimum time required to achieve 0.8 statistical
power.\label{fig:gam_example}}
\end{figure}

\begin{figure}[htbp]
\centering
\caption{Distribution of the minimum time required in order to detect a
significant trend (at the 0.05 level) in abundance according to a GAM
model given statistical power of 0.8. The smoothing parameter was set to
3 for each population.\label{fig:min_time_dist_gam}}
\end{figure}

\end{document}