### 2018-02-19 writing/manuscript_draft.Rmd
first update of short format version of the article

---
bibliography: White_bib.bib
fontsize: 12pt
geometry: margin=1in
header-includes: \usepackage{float} \usepackage{lineno} \usepackage[firstinits=true,style=authoryear-comp,natbib=true,doi=false,isbn=false,url=false,uniquename=false,uniquelist=false,sorting=none]{biblatex}

output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
  word_document: default
---

Title: Minimum time required to detect population trends: the need for long-term monitoring programs
\vspace{7 mm}

Author: \textsc{Easton R. White$^{1,2}$} 
\vspace{3 mm}

Address:\emph{
    \\$^1$Center for Population Biology \\
    University of California, Davis \\
    2320 Storer Hall \\
		University of California, Davis \\
		One Shields Avenue Davis, CA, USA}
\vspace{3 mm}

$^2$Corresponding author: eawhite@ucdavis.edu , 1 480 203 7931
\vspace{3 mm}

Number of words (including 300 words per figure):  6,950
\vspace{1 mm}

Number of figures: 5 
\vspace{1 mm}

Number of references: 34
\vspace{1 mm}

Supplementary material: 9 figures and 3 tables
\vspace{3 mm}

Data and code for all the figures and tables can be found at (https://github.com/erwhite1/time-series-project).

To be submitted to: \emph{Biological Conservation}
\vspace{3 mm}

Available as a preprint at: https://peerj.com/preprints/3168/
\vspace{3 mm}

Color does not need to be used for any figures in print
\vspace{3 mm}

\vspace{3 mm}

<!-- Shorter title for PeerJ
 \textbf{Minimum time required to detect population trends: the need for long-term monitoring programs}
 \vspace{3 mm}

Easton R. White$^{*}$

\begin{footnotesize} Center for Population Biology, University of California, Davis, California 95616 USA

*eawhite@ucdavis.edu
\end{footnotesize}

\vspace{5 mm}
-->
\linenumbers

\begin{abstract}
Long-term time series are necessary to better understand population dynamics, assess species' conservation status, and make management decisions. However, population data are often expensive, requiring a lot of time and resources. When is a population time series long enough to address a question of interest? I determine the minimum time series length required to detect significant increases or decreases in population abundance. To address this question, I use simulation methods and examine 822 populations of vertebrate species. Here I show that on average 15.9 years of continuous monitoring are required in order to achieve a high level of statistical power. However, there is a wide distribution around this average, casting doubt on simple rules of thumb. For both simulations and the time series data, the minimum time required depends on trend strength, population variability, and temporal autocorrelation. However, there were no life-history traits (e.g. generation length) that were predictive of the minimum time required. These results point to the importance of sampling populations over long periods of time. I argue that statistical power needs to be considered in monitoring program design and evaluation. Short time series are likely under-powered and potentially misleading.
\end{abstract}

Keywords: ecological time series, experimental design, monitoring, power analysis, statistical power, sampling design

# Introduction

Observational studies and population time series have become a cornerstone of modern ecological research and conservation biology [@Magurran2010; @Hughes2017]. Long-term data are necessary to both understand population dynamics and to assess species extinction risk. Even though many time series may now be considered "long-term" (e.g. continuous plankton recorder, @Giron-Nava2017), most are still short. Time series are typically short for a variety of reasons [@Field2007]. They are often coupled with an experiment, which may only last a couple of years. In addition, short funding cycles make it difficult to examine populations over longer periods of time [@Hughes2017].

How long of a time series is actually necessary? This question has important implications for both research and management [@Nichols2006]. Scientists need to know the time series length required to address a specific question. A short time series may lead to wrong conclusions given large natural year-to-year variability [@McCain2016]. Managers need to know when action is needed for a population. Therefore, managers must understand when population trend over time is actually meaningful. For example, the International Union for Conservation of Nature (IUCN) Red List Categories and Criteria suggest, under Criterion A2, a species qualifies as vulnerable if it has experienced a 30% decline over 10 years, or three generations [@IUCN2012]. For both scientific and management questions, because sampling is typically expensive, we also do not want to sample for longer than is necessary. For example, @Gerber1999 investigated the minimum time series required to estimate population growth of the endangered, but recovering, eastern North Pacific gray whale (\emph{Eschrichtius robustus}). They used a long-term census to retroactively determine the minimum time series required to assess threat status. They found that only 11 years were needed, eight years before the delisting decision was made. This highlights the importance of estimating the minimum time series required as an earlier decision would have saved time and money [@Gerber1999]. Further, waiting too long to make a decision can imperil a species where management action could have been taken earlier [@Martin2012; @Martin2017].

An important step in experimental design is to determine the number of samples required. For any particular experiment four quantities are intricately linked: significance level ($\alpha$), statistical power, effect size, and sample size [@Legg2006]. The exact relationship between these quantities depends on the specific statistical test. A type I error is a false positive, or incorrect rejection of a true null hypothesis. For example, if a time series was assessed as significantly increasing or decreasing---when there was no true significant trend---this would be a false positive. The false positive rate, or significance level ($\alpha$) is often set at 0.05 (although this is purely historical, @Mapstone1995). A type II error ($\beta$) is a failure to detect a true trend, or failure to reject a false null hypothesis. Formally, statistical power ($1-\beta$) is one minus the probability of a type II error ($\beta$). The effect size is a measure of the difference between two groups. Prior to an experiment, one could set appropriate levels of power, significance level, and the effect size to estimate the sample size required for the experiment. This approach, however, is not straight-forward for a time series, or more complicated scenarios [@Johnson2015], as data are clearly non-independent.

For time series data, two general approaches to estimating sample size are appropriate. <!--Typically, calculations of sample size are given by a formula specific to standard statistical tests (e.g. t-test). However, many ecological scenarios require the use of simulations [@Bolker2008; @Johnson2015]. As a first approach,-->Simulations can be designed for a specific population and question [@Bolker2008; @Johnson2015]. Simple models can be simulated with parameter values corresponding to a population of interest [@Gerrodette1987]. Statistical power is the proportion of simulations that meet some set of criteria. The specific criteria depend on the question at hand. For example, given a time series, when is the slope from linear regression significantly different from zero? In other words, when is the time series significantly increasing or decreasing? It is then possible to determine how power changes with a variable of interest. For example, time series can be simulated for different lengths of time. From these simulations, the minimum time series length required to meet certain levels of statistical significance and power is estimated [@Bolker2008].

In addition to using simulations, empirical time series can also be used. Multiple replicates of similar populations are usually not available, but it is possible to subsample an empirical time series [@Gerber1999; @Brashares2005]. Subsamples of different lengths can then be evaluated to estimate the proportion of subsamples meeting some criteria, again a measure of statistical power. Similar to the simulation approach, this measures of power can be used to determine the minimum time series required for a particular question of interest.

Past work has investigated questions related to the minimum time series required to estimate trends in population size over time [@Wagner2009; @Giron-Nava2017]. For example, @Rhodes2011 examined the optimal allocation of effort between spatial and temporal replicates. Using simple populations models, they found that the allocation of effort depends on environmental variation, spatial and temporal autocorrelation, and observer error. @Rueda-Cediel2015 also used a modeling approach, but parameterized a model specific for a threatened snail, \emph{Tasmaphena lamproides}. They found that for this short-lived organism, 15 years was adequate to assess long-term trends in abundance. However, these studies, and other past work, have typically focused either on theoretical aspects of monitoring design or focused on only a few species.

I use both simulations and empirical time series to determine the minimum number of years required to address several questions. I estimate the minimum time series length required ($T_{min}$) to assess long-term changes in abundance via simple linear regression. First, I estimate $T_{min}$ using a simulation approach. Then I examine 822 population time series to estimate $T_{min}$. In the supplementary material, I determine $T_{min}$ for related ideas: using more complicated population models, varying statistical level and power, and the use of generalized additive models.

# Methods

## Simulation approach

One approach to determining the minimum time series length needed is through repetitive simulations of a population model [@Gerrodette1987]. This is the same approach one might use in sample size calculations for any experimental design too complicated for simple power analyses [@Bolker2008;@Johnson2015]. I only briefly discuss this approach as it has been described elsewhere. Essentially, a population model is simulated repetitively for a number of years. This approach requires us to determine values for model parameters (e.g. population variability). As an example, we can take the following population model for population size $N$ at time $t$:

\begin{equation}
N(t + 1) = N(t) + r(t) + \epsilon \mbox{ with } \epsilon \sim N(\mu, \sigma)
\label{sim_model}
\end{equation}
where $\epsilon$ is a normally-distributed random noise term with mean $\mu$ and standard deviation $\sigma$. The rate of growth $r$ is also the trend strength of the increase or decrease (i.e. the rate of increase). It is important to note that any population model could be substituted for equation \ref{sim_model}, as in the supplementary material (Figs. \ref{fig:min_time_dist_log_pop}, \ref{fig:shark_example}).

Statistical power is then the proportion of simulations that meet some criteria. Here, our criteria is whether the slope parameter from linear regression is significant at the 0.05 threshold with statistical power of 0.8. Statistical power of 0.8 would indicate that, if there was a true trend in abundance, there would be a 0.8 probability of detecting the trend. Values of 0.05 for the significance level and 0.8 statistical power are historical and it is important to examine the effect of changing these values (Fig. \ref{fig:min_time_vs_alpha_beta}).

In Fig. \ref{fig:theoretical_approach}a, a number of simulated time series are shown for a set number of time periods ($t=40$). It is clear that statistical power increases quickly with increases in length of time sampled (Fig. \ref{fig:theoretical_approach}b). Where power is greater than 0.8 (the dotted line), that is the minimum time required ($T_{min}$) to be confident in the detection of a long-term trend in abundance.

```{r,echo=F,eval=T,fig.cap='(a) Example of a simulated time series for 40 time periods. (b) Statistical power versus the simulated time series length. The horizontal, dashed line is the desired statistical power of 0.8. The vertical, dashed line is the minimum time required to achieve the desired statistical power. (c) Minimum time required ($T_{min}$) for simulations with different values of the trend strength ($r$). (d) Minimum time required for different levels of population variability ($\\sigma$). In each case, the minimum time required is the minimum number of years to achieve 0.8 statistical power given a significance level of 0.05.\\label{fig:theoretical_approach}'}
source("../scripts/theoretical-models/calculate_power.R")
source("../scripts/theoretical-models/simulate_pop.R")

  # years_to_sample=seq(2,100,by=1)
  # power=lapply(X=years_to_sample,FUN = calculate_power,phi = 0.6,r=0.2,sigma = .5,trials=100)#should use 300+ trials
  # 
  # simulations = replicate(n=30,simulate_pop(phi = 0.6,r=0.2,sigma = .5,max.time=100))
  # min_time_needed = years_to_sample[tail(which(power<0.8),1)] + 1
  # 
  # save(simulations,years_to_sample,power,min_time_needed,file='../analysis-outputs/example_simulations_fig1.Rdata')
  load('../analysis-outputs/example_simulations_fig1.Rdata')

  #load command
  # Plotting commands
  par(mfrow=c(2,2),oma=c(1.5,2,0.5,0.5),mar=c(3,5,2.5,0.5))
    matplot(simulations,type='l',col=rgb(0.5,0.5,0.5,0.5),lty=1,ylab='',xlab='',las=1,cex.axis=1.2)
      mtext(text = 'time' ,side = 1,cex = 1.2,line=3)
      mtext(text = 'population size' ,side = 2,cex = 1.2,line=3.5)
      mtext(text = '(a)',side = 3,line = -1.2,adj=0.05,font = 2,cex=1.2)

    plot(years_to_sample,power,type='l',lwd=2,las=1,ylab='',xlab='',cex.axis=1.2)
      abline(v=min_time_needed,lty=2,lwd=2,col='black')
      abline(h=0.8,lty=2,lwd=2,col='black')
      mtext(text = 'time sampled' ,side = 1,cex = 1.2,line=3)
      mtext(text = 'statistical power' ,side = 2,cex = 1.2,line=3)
      mtext(text = '(b)',side = 3,line = -1.2,adj=0.05,font = 2,cex=1.2)

    #   
    # # Plot versus trend strength
    #   r_values = seq(1,3,by=0.1) #by=0.2
    #   sigma_values=seq(1,7,by=0.3) #by=0.2
    # 
    #   parameter_combinations = expand.grid(r_values,5)
    #   parameter_combinations$min_time_required = NA
    #   names(parameter_combinations)= c('r','sigma','min_time_required')
    #   
    #   for (i in 1:nrow(parameter_combinations)){
    #     years_to_sample=seq(2,50,by=1)
    #     y=lapply(X=years_to_sample,FUN = calculate_power,phi =0.5,r=parameter_combinations$r[i],sigma = parameter_combinations$sigma[i],trials=150)
    #     if (length(tail(which(y<0.8),1))>0){
    #     parameter_combinations$min_time_required[i] = years_to_sample[tail(which(y<0.8),1)] + 1
    #     }
    #     #print(i)
    #   }
    #   
    #   save(parameter_combinations,file='../analysis-outputs/min_time_vs_trend_strength.Rdata')
      load('../analysis-outputs/min_time_vs_trend_strength.Rdata')

      plot(parameter_combinations$r,parameter_combinations$min_time_required,type='l',lwd=2,las=1,ylab='',xlab='',cex.axis=1.2,ylim=c(0,50))
      mtext(text = 'trend strength (slope coefficient)' ,side = 1,cex = 1.2,line=3)
      mtext(text = expression(T[min]) ,side = 2,cex = 1.2,line=3)
      mtext(text = '(c)',side = 3,line = -1.2,adj=0.05,font = 2,cex=1.2)

      # 
      # 
      # 
      # # Plot versus sigma values
      # parameter_combinations = expand.grid(1.5,sigma_values)
      # parameter_combinations$min_time_required = NA
      # names(parameter_combinations)= c('r','sigma','min_time_required')
      # 
      # for (i in 1:nrow(parameter_combinations)){
      #   years_to_sample=seq(2,50,by=1)
      #   y=lapply(X=years_to_sample,FUN = calculate_power,phi =0.5,r=parameter_combinations$r[i],sigma = parameter_combinations$sigma[i],trials=150)
      #   if (length(tail(which(y<0.8),1))>0){
      #   parameter_combinations$min_time_required[i] = years_to_sample[tail(which(y<0.8),1)] + 1
      #   }
      #   #print(i)
      # }

      #save(parameter_combinations,file='../analysis-outputs/min_time_vs_pop_var.Rdata')
      load('../analysis-outputs/min_time_vs_pop_var.Rdata')
      plot(parameter_combinations$sigma,parameter_combinations$min_time_required,type='l',lwd=2,las=1,ylab='',xlab='',cex.axis=1.2,ylim=c(0,50))
      mtext(text = 'population variability' ,side = 1,cex = 1.2,line=3)
      mtext(text = expression(T[min]) ,side = 2,cex = 1.2,line=3)
      mtext(text = '(d)',side = 3,line = -1.2,adj=0.05,font = 2,cex=1.2)
```

<!-- code to pull in databases-->
```{r, echo=F}
# Pull in original database from Keith et al. 2015 and the cleaned version after our series of filtering steps

#Load keith et al. database (from GPDD database)
  load('../data/Keith_et_al_2015_data.Rdata')

#Load our cleaned data (see other code for cleaning steps)
  load('../cleaned-data/cleaned_timeseries_database_with_min_time_linear_regression.Rdata') #Do i need to filter out non-significant populations as well (where no trend was present)
 # pop_info=pop_info[pop_info$trend_p_value<0.05,]
 # long_dat=subset(long_dat,long_dat$ID %in% pop_info$ID)
 # long_time_names=subset(long_time_names,long_time_names %in% pop_info$ID)
```

## Data source
I use a database of `r length(table(dat$ID))` population time series compiled in [@Keith2015]; they compared the predictability of growth rates among populations. The data are originally from the Global Population Dynamics Database [@GPDD2010] and several other sources [@Keith2015]. I filtered out short time series (less than `r pop_info$series_length[1]` years), and those with missing data, leaving `r nrow(pop_info)` time series. The data includes information on `r length(table(pop_info$Binomial))` vertebrate species with a focus on mammals, birds, and fish. The data also includes information on generation length and survey specifications. For each time series, I also calculate other variables of interest: coefficient of variance in population size, long-term trend in abundance (slope coefficient from simple linear regression), and temporal autocorrelation. All analyses were conducted in R [@RCoreTeam2016].

```{r,echo=F}
# Pull in results of biological characteristics comparison to min time series length required
load('../cleaned-data/combined_databases_with_min_time_linear_regression.Rdata')
```
For a subset of populations ($n =$ `r nrow(LPI_pop_info)`), there is information on life-history traits available from another paper [@Myhrvold2015], including body size and generation length. All `r nrow(LPI_pop_info)` populations were birds. I examine how the minimum time required is related to these life-history traits (Fig. \ref{fig:biological_correlates}).

## Empirical approach

I assume that each time series is long enough to include all necessary information (e.g. variability) about the population. In other words, each time series is a representative sample. I first take all possible contiguous subsamples of each time series. For example, a time series of 35 years would have 34 possible contiguous subsamples of length 2, 34 possible contiguous subsamples of length 3, and continuing until 1 possible contiguous subsample of length 35 [@Gerber1999; @Giron-Nava2017]. Next, I run a linear regression for each subsampled time series. Then, I determine the proportion of subsamples of a particular length that have estimated slope coefficients which are statistically different from zero. I only look at the proportion of samples where the long-term, or ``true'', time series also has a significant slope. This proportion is a measure of statistical power. Lastly, I determine which subsample length is required to achieve a certain threshold of statistical power (0.8, @Cohen1992). The minimum subsampled length that met these criteria is the minimum time series length required ($T_{min}$).

In the supplementary material, I show how the same approach described here for more complicated population models. I also determine the minimum time required to estimate long-term trends according to generalized additive models, instead of the simple linear models used here (Fig. \ref{fig:gam_example}).

<!--
## Questions addressed in the appendix

I also examine how minimum time series calculations changed depending on the question of interest. In the supplementary material, I use similar calculations to determine the minimum time series required to estimate overall population growth rate (Fig. \ref{fig:growth_rate}). I also explore how the method can be applied to more biologically-realistic population models (Fig. \ref{fig:shark_example}. More specifically, I determine the minimum time required to estimate long-term population trends using a stochastic, age-structured model of lemon shark population dynamics in the Bahamas [@White2014]. Lastly, I determine the minimum time required to estimate long-term trends according to generalized additive models, instead of the simple linear models used here.
-->

# Results

I determined the minimum time series length ($T_{min}$) required to address a particular question of interest. What is the minimum time series length required to determine, via linear regression, the long-term population trend? Here, the minimum time series length required had high enough statistical power (greater than 0.8) for a set significance level ($\alpha$) of 0.05. It is also possible to alter statistical power and $\alpha$. Predictably, with increased statistical power or decreased $\alpha$, $T_{min}$ increased (Fig. \ref{fig:min_time_vs_alpha_beta}). I then estimated $T_{min}$ using two approaches. I briefly describe results from the simulation approach and then discuss the empirical approach.

## Simulation approach

I constructed a general population model where the trend strength (i.e. slope coefficient) over time could be a model parameter. I then simulated time series of different lengths. From these simulations I determined the minimum time series length required to achieve a certain level of statistical power. In line with past work [@Gerrodette1987], I found the $T_{min}$ increases (i.e. more time is required) with decreases in trend strength and with increases in population variability (Figs. \ref{fig:theoretical_approach}c,d).

I chose a simple model, but any other population model could be used (see example in Fig. \ref{fig:min_time_dist_log_pop}). Ideally, the specific model choice should be tailored to the population of interest. I explored how the simulation approach could be applied to more biologically-realistic population models (Fig. \ref{fig:shark_example}). Specifically, I determined the minimum time required to estimate long-term population trends using a stochastic, age-structured model of lemon shark population dynamics in the Bahamas [@White2014]. I found that over 27 years of continuous monitoring were needed in this particular scenario (Fig. \ref{fig:shark_example}). Similar to the simulation approach described above, the minimum time required for the lemon shark population was strongly dependent on model parameters.

## Empirical approach

```{r,echo=F,message=F,warning=F,fig.cap='(a) Distributions of the minimum time required for populations from four different biological classes. (b) Distribution of minimum time required for all populations regardless of biological class. The minimum time required calculation corresponds to a significance level of 0.05 and statistical power of 0.8.\\label{fig:min_time_dist}',fig.height=6}
# Pull in analysis of average min time needed for **empirical species** (save analysis as Rdata file or something)
 load("../cleaned-data/cleaned_timeseries_database_with_min_time_linear_regression.Rdata")

require(ggplot2)
plot1<- ggplot(aes(class,min_time_for_power),data=pop_info) + geom_violin() + geom_jitter(col=rgb(0.5,0.5,0.5,0.3),height = 0, width = 0.1) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"),axis.text = element_text(size=12),axis.title.x =element_text(size=16) ,axis.title.y =element_text(size=16) ) +
 labs(x='biological class',y='minimum time required') +
  geom_text(aes(x=4.4,y=33),label='(a)',size=6) +
  scale_y_continuous(limits = c(0, 40))

library(png)
library(grid)

fish <- annotation_custom(rasterGrob(readPNG("phylo-fish.png"), interpolate=TRUE), xmin=-2.5,ymin=35,ymax=40) 
bird <- annotation_custom(rasterGrob(readPNG("phylo-bird.png"), interpolate=TRUE), xmin=-0.5, ymin=35) 
shark <- annotation_custom(rasterGrob(readPNG("phylo-shark.png"), interpolate=TRUE), xmin=2.6,xmax=3.3, ymin=35) 
mammal <- annotation_custom(rasterGrob(readPNG("phylo-mammal.png"), interpolate=TRUE), xmin=3.3, ymin=35) 
plot1 <- plot1+ fish + bird + shark + mammal

#counts <- table(pop_info$class,pop_info$min_time_for_power)
plot2<- ggplot(aes(min_time_for_power),data=pop_info) + geom_bar() +  
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"),axis.text = element_text(size=12),axis.title.x =element_text(size=16) ,axis.title.y =element_text(size=16) ) +
 labs(y='count',x='minimum time required')+
  geom_text(aes(x=35,y=47),label='(b)',size=6)

require(gridExtra)
grid.arrange(plot1, plot2, nrow=2)

# Code for barplot in my first submission
# barplot(counts, main=" ",
#         xlab=" ", ylab="", col=gray.colors(4),
#         ylim=c(0,50),xlim=c(0,40),
#         las=1,cex.names =1.2,cex.axis = 1.2,#density=35,
#         angle=c(20,40,60,80))
# legend('topright',legend=rownames(counts),col = gray.colors(4),pch = 15,bty ="n")
# 
# mtext(text = 'Frequency',2,line=3,cex=1.2)
# mtext(text = 'Minimum time required',1,line=3,cex=1.2)
```

```{r poisson_model, echo=F,include=F}
# I should put this in the appendix
#Poisson regression
pop_info$abs_overall_trend = abs(pop_info$overall_trend)
time_series_correlates_poisson_model <-glm(min_time_for_power~ abs_overall_trend + autocorrelation + coefficient_variation,family = 'poisson',data=pop_info)
#par(mfrow=c(2,2),mar=c(5,5,1,2),oma=c(0,0,0.5,0))
#plot(poisson_model,cex.lab=1.2,las=1,cex.axis=1.2,main=' ')

#R^2 calculation (pg. 218 Zuur)
R2 <- 100*(time_series_correlates_poisson_model$null.deviance - time_series_correlates_poisson_model$deviance)/(time_series_correlates_poisson_model$null.deviance)

```

```{r,echo=F, fig.cap='Minimum time required to estimate change in abundance correlated with (a) trend strength (absolute value of slope coefficient estimated from linear regression), (b) coefficient of variation in interannual population size, and (c) temporal lag-1 autocorrelation.\\label{fig:correlates}',fig.height=7, fig.width=4}

# Plot minimum time required against trend strength, population variance, and autocorrelation  
  par(mfrow=c(3,1),mar=c(3,2,2,1),oma=c(1,4.5,0,0.5))
  plot(abs(pop_info$overall_trend),pop_info$min_time_for_power,pch=as.numeric(as.factor(pop_info$class))+14,col=rgb(0.5,0.5,0.5,0.5),las=1,ylab='',xlab='',cex.axis=1.2,cex=1.3,ylim=c(0,35))
  mtext('| trend strength |',1,line=3,cex=1.2)
  mtext('(a)',3,line=-1.5,adj = 0.985,cex=1.2,font = 2)
    plot(pop_info$coefficient_variation,pop_info$min_time_for_power,pch=as.numeric(as.factor(pop_info$class))+14,col=rgb(0.5,0.5,0.5,0.5),las=1,ylab='',xlab='',cex.axis=1.2, xlim = c(0.2,2.2),cex=1.3,ylim=c(0,35))
    mtext('coefficient of variation',1,line=3,cex=1.2)
    mtext('(b)',3,line=-1.5,adj = 0.985,cex=1.2,font = 2)
      plot(pop_info$autocorrelation,pop_info$min_time_for_power,pch=as.numeric(as.factor(pop_info$class))+14,col=rgb(0.5,0.5,0.5,0.5),las=1,ylab='',xlab='',cex.axis=1.2, xlim = c(-0.2,1),cex=1.3,ylim=c(0,35))
      mtext('autocorrelation',1,line=3,cex=1.2)
       mtext('(c)',3,line=-1.5,adj = 0.985,cex=1.2,font = 2)
       mtext('minimum time required',2,line=2.5,cex=1.2,outer = T)

legend('bottomleft',legend=names(table(pop_info$class)),col = 'grey',pch = as.numeric(names(table(as.numeric(as.factor(pop_info$class))+14))),bty ="n",cex=1.2)

```

I examined a database of `r nrow(pop_info)` separate population time series representing `r length(table(pop_info$Binomial))` species. This database consisted of vertebrate species with a variety of life-history characteristics (Fig. \ref{fig:biological_correlates}). I limited analyses to populations with at least 35 years of continuous sampling. I then examined the minimum time required to estimate long-term trends via linear regression.

Across all the populations, I found an average minimum time series length required ($T_{min}$) of `r round(mean(pop_info$min_time_for_power),1)` (SD=`r round((var(pop_info$min_time_for_power))^0.5,1)`), with a wide distribution (Fig. \ref{fig:min_time_dist}b). Estimates of $T_{min}$ varied between biological class (Fig. \ref{fig:min_time_dist}a). Ray-finned fish (class Actinopterygii) typically had estimates of $T_{min}$ over 20 years. Birds (class Aves) had a much wider distribution of $T_{min}$, but usually required less years of sampling. Differences between these classes can be explained by differences in variability in population size and strength of trends in abundance (Fig. \ref{fig:class}).

### Corrrelates for minimum time required

The minimum time series length required was strongly correlated with trend strength (i.e. estimated slope coefficient from linear regression), coefficient of variation in population size, and autocorrelation in population size (Fig. \ref{fig:correlates}). This is in line based on simulations here and those of others [@Rhodes2011]. Using a generalized linear model, with a Poisson error structure, all three of these explanatory variables were significant and had large effect sizes (see Table A1). Combined, trend strength, coefficient of variation in population, and autocorrelation account for `r round(100*(time_series_correlates_poisson_model$null.deviance - time_series_correlates_poisson_model$deviance)/(time_series_correlates_poisson_model$null.deviance),1)`% of the explained deviance [@Zuur2009] in minimum time series length required.

<!-- Minimum time required and biological correlates-->
```{r, echo=F}
load('../cleaned-data/combined_databases_with_min_time_linear_regression.Rdata')
```

For a subset of the populations I combined time series data with a data on life-history characteristics of amniotes [@Myhrvold2015]. There was life-history information available for `r nrow(LPI_pop_info)` populations representing `r length(table(LPI_pop_info$Binomial))` different species, all of which were birds (Aves class).

```{r, echo=F, fig.cap='Mimimum time required versus (a) generation length (years), (b) litter size (n), (c) log adult body mass (grams), (d) maximum longevity (years), and (e) incubation (days). The lines in each plot represent the best fit line from linear regression.\\label{fig:biological_correlates}',fig.height=6}

load('../cleaned-data/combined_databases_with_min_time_linear_regression.Rdata')
par(mfrow=c(3,2),oma=c(5,4,1,1),mar=c(5,3,0,0))
#x_axis_name_list = c('generation length (years)','litter size (n)','adult body mass (grams)','maximum longevity (years)','egg mass (grams)','incubation (days)')

  plot(LPI_pop_info$gen_len,LPI_pop_info$min_time_for_power,pch=as.numeric(as.factor(pop_info$class))+14,col=rgb(0.5,0.5,0.5,0.5),las=1,ylab='',xlab='',cex.axis=1.2,cex=1.2)
  abline(lm(LPI_pop_info$min_time_for_power~LPI_pop_info$gen_len),lwd=3)
  mtext(text = 'generation length (years)',side = 1,line = 3,cex = 1.2)
  mtext(text = "(a)",side = 3,line=-1.5,adj = 0.98,cex=1.2)

  plot(LPI_pop_info$litter_or_clutch_size_n,LPI_pop_info$min_time_for_power,pch=as.numeric(as.factor(pop_info$class))+14,col=rgb(0.5,0.5,0.5,0.5),las=1,ylab='',xlab='',cex.axis=1.2,cex=1.2)
  abline(lm(LPI_pop_info$min_time_for_power~LPI_pop_info$litter_or_clutch_size_n),lwd=3)
  mtext(text = 'litter size (n)',side = 1,line = 3,cex = 1.2)
  mtext(text = "(b)",side = 3,line=-1.5,adj = 0.98,cex=1.2)

    plot(log(LPI_pop_info$adult_body_mass_g),LPI_pop_info$min_time_for_power,pch=as.numeric(as.factor(pop_info$class))+14,col=rgb(0.5,0.5,0.5,0.5),las=1,ylab='',xlab='',cex.axis=1.2,cex=1.2)
  abline(lm(LPI_pop_info$min_time_for_power~LPI_pop_info$adult_body_mass_g),lwd=3)
  mtext(text = 'log(adult mass (grams))',side = 1,line = 3,cex = 1.2)
  mtext(text = "(c)",side = 3,line=-1.5,adj = 0.98,cex=1.2)

  plot(LPI_pop_info$maximum_longevity_y,LPI_pop_info$min_time_for_power,pch=as.numeric(as.factor(pop_info$class))+14,col=rgb(0.5,0.5,0.5,0.5),las=1,ylab='',xlab='',cex.axis=1.2,cex=1.2)
  abline(lm(LPI_pop_info$min_time_for_power~LPI_pop_info$maximum_longevity_y),lwd=3)
  mtext(text = 'maximum longevity (years)',side = 1,line = 3,cex = 1.2)
  mtext(text = "(d)",side = 3,line=-1.5,adj = 0.98,cex=1.2)

  plot(LPI_pop_info$incubation_d,LPI_pop_info$min_time_for_power,pch=as.numeric(as.factor(pop_info$class))+14,col=rgb(0.5,0.5,0.5,0.5),las=1,ylab='',xlab='',cex.axis=1.2,cex=1.2)
  abline(lm(LPI_pop_info$min_time_for_power~LPI_pop_info$incubation_d),lwd=3)
  mtext(text = 'incubation (days)',side = 1,line = 3,cex = 1.2)
  mtext(text = "(e)",side = 3,line=-1.5,adj = 0.98,cex=1.2)

#  boxplot(LPI_pop_info$min_time_for_power~LPI_pop_info$trophic,pch=as.numeric(as.factor(pop_info$class))+14,col=rgb(0.5,0.5,0.5,0.5),las=1,ylab='',xlab='',cex.axis=1.2)
#  stripchart(LPI_pop_info$min_time_for_power~LPI_pop_info$trophic, vertical = TRUE,method = "jitter", add = TRUE, pch = as.numeric(as.factor(pop_info$class))+14, col = rgb(0.5,0.5,0.5,0.5))
#  mtext(text = 'trophic position',side = 1,line = 3,cex = 1.2)
#  mtext(text = "(f)",side = 3,line=-1.5,adj = 0.98,cex=1.2)

plot(1, type="n", xlab="", ylab="", xaxt='n',yaxt='n',axes=FALSE)
legend('center',legend=names(table(pop_info$class)),col = 'grey',pch = as.numeric(names(table(as.numeric(as.factor(pop_info$class))+14))),bty ="n",cex=1.4)

mtext(text = 'minimum time required',side = 2,line = 2,cex = 1.4,outer = TRUE)

#######

biological_correlates_poisson_model <-glm(min_time_for_power~  litter_or_clutch_size_n  + gen_len + adult_body_mass_g + maximum_longevity_y  +incubation_d + trophic,family = 'poisson',data=LPI_pop_info)
R2_biological_correlates_poisson_model <- 100*(biological_correlates_poisson_model$null.deviance - biological_correlates_poisson_model$deviance)/(biological_correlates_poisson_model$null.deviance)

all_correlates_poisson_model <-glm(min_time_for_power~ abs(overall_trend) + coefficient_variation + autocorrelation +  litter_or_clutch_size_n  + gen_len + adult_body_mass_g + maximum_longevity_y  +incubation_d + trophic,family = 'poisson',data=LPI_pop_info)
R2_all_correlates_poisson_model <- 100*(all_correlates_poisson_model$null.deviance - all_correlates_poisson_model$deviance)/(all_correlates_poisson_model$null.deviance)
```

Some life-history traits were significant predictors for the minimum time required (Fig. \ref{fig:biological_correlates}, Tables A2,A3). However, none of these life-history traits explained a large part of the variation in minimum time required. In a generalized linear model, all of the life-history traits described in figure \ref{fig:biological_correlates} account for only `r round(R2_biological_correlates_poisson_model,3)`% of the explained deviance in minimum time series length required. In addition, when accounting for trend strength, coefficient of variation, and autocorrelation, no life-history traits were significant predictors of the minimum time required (Table A3).

### Evalulating the IUCN criteria

```{r, echo=F,fig.cap='Minimum time required to achieve 0.8 statistical power versus the minimum time required under IUCN criteria A2 to classify a species as vulnerable. Each point represents a single population, all of which saw declines of 30% or greater over a 10 year period.\\label{fig:IUCN_analysis}'}

declining_populations = subset(pop_info,pop_info$percent_change_10years>30 & pop_info$overall_trend<0)

par(mfrow=c(1,1),oma=c(0,0,0,0),mar=c(5,5,0.5,0.5))

shapes = as.numeric(as.factor(declining_populations$class))+14
shapes[shapes==17]= 18
plot(declining_populations$IUCN, declining_populations$min_time_for_power,
       ylim=c(0,35),xlim=c(0,85),pch=shapes,
       col=rgb(0.5,0.5,0.5,0.5),cex.lab=1.3,las=1,cex=1.5,
       xlab = 'minimum time required under IUCN criteria',ylab='')
mtext(text = 'minimum time required',side = 2,line = 3.8,outer = F,padj = 0.5,cex=1.3)
mtext(text = 'for 0.8 statistical power',side = 2,line = 2.8,outer = F,padj = 0.5,cex=1.3)
segments(0,0,100,100,lwd=2)

#mtext(text = paste('n = ',nrow(declining_populations),sep=''),side = 3,line = -2,outer = F,at=80,cex=1.2)
legend('bottomright',legend=names(table(declining_populations$class)),col = 'grey',pch = c(15,16,18),bty ="n",cex=1.2)

whale <- readPNG('phylo-sperm-whale.png')
rasterImage(whale,xleft=70,ybottom=12,xright=82,ytop=17)

puffin <- readPNG('phylo-puffin.png')
rasterImage(puffin,xleft=61,ybottom=30,xright=70,ytop=36)
```

The IUCN Red List Categories and Criteria suggest, under Criterion A2, a species qualifies as vulnerable if it has experienced a 30% decline over 10 years, or 3 generations (whichever is longer) [@IUCN2012]. I examined a subset of populations with observed declines of 30% or greater over 10 years, qualifying all of them as vulnerable. This resulted in n = `r nrow(declining_populations)` populations of fish, birds, and a single mammal. I then compared the minimum time required to achieve 0.8 statistical power ($T_{min}$) to the minimum time required under the IUCN criteria (Fig. \ref{fig:IUCN_analysis}). For populations below the identity line in figure \ref{fig:IUCN_analysis}, IUCN criteria would require more sampling compared to estimates for $T_{min}$. Further, populations above the identity line are cases where the IUCN criteria would classify a population as vulnerable despite not having sampled enough years to achieve high statistical power (Fig. \ref{fig:IUCN_analysis}). The silhouettes on figure \ref{fig:IUCN_analysis} highlight that species with longer generation times typically have larger discrepancies between $T_{min}$ and the minimum time required for IUCN assessments (Fig. \ref{fig:IUCN_correlates}).

### Sensitivity analysis
Lastly, I tested model sensitivity by using generalized additive models (GAMs) instead of simple linear regression. Again, I examined the minimum time required to estimate long-term population trends (Fig. \ref{fig:gam_example}). I found that although I obtain a similar distribution of minimum times required for GAMs, the minimum time required for GAMs is on average shorter than for linear regression (Fig. \ref{fig:min_time_dist_gam}).

# Discussion

I explored two approaches to estimate the minimum time series length required to address a particular question of interest. I asked, what is the minimum time series length required to determine long-term population trends using linear regression? This is one of the simplest questions one could ask of a time series. The simulation-based approach has been suggested by others, especially in situations more complicated than that suited for classic power analysis [@Gerrodette1987;@Johnson2015;@Bolker2008]. My simulations support past work that longer time series are needed when the trend strength (i.e. rate of increase or decrease) is weak or when population variability is high [@Gerrodette1987]. I also showed how the simulation model can be altered for a particular population (Fig. \ref{fig:shark_example}) or question (Figs. \ref{fig:min_time_dist_log_pop},\ref{fig:gam_example}).

Here, I focus on an empirical approach to estimate the minimum time series length required to assess changes in abundance over time. I examined `r nrow(pop_info)` population time series (all longer than 35 years). I then subsampled each to determine the minimum time required to achieve a desired significance level and power for linear regression. Statistical power is important as it provides on information as to the necessary samples required to determine a significant trend [@Legg2006]. I found that on average `r round(mean(pop_info$min_time_for_power),2)` years of continuous monitoring were typically necessary (Fig. \ref{fig:min_time_dist}b). However, the distribution of minimum time required was wide. This time-frame is in line with past work on a short-lived snail species [@Rueda-Cediel2015] and a long-lived whale species [@Gerber1999]. @Hatch2003 used seabird monitoring data to estimate minimum sampling requirements. He found that the time required ranged from 11 to 69 years depending on species, trend strength, and study design.

In line with theoretical predictions [@Rhodes2011], I also found $T_{min}$ was strongly correlated with the trend strength, variability in population size, and temporal autocorrelation (Fig. \ref{fig:correlates}). Contrary to my prior expectations, I also found that $T_{min}$ did not correlate with any life-history traits (Fig. \ref{fig:biological_correlates}). I initially hypothesized that species with longer lifespans or generation times may require a longer sampling period. This result could have been a result of at least two factors. First, the data I used may not include a diverse enough set of species with different life-history traits. Second, the question I poised, whether a population is increasing or decreasing, was specifically concerned with trends in population density over time. Therefore, life-history characteristics may be more important for other questions, like estimating species extinction risk [@Hutchings2012]. For example, @Blanchard2007 used detailed simulations of spatially-distributed fisheries to compare survey designs. They found that statistical power depended on survey design, temperature preferences, and the degree of population patchiness.

<!-- Similarly, @Bart2004 recommended a standard of 80% power to detect a 50% decline over 20 years in landbird populations. They found that only about 42% of populations in the Breeding Bird Survey met their criteria. @Seavy2007 asked whether statistical power was even a useful framework for assessing long-term population trends. They used 24 years of census data on Red-tailed Tropicbirds (\emph{Phaethon rubricauda}) in Hawaii and showed that to detect a 50% decline over 10 years almost always resulting in high statistical power (above 0.8). Therefore, they cautioned against only using power analyses to design monitoring schemes and instead argued for metrics that would increase precision. For example, @Seavy2007 suggest improving randomization, reducing bias, and increases detection probability when designing and evaluating monitoring programs.-->

An important related question, is the optimal allocation of sampling effort in space versus time. In a theoretical investigation of this question, @Rhodes2011 found that the optimal allocation of sampling depended strongly on temporal and spatial autocorrelation. If spatial population dynamics were highly correlated, then it was better to sample more temporally, and vice versa. My work supports this idea as populations with strong temporal autocorrelation needed less years of sampling (Fig. \ref{fig:correlates}). @Morrison2008 also studied the optimal allocation of sampling effort in space versus time, but used emprical data from a long-term survey of the collared pika (\emph{Ochotona collaris}) found in the Yukon. They estimated long-term growth rates among three subpopulations over a 10-year period. They found that surveys less than 5 years may be misleading and that extrapolating from one population to another, even when nearby geographically, may be untenable.

@Seavy2007 asked whether statistical power was even a useful framework for assessing long-term population trends. They used 24 years of census data on Red-tailed Tropicbirds (\emph{Phaethon rubricauda}) in Hawaii and showed that to detect a 50% decline over 10 years almost always resulting in high statistical power (above 0.8). Therefore, they cautioned against only using power analyses to design monitoring schemes and instead argued for metrics that would increase precision. For example, @Seavy2007 suggest improving randomization, reducing bias, and increases detection probability when designing and evaluating monitoring programs. I agree that power analyses should not be the only consideration when designing monitoring schemes. However, unlike @Seavy2007, my results indicate that longer than 10 years is often needed to achieve high statistical power.

This paper also has practical implications for the IUCN Redlist criteria. IUCN criteria A2 suggests that species that have experienced 30% declines over 10 years (or three generations) should be listed as vulnerable [@IUCN2012]. However, for the populations I examined, this criteria may be too simplistic (Fig. \ref{fig:IUCN_analysis}). For many populations, the IUCN criteria suggest more years than necessary are required to assess a population as vulnerable (points below diagonal line in Fig. \ref{fig:IUCN_analysis}). Conversely, for other populations the IUCN criteria suggest sampling times that are less than the minimum time required for statistical power. This suggests that the IUCN criteria are probably too simplistic as the minimum time required does not correlate with generation time (Fig. \ref{fig:biological_correlates}).

The design of monitoring programs should include calculations of statistical power, the allocation of sampling in space versus time [@Rhodes2011], and metrics to increase precision. Ideally, a formal decision analysis to evaluate these different factors would be conducted to design or assess any monitoring program [@Hauser2006; @McDonald-Madden2010]. This type of formal decision analysis would also include information on the costs of monitoring. These costs include the actual costs of sampling [@Brashares2005] and the ecological costs on inaction [@Thompson2000].

<!--
@Rueda-Cediel2015 found that 10-15 years was adequate to characterize population trends in a short lived animal. They used a matrix model of a particular species...

@Gerber1999 investigated the minimum time series required to estimate population growth of an endagered, but recovering, species. They used a long-term census to retroactively determine the minimum time series required to assess threat status. They found that only XX years were needed, XX years before the delisting decision was actually made. This highlights the importance of estimating the minimunm time series required as an earlier decision would have saved time and money [@Gerber1999].
-->

## Limitations

This paper has some limitations in determining the minimum time series length required. First, $T_{min}$ is particular to the specific question of interest. An additional complication is that for the empirical approach, the subsampling of the full time series allows for estimates of power, but the individual subsamples are clearly not independent of one another. Further, estimates of $T_{min}$ depend on chosen values of $\alpha$ and $\beta$ (Fig. \ref{fig:min_time_vs_alpha_beta}). In an ideal setting, a specific population model would be parameterized for each population of interest [@McCain2016]. Then, model simulations could be used to estimate the minimum time series required to address each specific question of interest. Clearly, this is not always practical, especially if conducting analyses for a wide array of species as I do here. In addition, the statistical models suggest that $T_{min}$ does not correlate with any life-history traits, at least for the question of linear regression (Fig. \ref{fig:biological_correlates}). Therefore, it is not possible to use these results to predict $T_{min}$ for another population, even if the population is of a species with a similar life-history to one in the database used.

## Conclusions

I used a database of `r nrow(pop_info)` populations to determine the minimum time series length required to detect population trends. This goes beyond previous work that either focused on theoretical investigations or a limited number of species. I show that to identify long-term changes in abundance, on average `r round(mean(pop_info$min_time_for_power),2)` years of continuous monitoring are often required (Fig. \ref{fig:min_time_dist}). However, there is wide distribution of estimated minimum times. Therefore, it is probably not wise to use a simple threshold number of years in monitoring design.

In line with theoretical predictions [@Gerrodette1987], I also show that $T_{min}$ is strongly correlated with the long-term population trend (i.e. rate of increase), variability in population size, and the temporal autocorrelation (Fig. \ref{fig:correlates}). Contrary to my initial hypotheses, minimum time required did not correlate with generation time or any other life-history traits (Fig. \ref{fig:biological_correlates}). This result argues against overly simplified measures of minimum sampling time based on generation length (Fig. \ref{fig:IUCN_analysis}).

My work implies that for many populations, short time series are probably not reliable for detecting population trends. This result highlights the importance of long-term monitoring programs. From both a scientific and management perspective estimates of $T_{min}$ are important. If a time series is too short, we lack the statistical power to reliably detect long-term population trends. In addition, a time series that is too long may be a poor use of already limited funds [@Gerber1999]. Further, more data is not always best in situations where management actions need to be taken [@Martin2012; @Martin2017]. When a population trend is detected, it may be too late for management action. In these situations, the precautionary principle may be more appropriate [@Thompson2000]. Future work should examine other species, with a wider range of life-history characteristics. In addition, similar approaches can be used to determine the minimum time series length required to address additional questions of interest.

<!--
## Time series length recommendations from other papers

- greater than 5 years [@Morrison2008]
- more than 10-15 years [@Rueda-Cediel2015]
- general advice (variability, correlations) [@Rhodes2011]
- more than 10 years [@Gerber1999]
- specific recommendations depend on amount of power and strength of trend (Bart 2004, Hatch 2003) [@Seavy2007]
- paper on syncrony from James
-->

# Supporting Information

In the supporting material, I provide an expanded methods sections, additional figures, minimum time calculations for determining exponential growth, simulations with a more complicated population model, and the use of generalized additive models to identify population trends. All code and data can be found at https://github.com/erwhite1/time-series-project

# Acknowledgements

ERW was partially supported by a National Science Foundation Graduate Fellowship. I would like to thank members of the Ecological Theory group at the University of California, Davis for their insight. I would also like to thank T. Dallas and E. Malcolm-White <!--X anonymous reviewers -->for their helpful comments.

# References

<div id="refs"></div>

\clearpage

# Supplementary material

 \setcounter{figure}{0}
 \renewcommand{\thefigure}{A\arabic{figure}}

\begin{figure}[htbp]
\centering
\caption{(a) Population size of Bigeye tuna (\emph{Thunnus obesus}) over
time. The line is the best fit line from linear regression. (b)
Statistical power for different subsets of the time series in panel
a.\label{fig:empirical_approach_example}}
\end{figure}

\begin{figure}[htbp]
\centering
\caption{Output of generalized linear model with a Poisson error
structure for predicting the minimum time required with explanatory
variables of the absolute value of the slope coefficient (or trend
strength), temporal autocorrelation, and variability in population
size.\label{fig:poisson_model}}
\end{figure}

\begin{figure}[htbp]
\centering
\caption{(a) Minimum time required to estimate change in abundance by biological class, (b) long-term trend (estimated slope coefficient) by class, (c) coefficient of variation in population size by class, and (d) temporal autocorrelation by class.\label{fig:class}}
\end{figure}

\begin{figure}[htbp]
\centering
\caption{Minimum time required to assess long-term trends in abundance
for values of statistical significance (\(\alpha\)) and power
(\(1-\beta\)).\label{fig:min_time_vs_alpha_beta}}
\end{figure}

\begin{figure}[htbp]
\centering
\caption{The difference between minimum time estimates is the minimum time required to achieve 0.8 statistical power versus the minimum time required under IUCN criteria A2 to classify a species as vulnerable. Each point represents a single population, all of which saw declines of 30\% or greater over a 10 year period. (a) Difference between minimum time estimates versus the coefficient of variation in population size. (b) Difference between minimum time estimates versus the generation length in years.\label{fig:IUCN_correlates}}
\end{figure}

\begin{figure}[htbp]
\centering
\caption{Distribution of the minimum time required in order to detect a
significant trend (at the 0.05 level) in log(abundance) given power of
0.8.\label{fig:min_time_dist_log_pop}}
\end{figure}

\begin{figure}[htbp]
\centering
\caption{Statistical power for different length of time series
simulations for a lemon shark population in Bimini,
Bahamas.\label{fig:shark_example}}
\end{figure}

<!--
\begin{figure}[htbp]
\centering
\caption{Example of calculating minimum time required for growth rate
estimation. (a) European herring gull (\emph{Larus argentatus}) scaled
density over time, (b) mean and standard deviation of growth rate for
subsamples of entire time series, and (c) the percent error between mean
estimated growth rate and the true long-term growth rate. The vertical
bar denotes the minimum time required to estimate growth rate within
20\% error.\label{fig:growth_rate}}
\end{figure}

\begin{figure}[htbp]
\centering
\caption{Histogram of the minimum time required in order to estimate the
long-term growth rate within 20\%
error.\label{fig:min_time_growth_dist}}
\end{figure}
-->

\begin{figure}[htbp]
\centering
\caption{(a) Time series for Bigeye tuna (\emph{Thunnus obesus}) with
corresponding fitted GAM model in red and (b) statistical power as a
function of the number of years sampled. The horizontal line at 0.8
indicates the minimum threshold for statistical power and the vertical
line denotes the minimum time required to achieve 0.8 statistical
power.\label{fig:gam_example}}
\end{figure}

\begin{figure}[htbp]
\centering
\caption{Distribution of the minimum time required in order to detect a
significant trend (at the 0.05 level) in abundance according to a GAM
model given statistical power of 0.8. The smoothing parameter was set to
3 for each population.\label{fig:min_time_dist_gam}}
\end{figure}

\vspace{1 cm}

Table A1: Output of generalized linear model to examine time series characteristics as correlates of the minimum time required for determining long-term population trends.

\vspace{1 cm}

Table A2: Output of generalized linear model to examine life-history trait correlates of the minimum time required for determine long-term population trends.

\vspace{1 cm}

Table A3: Output of generalized linear model to examine both time series characteristics and life-history trait correlates of the minimum time required for determine long-term population trends.