### 2018-02-19 writing/manuscript_draft.tex

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{setspace}  
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{}
  \pretitle{\vspace{\droptitle}}
  \posttitle{}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

\usepackage{float} \usepackage{lineno}
\usepackage[firstinits=true,style=authoryear-comp,natbib=true,doi=false,isbn=false,url=false,uniquename=false,uniquelist=false,sorting=none]{biblatex}

\begin{document}

Title: Minimum time required to detect population trends: the need for
long-term monitoring programs \vspace{7 mm}

Author: \textsc{Easton R. White$^{1,2}$} \vspace{3 mm}

Address:\emph{
    \\$^1$Center for Population Biology \\
    University of California, Davis \\
    2320 Storer Hall \\
        University of California, Davis \\
        One Shields Avenue Davis, CA, USA} \vspace{3 mm}

\(^2\)Corresponding author:
\href{mailto:eawhite@ucdavis.edu}{\nolinkurl{eawhite@ucdavis.edu}} , 1
480 203 7931 \vspace{3 mm}

Number of words (including 300 words per figure): 6,950 \vspace{1 mm}

Number of figures: 5 \vspace{1 mm}

Number of references: 34 \vspace{1 mm}

Supplementary material: 9 figures and 3 tables \vspace{3 mm}

Data and code for all the figures and tables can be found at
(\url{https://github.com/erwhite1/time-series-project}).

To be submitted to: \emph{Biological Conservation} \vspace{3 mm}

Available as a preprint at: \url{https://peerj.com/preprints/3168/}
\vspace{3 mm}

Color does not need to be used for any figures in print \vspace{3 mm}

\vspace{3 mm}

\begin{spacing}{2}

\linenumbers

\begin{abstract}
Long-term time series are necessary to better understand population dynamics, assess species' conservation status, and make management decisions. However, population data are often expensive, requiring a lot of time and resources. When is a population time series long enough to address a question of interest? I determine the minimum time series length required to detect significant increases or decreases in population abundance. To address this question, I use simulation methods and examine 822 populations of vertebrate species. Here I show that on average 15.9 years of continuous monitoring are required in order to achieve a high level of statistical power. However, there is a wide distribution around this average, casting doubt on simple rules of thumb. For both simulations and the time series data, the minimum time required depends on trend strength, population variability, and temporal autocorrelation. However, there were no life-history traits (e.g. generation length) that were predictive of the minimum time required. These results point to the importance of sampling populations over long periods of time. I argue that statistical power needs to be considered in monitoring program design and evaluation. Short time series are likely under-powered and potentially misleading.
\end{abstract}

Keywords: ecological time series, experimental design, monitoring, power
analysis, statistical power, sampling design

\section{Introduction}\label{introduction}

Observational studies and population time series have become a
cornerstone of modern ecological research and conservation biology
(Magurran et al. 2010; Hughes et al. 2017). Long-term data are necessary
to both understand population dynamics and to assess species extinction
risk. Even though many time series may now be considered ``long-term''
(e.g.~continuous plankton recorder, Giron-Nava et al. (2017)), most are
still short. Time series are typically short for a variety of reasons
(Field et al. 2007). They are often coupled with an experiment, which
may only last a couple of years. In addition, short funding cycles make
it difficult to examine populations over longer periods of time (Hughes
et al. 2017).

How long of a time series is actually necessary? This question has
important implications for both research and management (Nichols and
Williams 2006). Scientists need to know the time series length required
to address a specific question. A short time series may lead to wrong
conclusions given large natural year-to-year variability (McCain,
Szewczyk, and Knight 2016). Managers need to know when action is needed
for a population. Therefore, managers must understand when population
trend over time is actually meaningful. For example, the International
Union for Conservation of Nature (IUCN) Red List Categories and Criteria
suggest, under Criterion A2, a species qualifies as vulnerable if it has
experienced a 30\% decline over 10 years, or three generations (IUCN
2012). For both scientific and management questions, because sampling is
typically expensive, we also do not want to sample for longer than is
necessary. For example, Gerber, DeMaster, and Kareiva (1999)
investigated the minimum time series required to estimate population
growth of the endangered, but recovering, eastern North Pacific gray
whale (\emph{Eschrichtius robustus}). They used a long-term census to
retroactively determine the minimum time series required to assess
threat status. They found that only 11 years were needed, eight years
before the delisting decision was made. This highlights the importance
of estimating the minimum time series required as an earlier decision
would have saved time and money (Gerber, DeMaster, and Kareiva 1999).
Further, waiting too long to make a decision can imperil a species where
management action could have been taken earlier (Martin et al. 2012;
Martin et al. 2017).

An important step in experimental design is to determine the number of
samples required. For any particular experiment four quantities are
intricately linked: significance level (\(\alpha\)), statistical power,
effect size, and sample size (Legg and Nagy 2006). The exact
relationship between these quantities depends on the specific
statistical test. A type I error is a false positive, or incorrect
rejection of a true null hypothesis. For example, if a time series was
assessed as significantly increasing or decreasing---when there was no
true significant trend---this would be a false positive. The false
positive rate, or significance level (\(\alpha\)) is often set at 0.05
(although this is purely historical, Mapstone (1995)). A type II error
(\(\beta\)) is a failure to detect a true trend, or failure to reject a
false null hypothesis. Formally, statistical power (\(1-\beta\)) is one
minus the probability of a type II error (\(\beta\)). The effect size is
a measure of the difference between two groups. Prior to an experiment,
one could set appropriate levels of power, significance level, and the
effect size to estimate the sample size required for the experiment.
This approach, however, is not straight-forward for a time series, or
more complicated scenarios (P. C. Johnson et al. 2015), as data are
clearly non-independent.

For time series data, two general approaches to estimating sample size
are appropriate. Simulations can be designed for a specific population
and question (Bolker 2008; P. C. Johnson et al. 2015). Simple models can
be simulated with parameter values corresponding to a population of
interest (Gerrodette 1987). Statistical power is the proportion of
simulations that meet some set of criteria. The specific criteria depend
on the question at hand. For example, given a time series, when is the
slope from linear regression significantly different from zero? In other
words, when is the time series significantly increasing or decreasing?
It is then possible to determine how power changes with a variable of
interest. For example, time series can be simulated for different
lengths of time. From these simulations, the minimum time series length
required to meet certain levels of statistical significance and power is
estimated (Bolker 2008).

In addition to using simulations, empirical time series can also be
used. Multiple replicates of similar populations are usually not
available, but it is possible to subsample an empirical time series
(Gerber, DeMaster, and Kareiva 1999; Brashares and Sam 2005). Subsamples
of different lengths can then be evaluated to estimate the proportion of
subsamples meeting some criteria, again a measure of statistical power.
Similar to the simulation approach, this measures of power can be used
to determine the minimum time series required for a particular question
of interest.

Past work has investigated questions related to the minimum time series
required to estimate trends in population size over time (Wagner,
Vandergoot, and Tyson 2009; Giron-Nava et al. 2017). For example, Rhodes
and Jonzen (2011) examined the optimal allocation of effort between
spatial and temporal replicates. Using simple populations models, they
found that the allocation of effort depends on environmental variation,
spatial and temporal autocorrelation, and observer error. Rueda-Cediel
et al. (2015) also used a modeling approach, but parameterized a model
specific for a threatened snail, \emph{Tasmaphena lamproides}. They
found that for this short-lived organism, 15 years was adequate to
assess long-term trends in abundance. However, these studies, and other
past work, have typically focused either on theoretical aspects of
monitoring design or focused on only a few species.

I use both simulations and empirical time series to determine the
minimum number of years required to address several questions. I
estimate the minimum time series length required (\(T_{min}\)) to assess
long-term changes in abundance via simple linear regression. First, I
estimate \(T_{min}\) using a simulation approach. Then I examine 822
population time series to estimate \(T_{min}\). In the supplementary
material, I determine \(T_{min}\) for related ideas: using more
complicated population models, varying statistical level and power, and
the use of generalized additive models.

\section{Methods}\label{methods}

\subsection{Simulation approach}\label{simulation-approach}

One approach to determining the minimum time series length needed is
through repetitive simulations of a population model (Gerrodette 1987).
This is the same approach one might use in sample size calculations for
any experimental design too complicated for simple power analyses
(Bolker 2008; P. C. Johnson et al. 2015). I only briefly discuss this
approach as it has been described elsewhere. Essentially, a population
model is simulated repetitively for a number of years. This approach
requires us to determine values for model parameters (e.g.~population
variability). As an example, we can take the following population model
for population size \(N\) at time \(t\):

\begin{equation}
N(t + 1) = N(t) + r(t) + \epsilon \mbox{ with } \epsilon \sim N(\mu, \sigma)
\label{sim_model}
\end{equation}

where \(\epsilon\) is a normally-distributed random noise term with mean
\(\mu\) and standard deviation \(\sigma\). The rate of growth \(r\) is
also the trend strength of the increase or decrease (i.e.~the rate of
increase). It is important to note that any population model could be
substituted for equation \ref{sim_model}, as in the supplementary
material (Figs. \ref{fig:min_time_dist_log_pop},
\ref{fig:shark_example}).

Statistical power is then the proportion of simulations that meet some
criteria. Here, our criteria is whether the slope parameter from linear
regression is significant at the 0.05 threshold with statistical power
of 0.8. Statistical power of 0.8 would indicate that, if there was a
true trend in abundance, there would be a 0.8 probability of detecting
the trend. Values of 0.05 for the significance level and 0.8 statistical
power are historical and it is important to examine the effect of
changing these values (Fig. \ref{fig:min_time_vs_alpha_beta}).

In Fig. \ref{fig:theoretical_approach}a, a number of simulated time
series are shown for a set number of time periods (\(t=40\)). It is
clear that statistical power increases quickly with increases in length
of time sampled (Fig. \ref{fig:theoretical_approach}b). Where power is
greater than 0.8 (the dotted line), that is the minimum time required
(\(T_{min}\)) to be confident in the detection of a long-term trend in
abundance.

\begin{figure}[htbp]
\centering
\includegraphics{manuscript_draft_files/figure-latex/unnamed-chunk-1-1.pdf}
\caption{(a) Example of a simulated time series for 40 time periods. (b)
Statistical power versus the simulated time series length. The
horizontal, dashed line is the desired statistical power of 0.8. The
vertical, dashed line is the minimum time required to achieve the
desired statistical power. (c) Minimum time required (\(T_{min}\)) for
simulations with different values of the trend strength (\(r\)). (d)
Minimum time required for different levels of population variability
(\(\sigma\)). In each case, the minimum time required is the minimum
number of years to achieve 0.8 statistical power given a significance
level of 0.05.\label{fig:theoretical_approach}}
\end{figure}

\subsection{Data source}\label{data-source}

I use a database of 2444 population time series compiled in (Keith et
al. 2015); they compared the predictability of growth rates among
populations. The data are originally from the Global Population Dynamics
Database (NERC Centre for Population Biology 2010) and several other
sources (Keith et al. 2015). I filtered out short time series (less than
35 years), and those with missing data, leaving 822 time series. The
data includes information on 477 vertebrate species with a focus on
mammals, birds, and fish. The data also includes information on
generation length and survey specifications. For each time series, I
also calculate other variables of interest: coefficient of variance in
population size, long-term trend in abundance (slope coefficient from
simple linear regression), and temporal autocorrelation. All analyses
were conducted in R (R Core Team 2016).

For a subset of populations (\(n =\) 547), there is information on
life-history traits available from another paper (Myhrvold et al. 2015),
including body size and generation length. All 547 populations were
birds. I examine how the minimum time required is related to these
life-history traits (Fig. \ref{fig:biological_correlates}).

\subsection{Empirical approach}\label{empirical-approach}

I assume that each time series is long enough to include all necessary
information (e.g.~variability) about the population. In other words,
each time series is a representative sample. I first take all possible
contiguous subsamples of each time series. For example, a time series of
35 years would have 34 possible contiguous subsamples of length 2, 34
possible contiguous subsamples of length 3, and continuing until 1
possible contiguous subsample of length 35 (Gerber, DeMaster, and
Kareiva 1999; Giron-Nava et al. 2017). Next, I run a linear regression
for each subsampled time series. Then, I determine the proportion of
subsamples of a particular length that have estimated slope coefficients
which are statistically different from zero. I only look at the
proportion of samples where the long-term, or ``true'', time series also
has a significant slope. This proportion is a measure of statistical
power. Lastly, I determine which subsample length is required to achieve
a certain threshold of statistical power (0.8, Cohen (1992)). The
minimum subsampled length that met these criteria is the minimum time
series length required (\(T_{min}\)).

In the supplementary material, I show how the same approach described
here for more complicated population models. I also determine the
minimum time required to estimate long-term trends according to
generalized additive models, instead of the simple linear models used
here (Fig. \ref{fig:gam_example}).

\section{Results}\label{results}

I determined the minimum time series length (\(T_{min}\)) required to
address a particular question of interest. What is the minimum time
series length required to determine, via linear regression, the
long-term population trend? Here, the minimum time series length
required had high enough statistical power (greater than 0.8) for a set
significance level (\(\alpha\)) of 0.05. It is also possible to alter
statistical power and \(\alpha\). Predictably, with increased
statistical power or decreased \(\alpha\), \(T_{min}\) increased (Fig.
\ref{fig:min_time_vs_alpha_beta}). I then estimated \(T_{min}\) using
two approaches. I briefly describe results from the simulation approach
and then discuss the empirical approach.

\subsection{Simulation approach}\label{simulation-approach-1}

I constructed a general population model where the trend strength
(i.e.~slope coefficient) over time could be a model parameter. I then
simulated time series of different lengths. From these simulations I
determined the minimum time series length required to achieve a certain
level of statistical power. In line with past work (Gerrodette 1987), I
found the \(T_{min}\) increases (i.e.~more time is required) with
decreases in trend strength and with increases in population variability
(Figs. \ref{fig:theoretical_approach}c,d).

I chose a simple model, but any other population model could be used
(see example in Fig. \ref{fig:min_time_dist_log_pop}). Ideally, the
specific model choice should be tailored to the population of interest.
I explored how the simulation approach could be applied to more
biologically-realistic population models (Fig. \ref{fig:shark_example}).
Specifically, I determined the minimum time required to estimate
long-term population trends using a stochastic, age-structured model of
lemon shark population dynamics in the Bahamas (White, Nagy, and Gruber
2014). I found that over 27 years of continuous monitoring were needed
in this particular scenario (Fig. \ref{fig:shark_example}). Similar to
the simulation approach described above, the minimum time required for
the lemon shark population was strongly dependent on model parameters.

\subsection{Empirical approach}\label{empirical-approach-1}

\begin{figure}[htbp]
\centering
\includegraphics{manuscript_draft_files/figure-latex/unnamed-chunk-4-1.pdf}
\caption{(a) Distributions of the minimum time required for populations
from four different biological classes. (b) Distribution of minimum time
required for all populations regardless of biological class. The minimum
time required calculation corresponds to a significance level of 0.05
and statistical power of 0.8.\label{fig:min_time_dist}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{manuscript_draft_files/figure-latex/unnamed-chunk-5-1.pdf}
\caption{Minimum time required to estimate change in abundance
correlated with (a) trend strength (absolute value of slope coefficient
estimated from linear regression), (b) coefficient of variation in
interannual population size, and (c) temporal lag-1
autocorrelation.\label{fig:correlates}}
\end{figure}

I examined a database of 822 separate population time series
representing 477 species. This database consisted of vertebrate species
with a variety of life-history characteristics (Fig.
\ref{fig:biological_correlates}). I limited analyses to populations with
at least 35 years of continuous sampling. I then examined the minimum
time required to estimate long-term trends via linear regression.

Across all the populations, I found an average minimum time series
length required (\(T_{min}\)) of 15.9 (SD=8.3), with a wide distribution
(Fig. \ref{fig:min_time_dist}b). Estimates of \(T_{min}\) varied between
biological class (Fig. \ref{fig:min_time_dist}a). Ray-finned fish (class
Actinopterygii) typically had estimates of \(T_{min}\) over 20 years.
Birds (class Aves) had a much wider distribution of \(T_{min}\), but
usually required less years of sampling. Differences between these
classes can be explained by differences in variability in population
size and strength of trends in abundance (Fig. \ref{fig:class}).

\subsubsection{Corrrelates for minimum time
required}\label{corrrelates-for-minimum-time-required}

The minimum time series length required was strongly correlated with
trend strength (i.e.~estimated slope coefficient from linear
regression), coefficient of variation in population size, and
autocorrelation in population size (Fig. \ref{fig:correlates}). This is
in line based on simulations here and those of others (Rhodes and Jonzen
2011). Using a generalized linear model, with a Poisson error structure,
all three of these explanatory variables were significant and had large
effect sizes (see Table A1). Combined, trend strength, coefficient of
variation in population, and autocorrelation account for 75.1\% of the
explained deviance (Zuur et al. 2009) in minimum time series length
required.

For a subset of the populations I combined time series data with a data
on life-history characteristics of amniotes (Myhrvold et al. 2015).
There was life-history information available for 547 populations
representing 315 different species, all of which were birds (Aves
class).

\begin{figure}[htbp]
\centering
\includegraphics{manuscript_draft_files/figure-latex/unnamed-chunk-7-1.pdf}
\caption{Mimimum time required versus (a) generation length (years), (b)
litter size (n), (c) log adult body mass (grams), (d) maximum longevity
(years), and (e) incubation (days). The lines in each plot represent the
best fit line from linear regression.\label{fig:biological_correlates}}
\end{figure}

Some life-history traits were significant predictors for the minimum
time required (Fig. \ref{fig:biological_correlates}, Tables A2,A3).
However, none of these life-history traits explained a large part of the
variation in minimum time required. In a generalized linear model, all
of the life-history traits described in figure
\ref{fig:biological_correlates} account for only 5.99\% of the explained
deviance in minimum time series length required. In addition, when
accounting for trend strength, coefficient of variation, and
autocorrelation, no life-history traits were significant predictors of
the minimum time required (Table A3).

\subsubsection{Evalulating the IUCN
criteria}\label{evalulating-the-iucn-criteria}

\begin{figure}[htbp]
\centering
\includegraphics{manuscript_draft_files/figure-latex/unnamed-chunk-8-1.pdf}
\caption{Minimum time required to achieve 0.8 statistical power versus
the minimum time required under IUCN criteria A2 to classify a species
as vulnerable. Each point represents a single population, all of which
saw declines of 30\% or greater over a 10 year
period.\label{fig:IUCN_analysis}}
\end{figure}

The IUCN Red List Categories and Criteria suggest, under Criterion A2, a
species qualifies as vulnerable if it has experienced a 30\% decline
over 10 years, or 3 generations (whichever is longer) (IUCN 2012). I
examined a subset of populations with observed declines of 30\% or
greater over 10 years, qualifying all of them as vulnerable. This
resulted in n = 162 populations of fish, birds, and a single mammal. I
then compared the minimum time required to achieve 0.8 statistical power
(\(T_{min}\)) to the minimum time required under the IUCN criteria (Fig.
\ref{fig:IUCN_analysis}). For populations below the identity line in
figure \ref{fig:IUCN_analysis}, IUCN criteria would require more
sampling compared to estimates for \(T_{min}\). Further, populations
above the identity line are cases where the IUCN criteria would classify
a population as vulnerable despite not having sampled enough years to
achieve high statistical power (Fig. \ref{fig:IUCN_analysis}). The
silhouettes on figure \ref{fig:IUCN_analysis} highlight that species
with longer generation times typically have larger discrepancies between
\(T_{min}\) and the minimum time required for IUCN assessments (Fig.
\ref{fig:IUCN_correlates}).

\subsubsection{Sensitivity analysis}\label{sensitivity-analysis}

Lastly, I tested model sensitivity by using generalized additive models
(GAMs) instead of simple linear regression. Again, I examined the
minimum time required to estimate long-term population trends (Fig.
\ref{fig:gam_example}). I found that although I obtain a similar
distribution of minimum times required for GAMs, the minimum time
required for GAMs is on average shorter than for linear regression (Fig.
\ref{fig:min_time_dist_gam}).

\section{Discussion}\label{discussion}

I explored two approaches to estimate the minimum time series length
required to address a particular question of interest. I asked, what is
the minimum time series length required to determine long-term
population trends using linear regression? This is one of the simplest
questions one could ask of a time series. The simulation-based approach
has been suggested by others, especially in situations more complicated
than that suited for classic power analysis (Gerrodette 1987; P. C.
Johnson et al. 2015; Bolker 2008). My simulations support past work that
longer time series are needed when the trend strength (i.e.~rate of
increase or decrease) is weak or when population variability is high
(Gerrodette 1987). I also showed how the simulation model can be altered
for a particular population (Fig. \ref{fig:shark_example}) or question
(Figs. \ref{fig:min_time_dist_log_pop},\ref{fig:gam_example}).

Here, I focus on an empirical approach to estimate the minimum time
series length required to assess changes in abundance over time. I
examined 822 population time series (all longer than 35 years). I then
subsampled each to determine the minimum time required to achieve a
desired significance level and power for linear regression. Statistical
power is important as it provides on information as to the necessary
samples required to determine a significant trend (Legg and Nagy 2006).
I found that on average 15.91 years of continuous monitoring were
typically necessary (Fig. \ref{fig:min_time_dist}b). However, the
distribution of minimum time required was wide. This time-frame is in
line with past work on a short-lived snail species (Rueda-Cediel et al.
2015) and a long-lived whale species (Gerber, DeMaster, and Kareiva
1999). Hatch (2003) used seabird monitoring data to estimate minimum
sampling requirements. He found that the time required ranged from 11 to
69 years depending on species, trend strength, and study design.

In line with theoretical predictions (Rhodes and Jonzen 2011), I also
found \(T_{min}\) was strongly correlated with the trend strength,
variability in population size, and temporal autocorrelation (Fig.
\ref{fig:correlates}). Contrary to my prior expectations, I also found
that \(T_{min}\) did not correlate with any life-history traits (Fig.
\ref{fig:biological_correlates}). I initially hypothesized that species
with longer lifespans or generation times may require a longer sampling
period. This result could have been a result of at least two factors.
First, the data I used may not include a diverse enough set of species
with different life-history traits. Second, the question I poised,
whether a population is increasing or decreasing, was specifically
concerned with trends in population density over time. Therefore,
life-history characteristics may be more important for other questions,
like estimating species extinction risk (J. A. Hutchings et al. 2012).
For example, Blanchard, Maxwell, and Jennings (2007) used detailed
simulations of spatially-distributed fisheries to compare survey
designs. They found that statistical power depended on survey design,
temperature preferences, and the degree of population patchiness.

An important related question, is the optimal allocation of sampling
effort in space versus time. In a theoretical investigation of this
question, Rhodes and Jonzen (2011) found that the optimal allocation of
sampling depended strongly on temporal and spatial autocorrelation. If
spatial population dynamics were highly correlated, then it was better
to sample more temporally, and vice versa. My work supports this idea as
populations with strong temporal autocorrelation needed less years of
sampling (Fig. \ref{fig:correlates}). Morrison and Hik (2008) also
studied the optimal allocation of sampling effort in space versus time,
but used emprical data from a long-term survey of the collared pika
(\emph{Ochotona collaris}) found in the Yukon. They estimated long-term
growth rates among three subpopulations over a 10-year period. They
found that surveys less than 5 years may be misleading and that
extrapolating from one population to another, even when nearby
geographically, may be untenable.

Seavy and Reynolds (2007) asked whether statistical power was even a
useful framework for assessing long-term population trends. They used 24
years of census data on Red-tailed Tropicbirds
(\emph{Phaethon rubricauda}) in Hawaii and showed that to detect a 50\%
decline over 10 years almost always resulting in high statistical power
(above 0.8). Therefore, they cautioned against only using power analyses
to design monitoring schemes and instead argued for metrics that would
increase precision. For example, Seavy and Reynolds (2007) suggest
improving randomization, reducing bias, and increases detection
probability when designing and evaluating monitoring programs. I agree
that power analyses should not be the only consideration when designing
monitoring schemes. However, unlike Seavy and Reynolds (2007), my
results indicate that longer than 10 years is often needed to achieve
high statistical power.

This paper also has practical implications for the IUCN Redlist
criteria. IUCN criteria A2 suggests that species that have experienced
30\% declines over 10 years (or three generations) should be listed as
vulnerable (IUCN 2012). However, for the populations I examined, this
criteria may be too simplistic (Fig. \ref{fig:IUCN_analysis}). For many
populations, the IUCN criteria suggest more years than necessary are
required to assess a population as vulnerable (points below diagonal
line in Fig. \ref{fig:IUCN_analysis}). Conversely, for other populations
the IUCN criteria suggest sampling times that are less than the minimum
time required for statistical power. This suggests that the IUCN
criteria are probably too simplistic as the minimum time required does
not correlate with generation time (Fig.
\ref{fig:biological_correlates}).

The design of monitoring programs should include calculations of
statistical power, the allocation of sampling in space versus time
(Rhodes and Jonzen 2011), and metrics to increase precision. Ideally, a
formal decision analysis to evaluate these different factors would be
conducted to design or assess any monitoring program (Hauser, Pople, and
Possingham 2006; McDonald-Madden et al. 2010). This type of formal
decision analysis would also include information on the costs of
monitoring. These costs include the actual costs of sampling (Brashares
and Sam 2005) and the ecological costs on inaction (Thompson et al.
2000).

\subsection{Limitations}\label{limitations}

This paper has some limitations in determining the minimum time series
length required. First, \(T_{min}\) is particular to the specific
question of interest. An additional complication is that for the
empirical approach, the subsampling of the full time series allows for
estimates of power, but the individual subsamples are clearly not
independent of one another. Further, estimates of \(T_{min}\) depend on
chosen values of \(\alpha\) and \(\beta\) (Fig.
\ref{fig:min_time_vs_alpha_beta}). In an ideal setting, a specific
population model would be parameterized for each population of interest
(McCain, Szewczyk, and Knight 2016). Then, model simulations could be
used to estimate the minimum time series required to address each
specific question of interest. Clearly, this is not always practical,
especially if conducting analyses for a wide array of species as I do
here. In addition, the statistical models suggest that \(T_{min}\) does
not correlate with any life-history traits, at least for the question of
linear regression (Fig. \ref{fig:biological_correlates}). Therefore, it
is not possible to use these results to predict \(T_{min}\) for another
population, even if the population is of a species with a similar
life-history to one in the database used.

\subsection{Conclusions}\label{conclusions}

I used a database of 822 populations to determine the minimum time
series length required to detect population trends. This goes beyond
previous work that either focused on theoretical investigations or a
limited number of species. I show that to identify long-term changes in
abundance, on average 15.91 years of continuous monitoring are often
required (Fig. \ref{fig:min_time_dist}). However, there is wide
distribution of estimated minimum times. Therefore, it is probably not
wise to use a simple threshold number of years in monitoring design.

In line with theoretical predictions (Gerrodette 1987), I also show that
\(T_{min}\) is strongly correlated with the long-term population trend
(i.e.~rate of increase), variability in population size, and the
temporal autocorrelation (Fig. \ref{fig:correlates}). Contrary to my
initial hypotheses, minimum time required did not correlate with
generation time or any other life-history traits (Fig.
\ref{fig:biological_correlates}). This result argues against overly
simplified measures of minimum sampling time based on generation length
(Fig. \ref{fig:IUCN_analysis}).

My work implies that for many populations, short time series are
probably not reliable for detecting population trends. This result
highlights the importance of long-term monitoring programs. From both a
scientific and management perspective estimates of \(T_{min}\) are
important. If a time series is too short, we lack the statistical power
to reliably detect long-term population trends. In addition, a time
series that is too long may be a poor use of already limited funds
(Gerber, DeMaster, and Kareiva 1999). Further, more data is not always
best in situations where management actions need to be taken (Martin et
al. 2012; Martin et al. 2017). When a population trend is detected, it
may be too late for management action. In these situations, the
precautionary principle may be more appropriate (Thompson et al. 2000).
Future work should examine other species, with a wider range of
life-history characteristics. In addition, similar approaches can be
used to determine the minimum time series length required to address
additional questions of interest.

\section{Supporting Information}\label{supporting-information}

In the supporting material, I provide an expanded methods sections,
additional figures, minimum time calculations for determining
exponential growth, simulations with a more complicated population
model, and the use of generalized additive models to identify population
trends. All code and data can be found at
\url{https://github.com/erwhite1/time-series-project}

\section{Acknowledgements}\label{acknowledgements}

ERW was partially supported by a National Science Foundation Graduate
Fellowship. I would like to thank members of the Ecological Theory group
at the University of California, Davis for their insight. I would also
like to thank T. Dallas and E. Malcolm-White for their helpful comments.

\section{References}\label{references}

\hypertarget{refs}{}
\hypertarget{ref-Blanchard2007}{}
Blanchard, Julia L, David L Maxwell, and Simon Jennings. 2007. ``Power
of monitoring surveys to detect abundance trends in depleted
populations: the effects of density-dependent habitat use, patchiness,
and climate change.'' \emph{ICES Journal of Marine Science} 65 (1):
111--20.

\hypertarget{ref-Bolker2008}{}
Bolker, Benjamin M. 2008. \emph{Ecological Models and Data in R}. 1st
ed. Princeton, New Jersey: Princeton University Press.

\hypertarget{ref-Brashares2005}{}
Brashares, Justin S, and Moses K Sam. 2005. ``How much is enough?
Estimating the minimum sampling required for effective monitoring of
African reserves.'' \emph{Biodiversity and Conservation} 14: 2709--22.
doi:\href{https://doi.org/10.1007/s10531-005-8404-z}{10.1007/s10531-005-8404-z}.

\hypertarget{ref-Cohen1992}{}
Cohen, Jacob. 1992. ``A power primer.'' \emph{Psychological Bulletin}
112 (1): 155--59.
doi:\href{https://doi.org/10.1037/0033-2909.112.1.155}{10.1037/0033-2909.112.1.155}.

\hypertarget{ref-Field2007}{}
Field, Scott A, Patrick J O Connor, Andrew J Tyre, and Hugh P
Possingham. 2007. ``Making monitoring meaningful.'' \emph{Austral
Ecology} 32: 485--91.
doi:\href{https://doi.org/10.1111/j.1442-9993.2007.01715.x}{10.1111/j.1442-9993.2007.01715.x}.

\hypertarget{ref-Gerber1999}{}
Gerber, L R, D P DeMaster, and P M Kareiva. 1999. ``Gray whales and the
value of monitering data in implementing the U.S. endangered species
act.'' \emph{Conservation Biology} 13 (5): 1215--9.

\hypertarget{ref-Gerrodette1987}{}
Gerrodette, Tim. 1987. ``A power analysis for detecting trends.''
doi:\href{https://doi.org/10.2307/1939220}{10.2307/1939220}.

\hypertarget{ref-Giron-Nava2017}{}
Giron-Nava, Alfredo, Chase C James, Andrew F Johnson, David Dannecker,
Bethany Kolody, Adrienne Lee, Maitreyi Nagarkar, et al. 2017.
``Quantitative argument for long-term ecological monitoring.''
\emph{Marine Ecology Progress Series} 572: 269--74.

\hypertarget{ref-Hatch2003}{}
Hatch, S A. 2003. ``Statistical power for detecting trends with
applications to seabirds monitoring.'' \emph{Biological Conservation}
111: 317--29.

\hypertarget{ref-Hauser2006}{}
Hauser, Cindy E., Anthony R. Pople, and Hugh P. Possingham. 2006.
``Should managed populations be monitored every year?'' \emph{Ecological
Applications} 16 (2): 807--19.

\hypertarget{ref-Hughes2017}{}
Hughes, Brent B, Rodrigo Beas-luna, Allison K Barner, Kimberly Brewitt,
Daniel R Brumbaugh, Elizabeth B. Cerny-Chipman, Sarah L. Close, et al.
2017. ``Long-term studies contribute disproportionately to ecology and
policy.'' \emph{BioScience} 67 (3): 271--81.
doi:\href{https://doi.org/10.1093/biosci/biw185}{10.1093/biosci/biw185}.

\hypertarget{ref-Hutchings2012}{}
Hutchings, Jeffrey A, Ransom A Myers, Verónica B García, Luis O
Lucifora, and Anna Kuparinen. 2012. ``Life-history correlates of
extinction risk and recovery potential.'' \emph{Ecological Applications}
22 (4): 1061--7.

\hypertarget{ref-IUCN2012}{}
IUCN. 2012. ``IUCN Red List Categories and Criteria: Version 3.1.''
doi:\href{https://doi.org/10.9782-8317-0633-5}{10.9782-8317-0633-5}.

\hypertarget{ref-Johnson2015}{}
Johnson, Paul CD, Sarah JE Barry, Heather M Ferguson, and Pie Müller.
2015. ``Power analysis for generalized linear mixed models in ecology
and evolution.'' \emph{Methods in Ecology and Evolution} 6 (2): 133--42.
doi:\href{https://doi.org/10.1111/2041-210X.12306}{10.1111/2041-210X.12306}.

\hypertarget{ref-Keith2015}{}
Keith, David, H. Resit Akçakaya, Stuart H.M. Butchart, Ben Collen,
Nicholas K. Dulvy, Elizabeth E. Holmes, Jeffrey A. Hutchings, et al.
2015. ``Temporal correlations in population trends: Conservation
implications from time-series analysis of diverse animal taxa.''
\emph{Biological Conservation} 192. Elsevier B.V.: 247--57.
doi:\href{https://doi.org/10.1016/j.biocon.2015.09.021}{10.1016/j.biocon.2015.09.021}.

\hypertarget{ref-Legg2006}{}
Legg, Colin J, and Laszlo Nagy. 2006. ``Why most conservation monitoring
is, but need not be, a waste of time.'' \emph{Journal of Environmental
Management} 78: 194--99.
doi:\href{https://doi.org/10.1016/j.jenvman.2005.04.016}{10.1016/j.jenvman.2005.04.016}.

\hypertarget{ref-Magurran2010}{}
Magurran, Anne E, Stephen R Baillie, Stephen T Buckland, Jan Mcp Dick,
David A Elston, E Marian Scott, Rognvald I Smith, Paul J Somerfield, and
Allan D Watt. 2010. ``Long-term datasets in biodiversity research and
monitoring : assessing change in ecological communities through time.''
\emph{Trends in Ecology and Evolution} 25: 574--82.
doi:\href{https://doi.org/10.1016/j.tree.2010.06.016}{10.1016/j.tree.2010.06.016}.

\hypertarget{ref-Mapstone1995}{}
Mapstone, Bruce D. 1995. ``Scalable decision rules for environmental
impact studies : effect Size , type I , and type II errors.''
\emph{Ecological Applications} 5 (2): 401--10.

\hypertarget{ref-Martin2017}{}
Martin, Tara G., Abbey E. Camaclang, Hugh P. Possingham, Lynn A.
Maguire, and Iadine Chadès. 2017. ``Timing of Protection of Critical
Habitat Matters.'' \emph{Conservation Letters} 10 (3): 308--16.
doi:\href{https://doi.org/10.1111/conl.12266}{10.1111/conl.12266}.

\hypertarget{ref-Martin2012}{}
Martin, Tara G., Simon Nally, Andrew A. Burbidge, Sophie Arnall, Stephen
T. Garnett, Matt W. Hayward, Linda F. Lumsden, Peter Menkhorst, Eve
Mcdonald-Madden, and Hugh P. Possingham. 2012. ``Acting fast helps avoid
extinction.'' \emph{Conservation Letters} 5 (4): 274--80.
doi:\href{https://doi.org/10.1111/j.1755-263X.2012.00239.x}{10.1111/j.1755-263X.2012.00239.x}.

\hypertarget{ref-McCain2016}{}
McCain, Christy Marie, Tim Szewczyk, and Kevin Bracy Knight. 2016.
``Population variability complicates the accurate detection of climate
change responses.'' \emph{Global Change Biology} 22 (6): 2081--93.
doi:\href{https://doi.org/10.1111/gcb.13211}{10.1111/gcb.13211}.

\hypertarget{ref-McDonald-Madden2010}{}
McDonald-Madden, Eve, Peter W J Baxter, Richard A. Fuller, Tara G.
Martin, Edward T. Game, Jensen Montambault, and Hugh P. Possingham.
2010. ``Monitoring does not always count.'' \emph{Trends in Ecology and
Evolution} 25 (10): 547--50.
doi:\href{https://doi.org/10.1016/j.tree.2010.07.002}{10.1016/j.tree.2010.07.002}.

\hypertarget{ref-Morrison2008}{}
Morrison, Shawm, and David S. Hik. 2008. ``When? Where? And for how
long? Census design considerations for an Alpine Lagomorph, the Collared
pika.'' In \emph{Lagomorph Biology}, 103--13. Springer Berlin
Heidelberg.
doi:\href{https://doi.org/10.1007/978-3-540-72446-9}{10.1007/978-3-540-72446-9}.

\hypertarget{ref-Myhrvold2015}{}
Myhrvold, Nathan P., Elita Baldridge, Benjamin Chan, Dhileep Sivam,
Daniel L. Freeman, and S.K. Morgan Ernest. 2015. ``An amniote
life-history database to perform comparative analyses with birds,
mammals, and reptiles.'' \emph{Ecology} 96 (11): 3109.

\hypertarget{ref-GPDD2010}{}
NERC Centre for Population Biology, Imperial College. 2010. ``The Global
Population Dynamics Database Version 2.''

\hypertarget{ref-Nichols2006}{}
Nichols, James D., and Bryon K. Williams. 2006. ``Monitoring for
conservation.'' \emph{Trends in Ecology and Evolution} 21 (12): 668--73.
doi:\href{https://doi.org/10.1016/j.tree.2006.08.007}{10.1016/j.tree.2006.08.007}.

\hypertarget{ref-RCoreTeam2016}{}
R Core Team. 2016. ``R: A language and environment for statistical
computing.'' Vienna, Austria: R Foundation for Statistical Computing.
\url{https://www.r-project.org/}.

\hypertarget{ref-Rhodes2011}{}
Rhodes, Jonathan R., and Niclas Jonzen. 2011. ``Monitoring temporal
trends in spatially structured populations: how should sampling effort
be allocated between space and time?'' \emph{Ecography} 34 (6): 1040--8.
doi:\href{https://doi.org/10.1111/j.1600-0587.2011.06370.x}{10.1111/j.1600-0587.2011.06370.x}.

\hypertarget{ref-Rueda-Cediel2015}{}
Rueda-Cediel, Pamela, Kurt E Anderson, Tracey J Regan, Janet Franklin,
and M Regan. 2015. ``Combined influences of model choice, data quality,
and data quantity when estimating population trends.'' \emph{PLoSONE} 10
(7): e0132255.
doi:\href{https://doi.org/10.1371/journal.pone.0132255}{10.1371/journal.pone.0132255}.

\hypertarget{ref-Seavy2007}{}
Seavy, Nathaniel E., and Michelle H. Reynolds. 2007. ``Is statistical
power to detect trends a good assessment of population monitoring?''
\emph{Biological Conservation} 140 (1-2): 187--91.
doi:\href{https://doi.org/10.1016/j.biocon.2007.08.007}{10.1016/j.biocon.2007.08.007}.

\hypertarget{ref-Thompson2000}{}
Thompson, Paul M, Ben Wilson, Kate Grellier, and Philip S Hammond. 2000.
``Combining power analysis and population viability analysis to compare
traditional and precautionary approaches to conservation of coastal
cetaceans.'' \emph{Conservation Biology} 14 (5): 1253--63.

\hypertarget{ref-Wagner2009}{}
Wagner, Tyler, Christopher S. Vandergoot, and Jeff Tyson. 2009.
``Evaluating the power to detect temporal trends in fishery-independent
surveys - A case study based on gill nets set in the Ohio waters of Lake
Erie for walleyes.'' \emph{North American Journal of Fisheries
Management} 29: 805--16.
doi:\href{https://doi.org/10.1577/M08-197.1}{10.1577/M08-197.1}.

\hypertarget{ref-White2014}{}
White, Easton R, John D Nagy, and Samuel H Gruber. 2014. ``Modeling the
population dynamics of lemon sharks.'' \emph{Biology Direct} 9 (23):
1--18.

\hypertarget{ref-Zuur2009}{}
Zuur, Alain F., Elena N. Ieno, Neil J. Walker, Anatoly A. Saveliev, and
Graham M. Smith. 2009. \emph{Mixed Effects Models and Extensions in
Ecology with R}. New York: Springer.

\end{spacing}
\end{document}