### 2018-02-19 writing/supp_mat_draft.Rmd

---
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: true
fontsize: 12pt
geometry: margin=1in
bibliography: White_bib.bib
header-includes: 
    \usepackage{float}
    \renewcommand{\thepage}{S\arabic{page}} 
    \renewcommand{\thesection}{S\arabic{section}}  
    \renewcommand{\thetable}{A\arabic{table}}  
    \renewcommand{\thefigure}{A\arabic{figure}}
    \usepackage[round]{natbib} \bibpunct[; ]{(}{)}{,}{a}{}{;}
---

\vspace{2cm}

\begin{center}
 \textbf{Supplementary Material: Minimum time required to detect population trends: the need for long-term monitoring programs}

Easton R. White
\vspace{3 mm}

Center for Population Biology, University of California, Davis, California 95616 USA

eawhite@ucdavis.edu

 \end{center}

\vspace{2cm}

\tableofcontents

\vspace{1cm}

Data and code for all the figures and tables can be found at (https://github.com/erwhite1/time-series-project). All analyses were run using R [@RCoreTeam2016].

\vspace{2cm}

\clearpage

# Detailed example of subsampling and power calculations

```{r,echo=F}
 #
 load("../cleaned-data/cleaned_timeseries_database_with_min_time_linear_regression.Rdata")
    pop = subset(long_dat,long_dat$ID=='7558')
    model_example = lm(pop$popvalue~c(1:nrow(pop)))
```

Here, I illustrate how we performed the subsampling and power calculations for a specific population. As an example, I examine a 35-year time series of Bigeye tuna (\emph{Thunnus obesus}), one species in the Global Population Dynamics Database [@GPDD2010]. Simple linear regression indicates a significant decrease for this population with an estimated slope coefficient of `r as.numeric(coef(model_example)[2])`. I assume that this significant increase over 35 years is in fact the ``true trend". In statistical jargon, the 35-year trend is an effect that is actually present; we can reject the null hypothesis of no trend. We can then use this as a benchmark to see if subsamples of the time series also indicate a significant increase.

<!-- example of how subsampling and empirical approach works -->
```{r, echo=F,message=F,warning=F,fig.cap='(a) Population size of Bigeye tuna (\\emph{Thunnus obesus}) over time. The line is the best fit line from linear regresssion. (b) Statistical power for different subsets of the time series in panel a.\\label{fig:empirical_approach_example}'}
# Example of subsampling empirical data
 load("../cleaned-data/cleaned_timeseries_database_with_min_time_linear_regression.Rdata")
  par(mfrow=c(1,2),oma=c(4,4,0.5,0.5))
    pop = subset(long_dat,long_dat$ID=='7558') #13192 for Harp seals, 7558 for Thunnus obesus
    plot(pop$year,pop$popvalue,cex.axis=1.2,cex.lab=1.2,ylab='',xlab='time (years)',pch=16,las=1)
    abline(lm(pop$popvalue~pop$year),lwd=2,col='red')
    mtext(text = '(a)',side = 3,line = -1.2,adj=0.95,font = 2,cex=1.2)
    mtext(text='population size', side=2,line=4,cex=1.2)
# could insert all possible regressions of length n
source("../scripts/calculate_power_metric.R")
    power_example=calculate_power_for_subsamples(pop$popvalue,0.05)
    plot(power_example,cex.axis=1.2,cex.lab=1.2,ylab='statistical power',xlab='years sampled',las=1,pch=16)
    mtext(text = '(b)',side = 3,line = -1.2,adj=0.05,font = 2,cex=1.2)
# could also insert standard error bars
    abline(h=0.8,col='red',lwd=2,lty=2)
    abline(v=head(which(power_example>0.8),1),col='red',lwd=2,lty=2)

```

I then perform a subsampling routine to estimate the minimum time required $T_{min}$ (similar to @Gerber1999). This is the same routine I used for the results in the main manuscript.
\begin{enumerate}
  \item We first extract all contiguous subsamples of the time series. This leads to 34 two-year subsamples, 33 two-year subsamples, and so forth until a single 35-year subsample.  
  \item For each subsample, we conduct linear regression and extract model coefficients and p-values.    \item We can call each set of subsamples, of the same length, a set. The proportion of subsamples within a set that show significant trends (significant slope coefficient $\alpha$ less than 0.05) is the statistical power. It is important to note that we only consider subsamples to be significant if they are significant in the same direction as the complete 35-year time series. In other words, we are conducting a one-tailed test.
  \item We can then plot statistical power as a function of time series length (Fig. \ref{fig:empirical_approach_example}b). As expected, we can see that power increases with the more years that are sampled.
  \item  Then, we determine an appropriate level of statistical power that we find acceptable. Traditionally, this has been at 0.8, however, this is purely historical. Statistical power of 0.8 implies if a true trend is present, or there is a real change in abundance, then we will detect this trend 0.8 proportion of the time. 
  \item We then determine the minimum time series length ($T_{min}$) required to achieve that level of statistical power. Here, $T_{min}$ is the first point in Fig. \ref{fig:empirical_approach_example}b where following points are also above 0.8. In this example, $T_{min}$ is `r head(which(power_example>0.8),1)`. 
\end{enumerate}

Therefore, a minimum of `r head(which(power_example>0.8),1)` years of continuous monitoring are required (for 0.8 statistical power at 0.05 significance level) to determine long-term changes in abundance.

\pagebreak

# Additional results from the main manuscript

## Correlates of minimum time required

```{r poisson_model, echo=F,fig.cap='Output of generalized linear model with a Poisson error structure for predicting the minimum time required with explanatory variables of the absolute value of the slope coefficient (or trend strength), temporal autocorrelation, and coefficient of variation of population size.\\label{fig:poisson_model}'}

load("../cleaned-data/cleaned_timeseries_database_with_min_time_linear_regression.Rdata")
load("../cleaned-data/combined_databases_with_min_time_linear_regression.Rdata")

# GLM with Poisson error structure (include all terms - it is best not to include abs_overall_trend with variance as they are strongly corrrelated to one another)
pop_info$abs_overall_trend = abs(pop_info$overall_trend)
poisson_model <-glm(min_time_for_power~ abs_overall_trend + autocorrelation + coefficient_variation,family = 'poisson',data=pop_info)

par(mfrow=c(2,2),mar=c(5,5,1,2),oma=c(0,0,0.5,0))
plot(poisson_model,cex.lab=1.2,las=1,cex.axis=1.2,main=' ',pch=16, col=rgb(0.5,0.5,0.5,0.4))
#R^2 calculation (pg. 218 Zuur et al. 2009 mixed effects book)

time_series_correlates_poisson_model <-glm(min_time_for_power~ abs_overall_trend + autocorrelation + coefficient_variation,family = 'poisson',data=pop_info)
R2_time_series_correlates <- 100*(time_series_correlates_poisson_model$null.deviance - time_series_correlates_poisson_model$deviance)/(time_series_correlates_poisson_model$null.deviance)

biological_correlates_poisson_model <-glm(min_time_for_power~  litter_or_clutch_size_n  + gen_len + adult_body_mass_g + maximum_longevity_y  +incubation_d + trophic,family = 'poisson',data=LPI_pop_info)
R2_biological_correlates_poisson_model <- 100*(biological_correlates_poisson_model$null.deviance - biological_correlates_poisson_model$deviance)/(biological_correlates_poisson_model$null.deviance)

all_correlates_poisson_model <-glm(min_time_for_power~ abs(overall_trend) + coefficient_variation + autocorrelation +  litter_or_clutch_size_n  + gen_len + adult_body_mass_g + maximum_longevity_y  +incubation_d + trophic,family = 'poisson',data=LPI_pop_info)
R2_all_correlates_poisson_model <- 100*(all_correlates_poisson_model$null.deviance - all_correlates_poisson_model$deviance)/(all_correlates_poisson_model$null.deviance)

```

In the main text, I explained how the minimum time required strongly correlated with the trend strength, temporal autocorrelation, and coefficient of variation in population size. Here, I use a generalized linear model framework with a Poisson error structure to determine explanatory variables of the minimum time required. I use the same `r nrow(pop_info)` populations as in the main text. In figure \ref{fig:poisson_model}, I show a set of residual plots for the regression analysis. I then show the coefficient estimates and levels of significance in table \ref{table:model_summary_time_series_correlates}.

```{r,echo=F,message=F,warning=F}
require(knitr)
knitr::kable(summary(time_series_correlates_poisson_model)$coef,caption = 'Output of generalized linear model to examine time series characteristics as correlates of the minimum time required for determining long-term population trends.\\label{table:model_summary_time_series_correlates}')
```

```{r,echo=F,message=F,warning=F}
require(knitr)
knitr::kable(summary(biological_correlates_poisson_model)$coef,caption = 'Output of generalized linear model to examine life-history trait correlates of the minimum time required for determining long-term population trends.\\label{table:model_summary_biological_correlates}')
```

```{r,echo=F,message=F,warning=F}
require(knitr)
knitr::kable(summary(all_correlates_poisson_model)$coef,caption = 'Output of generalized linear model to examine both time series characteristics and life-history trait correlates of the minimum time required for determining long-term population trends.\\label{table:model_summary_all_correlates}')
```

The model with trend strength, autocorrelation, and the coefficient of variation in population size `r round(R2_time_series_correlates,2)`% of the variation in the minimum time required (Table \ref{table:model_summary_time_series_correlates}).

For a subset of populations, I examined life-history correlates as potential explanatory variables for the minimum time required. Table \ref{table:model_summary_biological_correlates} and \ref{table:model_summary_all_correlates} summarize results for a model with only life-history traits and an additional model with life-history traits and time series characteristics combined, respectively. The model with all life-history traits explained only `r round(R2_biological_correlates_poisson_model,2)`% of the variation in the minimum time required (Table \ref{table:model_summary_biological_correlates}). The model with all life-history traits and time series characteristics explained `r round(R2_all_correlates_poisson_model,2)`% of the variation in the minimum time required (Table \ref{table:model_summary_all_correlates}). When accounting for characteristics of the time series, no life-history variables were significant.

\pagebreak

<!--
## Minimum time required and biological correlates
```{r, echo=F}

load('../cleaned-data/combined_databases_with_min_time_linear_regression.Rdata')
```

In the main manuscript, I examined the minimum time required to detect a significant trend in abundance over time using linear regression. As detailed in the main manuscript the minimum time required was around 15, but there was a wide distribution. Therefore, I were interested in potential explanatory variables of the minimum time required. In the main manuscript, I examined characteristics of the time series itself, like variability, autocorrelation, and the trend in abundance over time. Here, I combined our time series data with a data on life history characteristics of amniotes from @Myhrvold2015. There was life history information available for `r nrow(LPI_pop_info)` populations representing `r length(table(LPI_pop_info$Binomial))` different species, all of which were birds (Aves class).

```{r, echo=F, fig.cap='Mimimum time required versus (a) generation length (years), (b) litter size (n), (c) adult body mass (grams), (d) maximum longevity (years), (e) egg mass (grams), and (f) incubation (days). The lines in each plot represent the best fit line from linear regression.\\label{fig:biological_correlates}',fig.height=6}

load('../cleaned-data/combined_databases_with_min_time_linear_regression.Rdata')
par(mfrow=c(3,2),oma=c(5,4,1,1),mar=c(5,3,0,0))

n=1
x_axis_name_list = c('generation length (years)','litter size (n)','adult body mass (grams)','maximum longevity (years)','egg mass (grams)','incubation (days)')
for (x_axis_name in names(LPI_pop_info)[21:26]){
  plot(LPI_pop_info[[x_axis_name]],LPI_pop_info$min_time_for_power,pch=16,col=rgb(0.5,0.5,0.5,0.5),las=1,ylab='',xlab='',cex.axis=1.2)
  abline(lm(LPI_pop_info$min_time_for_power~LPI_pop_info[[x_axis_name]]),lwd=3)

  mtext(text = x_axis_name_list[n],side = 1,line = 3,cex = 1.2)
  mtext(text = paste('(',letters[n],')',sep=''),side = 3,line=-1.5,adj = 0.98,cex=1.2)
  n=n+1
}

mtext(text = 'minimum time required',side = 2,line = 2,cex = 1.4,outer = TRUE)

#######

poisson_model <-glm(min_time_for_power~ gen_len + litter_or_clutch_size_n + adult_body_mass_g +maximum_longevity_y + egg_mass_g + incubation_d,family = 'poisson',data=LPI_pop_info)

#  par(mfrow=c(2,2),mar=c(5,5,1,2),oma=c(0,0,0.5,0))
#    plot(poisson_model,cex.lab=1.2,las=1,cex.axis=1.2,main=' ',pch=16, col=rgb(0.5,0.5,0.5,0.4))
#R^2 calculation (pg. 218 Zuur et al. 2009 mixed effects book)
R2_fullmodel <- 100*(poisson_model$null.deviance - poisson_model$deviance)/(poisson_model$null.deviance)
```

I then correlated minimum time required for each population with its corresponding life history characteristics. In figure \ref{fig:biological_correlates} I examined minimum time required versus generation length (years), litter size (n), adult body mass (grams), maximum longevity (years), egg mass (grams), and incubation (days). None of these variables had much explanatory power in accounting for the variance in the minimum time required. We ran a generalized linear model (with a Poisson error structure) and found that these six variables only accounted for `r round(100*(poisson_model$null.deviance - poisson_model$deviance)/(poisson_model$null.deviance),2)`% of the explained deviance in minimum time required.
-->

\pagebreak

## Minimum time series and biological class

Minimum time required differed for different biological classes. For instance, more time is required for species within the Actinopterygii class compared to other species (Fig. \ref{fig:class}a). These differences between biological classes can be explained by differences in population variability, with species in the Actinopterygii class experiencing larger inter-annual variability in population size (Fig. \ref{fig:class}c). In addition, populations in the Actinopterygii class generally had weaker trends in abundance (i.e. smaller slope coefficients) compared to the Aves and Mammalia populations.

```{r,echo=F,fig.cap='(a) Minimum time required to estimate change in abundance by biological class, (b) long-term trend (estimated slope coefficient) by class, (c) coefficient of variation in population size by class, and (d) temporal autocorrelation by class.\\label{fig:class}',fig.height=6}

# Pull in data

load("../cleaned-data/cleaned_timeseries_database_with_min_time_linear_regression.Rdata")
# Biological correlates plots
par(mfrow=c(2,2),oma=c(8,1,1.5,0.5),mar=c(0.5,4,0,1))

boxplot(pop_info$min_time_for_power~pop_info$class, lwd = 1, ylab = 'minimum time required',col='white',cex.lab=1.4,cex.axis=1.2,las=2,xaxt="n",outline=FALSE)
mtext(text = '(a)',side = 3,line = -1.2,adj=0.95,font = 2,cex=1.2)
stripchart(pop_info$min_time_for_power~pop_info$class, vertical = TRUE,method = "jitter", add = TRUE, pch = 20, col = rgb(0.5,0.5,0.5,0.5))

boxplot(abs(pop_info$overall_trend)~pop_info$class, lwd = 1, ylab = ' ',col='white',cex.lab=1.4,cex.axis=1.2,las=2,xaxt="n",outline=FALSE)
stripchart(abs(pop_info$overall_trend)~pop_info$class, vertical = TRUE,method = "jitter", add = TRUE, pch = 20, col = rgb(0.5,0.5,0.5,0.5))
  mtext('| slope coefficient |',2,line=3.5,cex=1.2)
  mtext(text = '(b)',side = 3,line = -1.2,adj=0.95,font = 2,cex=1.2)

boxplot(pop_info$coefficient_variation~pop_info$class, lwd = 1, ylab = 'coefficient of variation',col='white',cex.lab=1.4,cex.axis=1.2,las=2,outline=FALSE)
stripchart(pop_info$coefficient_variation~pop_info$class, vertical = TRUE,method = "jitter", add = TRUE, pch = 20, col = rgb(0.5,0.5,0.5,0.5))
mtext(text = '(c)',side = 3,line = -1.2,adj=0.95,font = 2,cex=1.2)

boxplot(pop_info$autocorrelation~pop_info$class, lwd = 1, ylab = 'autocorrelation',col='white',cex.lab=1.4,cex.axis=1.2,las=2,ylim=c(-0.2,1.2),outline=FALSE)
stripchart(pop_info$autocorrelation~pop_info$class, vertical = TRUE,method = "jitter", add = TRUE, pch = 20, col = rgb(0.5,0.5,0.5,0.5))
mtext(text = '(d)',side = 3,line = -1.2,adj=0.95,font = 2,cex=1.2)
```

\pagebreak

## Sensitivity analysis of significance level and power

Estimates of $T_{min}$ depend strongly on the values used for the statistical significance level ($\alpha$) and the probability of type II error ($\beta$), both of which are set by the practitioner. Again, statistical power is $1-\beta$. Here we used simulations of the model described in the main manuscript. The model simulates linear trends in population abundance. We explored how estimates of $T_{min}$ are affected by changes in each of these parameters. We see that the minimum time required increases with increases in statistical power or decreases with increases in the threshold for statistical significance (Fig \ref{fig:min_time_vs_alpha_beta}).

```{r,echo=F,message=F,error=F,fig.cap='Minimum time required to assess long-term trends in abundance for values of statistical significance ($\\alpha$) and power ($1-\\beta$).\\label{fig:min_time_vs_alpha_beta}'}
# Bring in theoretical code
# Build plot of $T_min$ vs alpha and another versus beta
# Explain results after building figures

source('../scripts/calculate_slope.R') 
source('../scripts/theoretical-models/simulate_pop.R')
# calculate_power = function(r,phi,sigma,trials,max.time,alpha){
#   trials   = trials
#   max.time = max.time
#   r        = r
#   phi      = phi
#   sigma    = sigma
#   
#   y=replicate(n=trials,simulate_pop(r,phi,sigma,max.time)) 
#   
#   power=sum((apply(y,2,calculate_p_value)<alpha) & sign(apply(y,2,calculate_slope))==sign(r))/trials 
#   
#   return(power)
# }
# 
# 
# min_time_needed= function(alpha,power){
#   years_to_sample=seq(2,30,by=1)
#   y=lapply(X=years_to_sample,FUN = calculate_power,phi =0.5,r=0.5,sigma = 1,trials=100,alpha=alpha)
#   min_time_needed = years_to_sample[tail(which(y<power),1)] + 1
#   return(min_time_needed)
# }
# 
# alpha_values=seq(0.001,0.2,by=0.01)
# power_values=seq(0.65,0.95,by=0.01)
# parameter_combinations = expand.grid(alpha_values,power_values)
# names(parameter_combinations) = c('alpha','power')
# parameter_combinations$min_time_required = NA
# 
# for (i in 1:nrow(parameter_combinations)){
# 
#   parameter_combinations$min_time_required[i] =  min_time_needed(parameter_combinations[i,1],parameter_combinations[i,2])
#  
#   #print(i)
# }

#write.csv(parameter_combinations,file = 'analysis-outputs/min_time_vs_alpha_and_power.csv',quote = F,row.names = F)

# Pull in data to prevent R from running for a long time to create data from scratch
parameter_combinations = read.csv('../analysis-outputs/min_time_vs_alpha_and_power.csv',header=T)

require(ggplot2)
ggplot(parameter_combinations, aes(x = alpha, y = power, z = min_time_required, fill=min_time_required)) +
  geom_tile() + 
  labs(title = " ", x = expression(alpha), y = expression(paste('power ',(1-beta))), fill = expression(T[min])) +
  scale_fill_distiller(palette="Spectral", na.value="white") +
  theme_bw()
```

## Additional results for IUCN analysis

In the main text, I presented results detailing the difference between the minimum time required to achieve a high level of statistical power and the number of years suggested by IUCN based on generation length to assess a species as vulnerable. There were deviations between the two approaches. These deviations were mostly explained by the differences in inter-annual population variability (coefficient of variation) and generation time (Fig. \ref{fig:IUCN_correlates}).

```{r, echo=F,message=F,warning=F,fig.cap='The difference between minimum time estimates is the minimum time required to achieve 0.8 statistical power versus the minimum time required under IUCN criteria A2 to classify a species as vulnerable. Each point represents a single population, all of which saw declines of 30% or greater over a 10 year period. (a) Difference between minimum time estimates versus the coefficient of variation in population size. (b) Difference between minimum time estimates versus the generation length in years.\\label{fig:IUCN_correlates}',fig.height=6}

library(viridis)
pop_info$color = as.numeric(as.factor(pop_info$class))
pop_info$color = viridis_pal(alpha=0.3,option='viridis')(5)[pop_info$color]
pop_info$shape = as.numeric(as.factor(pop_info$color)) + 14

declining_populations = subset(pop_info,pop_info$percent_change_10years>30 & pop_info$overall_trend<0)

par(mfrow=c(2,1),oma=c(0,5,0,0),mar=c(5,0.5,0.5,0.5))
error=(declining_populations$IUCN - declining_populations$min_time_for_power)
plot(declining_populations$coefficient_variation,error,
     ylab = '', las=1,
     xlab = 'coefficient of variation', col=declining_populations$color,
       pch=declining_populations$shape)
abline(lm(error~declining_populations$coefficient_variation),lwd=2)
mtext(text = '(a)',side = 3,line = -1.2,adj=0.05,font = 2,cex=1.2)

plot(declining_populations$gen_len,error,
     ylab = '', las=1,
     xlab = 'generation length (years)', col=declining_populations$color,
       pch=declining_populations$shape)
abline(lm(error~declining_populations$gen_len),lwd=2)
mtext(text = '(b)',side = 3,line = -1.2,adj=0.05,font = 2,cex=1.2)

mtext(text = 'Difference between minimum time estimates',side = 2,
      outer = T,at=0.6,line = 2,cex = 1.2)

legend('bottomright',legend=c('Aves','Actinopterygii','Mammalia'),pch=as.numeric(names(table(declining_populations$shape))),col=c(names(table(declining_populations$color))),bty ="n",cex=1.2,pt.cex = 1.4) 
```

\pagebreak

# Minimum time calculations testing exponential growth

In the main text, I evaluated the minimum time required to determine long-term trends in abundance via linear regression. This process examined linear trends in abundance over time. Here I examine the minimum time required to estimate long-term trends that are either exponential growth or decay. I use the same methods as described in the previous sections, but we take the $log$ of population density, or abundance.

```{r,echo=F,message=F,error=F,fig.cap='Distribution of the minimum time required in order to detect a significant trend (at the 0.05 level) in log(abundance) given power of 0.8.\\label{fig:min_time_dist_log_pop}'}

# load('../cleaned-data/cleaned_timeseries_database.Rdata')
# 
# source('../scripts/calculate_slope.R')
# source('../scripts/calculate_power_metric.R')

# save(pop_info,long_dat,long_time_names,file = '../cleaned-data/cleaned_timeseries_database_with_min_time_linear_regression_for_log_pop.Rdata')

# LOAD in previously run script from this code chuck and make histogram plot
load('../cleaned-data/cleaned_timeseries_database_with_min_time_linear_regression_for_log_pop.Rdata')

#plot(density(pop_info$min_time_for_power),las=1,ylab='',xlab='',cex.axis=1.2,main='')
#mtext(text = 'Frequency',2,line=3,cex=1.2)
#mtext(text = 'Minimum time required',1,line=3,cex=1.2)
require(ggplot2)
ggplot(aes(min_time_for_power),data=pop_info) + geom_bar() +  
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"),axis.text = element_text(size=12),axis.title.x =element_text(size=16) ,axis.title.y =element_text(size=16) ) +
 labs(y='count',x='minimum time required')

```

We see that the distribution of $T_{min}$ is almost identical to that in the main manuscript (Fig. \ref{fig:min_time_dist_log_pop}). This is perhaps not surprising as most time series that would significantly increase or decrease linearly would probably also significantly increase or decrease at an exponential rate. Further, the calculations here and in the main manuscript both use linear regression. Therefore both calculations estimate the same number of parameters.

\pagebreak

# Simulations with more complicated population model

In the main text, I showed how a simple population model could be simulated repeatedly to estimate the power obtained with time series of increasing length [@Bolker2008;@Johnson2015]. The model in the main text simulated linear population growth with only a slope coefficient, y-intercept, and noise parameter required. This model is purely phenomenological and does not include any species life-history information. Here, I use the same routine as the main text, but simulate from a more biologically-realistic population model. I use the model described in @White2014. The model is a stochastic, age-structured population model that includes density-dependence for lemon sharks (\emph{Negaprion brevirostris}) in Bimini, Bahamas.

```{r, echo=F,fig.height=3,fig.cap='Statistical power for different length of time series simulations for a lemon shark population in Bimini, Bahamas.\\label{fig:shark_example}'}

par(mfrow=c(1,2),oma=c(0.5,0.5,0.5,0.5),mar=c(4,4,0,0))
#### Simulate lemon shark model (See White et al. 2014)
total_lemons=NULL
IC= c(30,30,10,15,15,5,5,3,3,3,3,3,2,2,2,2,2,2,1,1,1,1,2,2,2,2)
no.runs=50
max.time=50
for (g in 1:(no.runs)){
#set up matrix (Lemons) to track number of individuals in each age class over time and matrix (Stable_age) to track the age distribution through time
	Lemons = matrix(IC,nrow= 26,ncol=max.time)
	Stable_age =matrix(0,nrow=26,ncol=max.time)
	#initialize matrices
	Lemons[1:26,1] = IC
	#keep track of Juveniles, Subadults, and Adults
	J =  sum(Lemons[1:3,1])    #Juveniles and newborns in lagoon
	S =  sum(Lemons[4:12,1])   #subadults who have left the lagoon and cannot reproduce
	A =  sum(Lemons[13:26,1])  #adults who can reproduce

	#initialize parameters (these are described in the paper)
	b=6.087   #Number of offspring per female 8.3 from Feldheim 2002, 6.7 from 
	h=1   	  #hill coefficient
	k=100     #half saturation constant (Estimated from Gruber 2001)
	u=0.17    #adult mortality rate, 0.15 is the default rate

		for (j in 2:(max.time)){
				Lemons[1,j]= sum(rpois(sum(Lemons[13:26,j-1])/4,b))  #use 
				Lemons[2,j] = Lemons[1,j-1] - rbinom (1, Lemons[1,j-1],(Lemons[1,j-1]^h)/(k+Lemons[1,j-1]^h))
				Lemons[3,j] = Lemons[2,j-1] - rbinom (1, Lemons[2,j-1],u) 
				Lemons[4,j] = Lemons[3,j-1] - rbinom (1, Lemons[3,j-1], u)
					for (i in 5:25){
					    Lemons[i,j] = Lemons[i-1,j-1] - rbinom (1, Lemons[i-1,j-1], u)
					}
					Lemons[26,j] = Lemons[25,j-1] - rbinom (1, Lemons[25,j-1], 1)
		          J = c(J, sum(Lemons[1:3,j]))
			 	  S = c(S, sum(Lemons[4:12,j]))
			  	  A = c(A, sum(Lemons[13:26,j]))
		}#end of time loop

total_lemons = rbind(total_lemons,colSums(Lemons))
}#end of trial loop 
matplot(t(total_lemons),type='l',col=1,lty=1,las=1,ylab='population size',xlab='time (years)',cex.lab=1.2)
mtext(text = '(a)',side = 3,line = -1.2,adj=0.95,font = 2,cex=1.2)
###
#source('../scripts/LemonShark_PopulationModel.R')
years_to_sample=seq(5,50,by=1)
#shark_power=sapply(years_to_sample,power)

load('../analysis-outputs/example_power_for_shark_model.Rdata')
   plot(years_to_sample,shark_power,type='l',lwd=2,las=1,
        ylab='',xlab='',cex.axis=1, ylim = c(0,1.05))
   abline(v=years_to_sample[tail(which(shark_power<0.8),1)+1],lty=2,lwd=2,col='red')
   abline(h=0.8,lty=2,lwd=2,col='red')
   mtext(text = 'time sampled' ,side = 1,cex = 1.2,line=3)
   mtext(text = 'statistical power' ,side = 2,cex = 1.2,line=2.5)
   mtext(text = '(b)',side = 3,line = -1.2,adj=0.95,font = 2,cex=1.2)

#save(shark_power,file='../analysis-outputs/example_power_for_shark_model.Rdata')

```

Here, I parameterize the model for a situation where adult mortality rate is high enough to cause a population decline. As in the main text, I simulated this more biologically-realistic model for different lengths of time. For each length of time, I calculated the statistical power. Similar to results with the simpler model, statistical power generally increases with longer sampling time (Fig. \ref{fig:shark_example}). In this example, the minimum time required ($T_{min}$) to obtain at least 0.8 statistical power, given a significance level of 0.05, is `r years_to_sample[tail(which(shark_power<0.8),1)+1]` years.

<!--
## Optional section looking at Gompertz model
In the main text, we showed how a simple population model could be simulated repeatedly to estimate the power obtained with time series of increasing length. The model in the main text simulated linear population growth with only a slope coefficient, y-intercept, and noise parameter required. This model is purely phenomenological and does not include any species life history. Here, we use the same routine as the main text, but simulate from a more biologically-realistic population model, the stochastic Gompertz population model:

  \begin{equation}
    N_{t+1} = N_t e^{a - b \mbox{ log} N_t + \epsilon_t}
  \end{equation}

Here, $N_t$ is the population density in year $t$, $a$ is an intercept term, and $b$ is a parameter measuring the strength of density-dependence. Lastly, $\epsilon_t$ is process error that is assumed to be normally distributed. @Knape2012 used this model to investigate the frequency of density-dependence in time series of populations.

We simulate this model repetitively for different lengths of time series. Statistical power is the proportion of time series of a set length that showed significant trends in abundance, given a true trend in abundance. *How do I determine significance here? Through linear regression again, or do I look at non-zero estimation of the parameter values?*-->

<!-- REMOVING from paper as it takes away from main message
# Minimum time required to estimate geometric growth rate

Instead of detecting a trend over time with linear regression, we could also calculate the geometric growth rate of the population. In figure \ref{fig:growth_rate}, we show how to calculate growth rates for subsamples of a time series. First, we created subsamples of each possible length from the full 35 year time series, as we did in the main next. Next, we calculated the mean and standard deviation of growth rates for each possible time series length (Fig. \ref{fig:growth_rate}b). Lastly, we calculated the percent error $\mbox{percent error} = 100 \times \left| \frac{\mbox{observed} - \mbox{theoretical}}{\mbox{theoretical}} \right|$ between the mean of each time series length (observed) and the overall population growth rate (theoretical). In Fig. \ref{fig:growth_rate}c), we show the percent error as a function of time series length. Here we define the minimum time required as the minimum number of years to achieve less than 20% error.

```{r, echo=F,eval=F, fig.cap='Example of calculating minimum time required for growth rate estimation. (a) European herring gull (\\emph{Larus argentatus}) scaled density over time, (b) mean and standard deviation of growth rate for subsamples of entire time series, and (c) the percent error between mean estimated growth rate and the true long-term growth rate. The vertical bar denotes the minimum time required to estimate growth rate within 20% error.\\label{fig:growth_rate}'}
#Example of using population growth rate instead of power
source("../scripts/calculate_geometric_growth.R")
load('../cleaned-data/cleaned_timeseries_database.Rdata')

par(mfrow=c(1,3),oma=c(0,1,0,0))

  pop = subset(long_dat,long_dat$ID=='10008') #13192 for Harp seals, 10007 is California Gull, 10008 for European herring gull (Larus argentatus)
  source('../scripts/create_subsamples.R')
  source('../scripts/percent_error.R')
  aaa=calculate_average_geometric_growth(pop$popvalue + 0.0001)
  plot(pop$year,pop$popvalue,cex.axis=1.2,cex.lab=1.2,ylab='',xlab='',pch=16,las=1,ylim=c(0,1))
    #abline(lm(pop$popvalue~pop$year),lwd=2,col='red')
    mtext(text = 'year',1,line=3,cex=1.4)
    mtext(text = 'population size',2,line=3.5,cex=1.4)
    mtext(text = '(a)',side = 3,line = -1.2,adj=0.95,font = 2,cex=1.2)

  plot(rowMeans(aaa,na.rm=T),cex.axis=1.2,pch=16,ylab='',xlab='',las=1,ylim=c(0.8,1))
    mtext(text = 'years sampled',1,line=3,cex=1.4)
    mtext(text = 'population growth rate',2,line=3.7,cex=1.4)

    abline(h= -0.01*rowMeans(aaa,na.rm = T)[34] +rowMeans(aaa,na.rm = T)[34] )
    abline(h= 0.01*rowMeans(aaa,na.rm = T)[34] +rowMeans(aaa,na.rm = T)[34] )

    #arrows(1:34,rowMeans(aaa,na.rm=T),1:34,rowMeans(aaa,na.rm=T)+sd(apply(aaa,MARGIN = 1,FUN=var,na.rm=T),na.rm=T),angle = 90,length = 0.1)
    #arrows(1:34,rowMeans(aaa,na.rm=T),1:34,rowMeans(aaa,na.rm=T)-sd(apply(aaa,MARGIN = 1,FUN=var,na.rm=T),na.rm=T),angle = 90,length = 0.1)
    mtext(text = '(b)',side = 3,line = -1.2,adj=0.95,font = 2,cex=1.2)

    plot(percent_error(rowMeans(aaa,na.rm=T),rowMeans(aaa,na.rm=T)[34]),cex.axis=1.2,pch=16,ylab='',xlab='',las=1, ylim = c(0,10))
    mtext(text = 'years sampled',1,line=3,cex=1.4)
    mtext(text = 'percent error in growth rate',2,line=3.7,cex=1.4)
    abline(h=0.2,lty=2)
    abline(v=tail(which(percent_error(rowMeans(aaa,na.rm=T),rowMeans(aaa,na.rm=T)[34])>0.1),1),lty=2)
    mtext(text = '(c)',side = 3,line = -1.2,adj=0.95,font = 2,cex=1.2)
```

We applied the same calculations to `r nrow(pop_info)` population time series. We then obtain a distribution of the minimum time required to measure the ``true" long-term growth rate (Fig. \ref{fig:min_time_growth_dist}). We see a bimodal distribution with many populations required 30+ years to estimate the long-term growth rate. The large number of short years required is due to cases where the entire time series is consistently increasing or decreasing at the same rate.

```{r, echo=F,eval=F,fig.cap='Histogram of the minimum time required in order to estimate the long-term growth rate within 20% error.\\label{fig:min_time_growth_dist}'}
# Growth rates calculations for each species
load('../cleaned-data/cleaned_timeseries_database.Rdata')
# 
# source('../scripts/create_subsamples.R')
# source('../scripts/calculate_geometric_growth.R')
# source('../scripts/percent_error.R')
#  for (j in 1:nrow(pop_info)){
#     pop= subset(long_dat,long_dat$ID==pop_info$ID[j])
#     pop$popvalue=as.numeric(pop$popvalue)
#     pop$popvalue=(pop$popvalue - min(pop$popvalue))/(max(pop$popvalue) - min(pop$popvalue))
#     pop$popvalue = pop$popvalue + 0.001
# 
#    # source('../scripts/calculate_power_metric.R')
#     aaa=calculate_average_geometric_growth(pop$popvalue)
#     if (length(which(percent_error(rowMeans(aaa,na.rm=T),rowMeans(aaa,na.rm=T)[34])>0.2))==0){
#       pop_info$min_time_growth_rate[j] = 2
#     }else{
#       pop_info$min_time_growth_rate[j] = tail(which(percent_error(rowMeans(aaa,na.rm=T),rowMeans(aaa,na.rm=T)[34])>0.2),1)+1}
#     print(paste(j,':',pop_info$min_time_for_power[j],sep=' '))
#  }
# 
# save(pop_info,file = '../analysis-outputs/min_time_for_growth_rate0_2.Rdata')

load('../analysis-outputs/min_time_for_growth_rate0_2.Rdata')

# Create plot
#par(mfrow=c(1,1))
#hist(pop_info$min_time_growth_rate,las=1,ylab='',xlab='',cex.axis=1.2,main='',breaks=20,xlim=c(0,40),ylim=c(0,500))
#mtext(text = 'frequency',2,line=3,cex=1.2)
#mtext(text = 'minimum time required',1,line=3,cex=1.2)

ggplot(aes(min_time_growth_rate),data=pop_info) + geom_bar() +  
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"),axis.text = element_text(size=12),axis.title.x =element_text(size=16) ,axis.title.y =element_text(size=16) ) +
 labs(y='count',x='minimum time required')

```

\clearpage 
-->

\pagebreak

# Using Generalized additive model to identify significant trends

In the main text, I examined the minimum time required to identify a trend in abundance via linear regression. This approach allowed us to identify increases or decreases, but a linear model may not always be a good fit. Generalized additive models (GAMs) are more general than general linear models and allow more flexibility [@Wood2006]. GAMs are models where a response variable depends on unknown smooth functions of explanatory variables. GAMs, therefore, can identify relationships between response variables and explanatory variables that are non-linear and perhaps more complicated. The downside of GAMs is they typically require more data and are also prone to overfitting.

Here, I conduct the same analyses in the main text, but instead calculate the minimum time series required to detect trends over time according to a GAM model. I hypothesized that GAMs should require less time to detect a trend as they are more flexible than linear regression. I provide an example that shows statistical power increases with more time sampled (Fig. \ref{fig:gam_example}).

```{r,echo=F,message=F,warning=F,fig.cap='(a) Time series for Bigeye tuna (\\emph{Thunnus obesus}) with corresponding fitted GAM model in red (with a smoothing parameter of 3) and (b) statistical power as a function of the number of years sampled. The horizontal ine at 0.8 indicates the minimum threshold for statistical power and the vertical line denotes the minimum time required to achieve 0.8 statistical power.\\label{fig:gam_example}'}
# show gam example side-by-side plot

 load("../cleaned-data/cleaned_timeseries_database_with_min_time_linear_regression.Rdata")
 source('../scripts/calculate_power_metric_gam.R')
  par(mfrow=c(1,2),oma=c(4,4,0.5,0.5))
    pop = subset(long_dat,long_dat$ID=='7558') #13192 for Harp seals, 7558 for Thunnus obesus
    pop$popvalue=(pop$popvalue - min(pop$popvalue))/(max(pop$popvalue) - min(pop$popvalue)) 
    plot(pop$year,pop$popvalue,cex.axis=1.2,cex.lab=1.2,ylab='',xlab='time (years)',pch=16,las=1)
    example_gam_model=gam(pop$popvalue ~ s(pop$year,k=3),family=gaussian)
    points(pop$year,example_gam_model$fitted.values,col='red',type='l',lwd=2)
    mtext(text = '(a)',side = 3,line = -1.2,adj=0.95,font = 2,cex=1.2)
    mtext(text='population size', side=2,line=4,cex=1.2)
# could insert all possible regressions of length n

    power_example=calculate_power_for_subsamples(pop$popvalue,0.05)
    plot(power_example,cex.axis=1.2,cex.lab=1.2,ylab='statistical power',xlab='years sampled',las=1,pch=16)
    mtext(text = '(b)',side = 3,line = -1.2,adj=0.05,font = 2,cex=1.2)
# could also insert standard error bars
    abline(h=0.8,col='red',lwd=2,lty=2)
    abline(v=head(which(power_example>0.8),1),col='red',lwd=2,lty=2)
```

```{r,echo=F,fig.cap='Distribution of the minimum time required in order to detect a significant trend (at the 0.05 level) in abundance according to a GAM model given statistical power of 0.8. The smoothing parameter was set to 3 for each population.\\label{fig:min_time_dist_gam}'}
# show gam results for minimium time required
load("../cleaned-data/cleaned_timeseries_database_with_min_time_linear_regression.Rdata")
main_text = pop_info
# load('../cleaned-data/cleaned_timeseries_database.Rdata')
#  source('../scripts/calculate_power_metric_gam.R')
#  for (j in 1:nrow(pop_info)){
#     pop= subset(long_dat,long_dat$ID==pop_info$ID[j])
#     pop$popvalue=as.numeric(pop$popvalue)
#     pop$popvalue=(pop$popvalue - min(pop$popvalue))/(max(pop$popvalue) - min(pop$popvalue))
# 
#     pop$trend_p_value = calculate_p_value_gam(pop$popvalue)
# 
#     pop_info$min_time_for_power_gam[j] = min_time_needed(pop$popvalue,0.05,0.8)
#     print(paste(j,':',pop_info$min_time_for_power[j],sep=' '))
#  }
# 
# #pop_info$min_time_for_power_gam[pop_info$min_time_for_power_gam==-999]=33
# 
# pop_info=pop_info[pop_info$trend_p_value<0.05,]
# long_dat=subset(long_dat,long_dat$ID %in% pop_info$ID)
# long_time_names=subset(long_time_names,long_time_names %in% pop_info$ID)
# 
# save(pop_info,long_dat,long_time_names,file='../analysis-outputs/min_time_required_GAM.Rdata')

load('../analysis-outputs/min_time_required_GAM.Rdata')
# Create plot
#par(mfrow=c(1,1))
pop_info = subset(pop_info,pop_info$min_time_for_power_gam!=-999)
#plot(density(min_time_gam),las=1,ylab='',xlab='',cex.axis=1.2,main='',xlim=c(0,40))
#mtext(text = 'Frequency',2,line=3,cex=1.2)
#mtext(text = 'Minimum time required',1,line=3,cex=1.2)
require(ggplot2)
ggplot(aes(min_time_for_power_gam),data=pop_info) + geom_bar() +  
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"),axis.text = element_text(size=12),axis.title.x =element_text(size=16) ,axis.title.y =element_text(size=16) ) +
 labs(y='count',x='minimum time required')
```

We then fit GAM models for `r nrow(pop_info)` populations. We found a similar distribution of minimum time required as in the main text for linear regression (Fig. \ref{fig:min_time_dist_gam}). However, in line with our hypothesis, the GAM models did result in a lower mean minimum time required of `r round(mean(pop_info$min_time_for_power_gam),2)` years compared to the results from the main text of `r round(mean(main_text$min_time_for_power),2)` years.

<!--
## Minimum time to detect density-dependence in time series

For decades, the importance of density-dependence in natural populations has been an important question in ecology (citations). To address this question, @Knape2012 used data from the Global Population Dynamics Database [@GPDD2010]. They fitted 627 time series with a Gompertz population model and found evidence for density-dependence in 45% of the time series. Each time series in their database contained at least 15 data points, and ranged from 15 to 157. A follow up question to their work would be, what time series length is required to confidently estimate density-dependence. 
-->

\clearpage

# References